### 2024-01-25 src/tex/ms.tex
[showyourwork] first commit

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

% Begin!
\begin{document}

% Title
\title{An open source scientific article}

% Author list
\author{@lmorabit}

% Abstract with filler text
\begin{abstract}
    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
    Phasellus eu tellus sit amet tortor gravida placerat.
    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
    Praesent eget sem vel leo ultrices bibendum.
    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
    Curabitur auctor semper nulla donec varius orci eget risus.
    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
    Duis eget orci sit amet orci dignissim rutrum.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
Donec vehicula augue eu neque, morbi tristique senectus et netus et.
Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
Phasellus eu tellus sit amet tortor gravida placerat.
Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
Praesent eget sem vel leo ultrices bibendum.
Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
Curabitur auctor semper nulla donec varius orci eget risus.
Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
Pellentesque cursus luctus mauris \citep{Luger2021}.

\bibliography{bib}

\end{document}

### 2024-01-25 src/scripts/testplot.py
testing

@@ -1,0 +1,30 @@
+import paths
+import numpy as np
+import matplotlib
+from matplotlib import pyplot as plt
+from astropy.table import Table, Column, join
+## WMAP9 is Hinshaw et al. 2013, H_0=69.3, Omega=0.287
+
+si = -0.8
+
+## read in SIMBA
+Simba_SF = Table.read( 'RLFS_50MYR_SF.csv', format='csv', delimiter=',' )
+Simba_AGN = Table.read( 'RLFS_50MYR_AGN.csv', format='csv', delimiter=',' )
+## shift using spectral index
+Simba_SF['x']  = Simba_SF['x'] + np.log10( np.power( (144./1400.), si ) )
+Simba_AGN['x']  = Simba_AGN['x'] + np.log10( np.power( (144./1400.), si ) )
+
+fig = plt.figure( figsize=(5,5) )
+## plot simba
+plt.plot( Simba_SF['x'], Simba_SF['Curve1'], color='red', linewidth=2.5, label='Simba SF' )
+plt.plot( Simba_AGN['x'], Simba_AGN['Curve2'], color='blue', linewidth=2.5, label='Simba AGN' )
+plt.xlim((20,28))
+plt.ylim(-7.5,-2)
+plt.xlabel('log('+r'$L_{\mathrm{144 MHz}}$'+' W Hz'+r'$^{-1}$'+'])')
+plt.ylabel('log('+r'$\Phi$'+' [mag'+r'$^{-1}$'+' Mpc'+r'$^{-3}$'+'])')
+plt.legend()
+plt.savefig(paths.figures / 'test.png',dpi=300)
+fig.clear()
+
+
+

### 2024-01-25 src/scripts/testplot.py
added data paths

@@ -8,8 +8,8 @@ from astropy.table import Table, Column, join
 si = -0.8
 
 ## read in SIMBA
-Simba_SF = Table.read( 'RLFS_50MYR_SF.csv', format='csv', delimiter=',' )
-Simba_AGN = Table.read( 'RLFS_50MYR_AGN.csv', format='csv', delimiter=',' )
+Simba_SF = Table.read( paths.data / 'RLFS_50MYR_SF.csv', format='csv', delimiter=',' )
+Simba_AGN = Table.read( paths.data / 'RLFS_50MYR_AGN.csv', format='csv', delimiter=',' )
 ## shift using spectral index
 Simba_SF['x']  = Simba_SF['x'] + np.log10( np.power( (144./1400.), si ) )
 Simba_AGN['x']  = Simba_AGN['x'] + np.log10( np.power( (144./1400.), si ) )

### 2024-01-25 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Define document class
-\documentclass[twocolumn]{aastex631}
+\documentclass[usenatbib,fleqn]{mnras}
 \usepackage{showyourwork}
+\usepackage{graphicx}
+\usepackage{amssymb,amsmath}
+\usepackage{color}
+\usepackage{multirow}
+\usepackage{soul}
+\usepackage{orcidlink}
+
+\newcommand{\sarc}{$^{\prime\prime}\!\!$}
+\newcommand{\myreferences}{references.bib}
+\newcommand{\frii}{FR$\,$II}
+\newcommand{\fri}{FR$\,$I}
+\newcommand{\wphz}{$\,$W$\,$Hz$^{-1}$}
+\newcommand{\tbpeak}{$T_{b,\mathrm{peak}}$}
+\newcommand{\tbtotal}{$T_{b,\mathrm{total}}$}
+
+% Title
+\title[Running head]{An open source scientific article}
+
+% Author list
+\author[L.K. Morabito]{\parbox{\textwidth}{Leah K. Morabito$^{1,2}$\thanks{E-mail: leah.k.morabito@durham.ac.uk}\orcidlink{0000-0003-0487-6651}\\}\\ 
+$^{1}$Centre for Extragalactic Astronomy, Department of Physics, Durham University, Durham DH1 3LE, UK \\
+$^{2}$Institute for Computational Cosmology, Department of Physics, University of Durham, South Road, Durham DH1 3LE, UK \\ }
 
@@
 
-% Title
-\title{An open source scientific article}
+\date{}
+\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2021}
+\maketitle
 
-% Author list
-\author{@lmorabit}
+\label{firstpage}
 
@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{figures/test.png}
+    \caption{Caption}
+    \label{fig:enter-label}
+    \script{testplot.py}
+\end{figure}
+
 \bibliography{bib}

### 2024-01-30 src/scripts/mauchsadler.py
initial mauchsadler

@@ -1,0 +1,31 @@
+import paths
+import numpy as np
+import matplotlib
+from matplotlib import pyplot as plt
+from astropy.table import Table, Column, join
+## WMAP9 is Hinshaw et al. 2013, H_0=69.3, Omega=0.287
+
+si = -0.8
+
+## read in SIMBA
+mauch_sadler = Table.read( paths.static / 'mauch_sadler_table5.csv', format='csv', delimiter=',' )
+## shift using spectral index
+Simba_SF['x']  = Simba_SF['x'] + np.log10( np.power( (144./1400.), si ) )
+ms_144MHz = mauch_sadler['log10_P1p4GHz'] + np.log10( np.power( (144./1400.), si ) ) 
+
+
+fig = plt.figure( figsize=(5,5) )
+## plot
+plt.plot( ms_144MHz, mauch_sadler['log10RLF_all'], color='black', linewidth=2.5, label='All' )
+plt.plot( ms_144MHz, mauch_sadler['log10RLF_SF'], color='magenta', linewidth=2.5, label='SFG' )
+plt.plot( ms_144MHz, mauch_sadler['log10RLF_AGN'], color='orange', linewidth=2.5, label='AGN' )
+plt.xlim((20,28))
+plt.ylim(-7.5,-2)
+plt.xlabel('log('+r'$L_{\mathrm{144 MHz}}$'+' W Hz'+r'$^{-1}$'+'])')
+plt.ylabel('log('+r'$\Phi$'+' [mag'+r'$^{-1}$'+' Mpc'+r'$^{-3}$'+'])')
+plt.legend()
+plt.savefig(paths.figures / 'mauch_sadler_RLFs.png',dpi=300)
+fig.clear()
+
+
+

### 2024-01-30 src/scripts/mauchsadler.py
remove simba ref

@@ -10,7 +10,6 @@ si = -0.8
 ## read in SIMBA
 mauch_sadler = Table.read( paths.static / 'mauch_sadler_table5.csv', format='csv', delimiter=',' )
 ## shift using spectral index
-Simba_SF['x']  = Simba_SF['x'] + np.log10( np.power( (144./1400.), si ) )
 ms_144MHz = mauch_sadler['log10_P1p4GHz'] + np.log10( np.power( (144./1400.), si ) ) 
 
 

### 2024-01-30 src/scripts/mauchsadler.py
colname

@@ -17,7 +17,7 @@ fig = plt.figure( figsize=(5,5) )
 ## plot
 plt.plot( ms_144MHz, mauch_sadler['log10RLF_all'], color='black', linewidth=2.5, label='All' )
 plt.plot( ms_144MHz, mauch_sadler['log10RLF_SF'], color='magenta', linewidth=2.5, label='SFG' )
-plt.plot( ms_144MHz, mauch_sadler['log10RLF_AGN'], color='orange', linewidth=2.5, label='AGN' )
+plt.plot( ms_144MHz, mauch_sadler['log10RLF_RLAGN'], color='orange', linewidth=2.5, label='AGN' )
 plt.xlim((20,28))
 plt.ylim(-7.5,-2)
 plt.xlabel('log('+r'$L_{\mathrm{144 MHz}}$'+' W Hz'+r'$^{-1}$'+'])')

### 2024-01-30 src/scripts/mauchsadler.py
changes

@@ -3,15 +3,45 @@ import numpy as np
 import matplotlib
 from matplotlib import pyplot as plt
 from astropy.table import Table, Column, join
+from helper_functions import *
+from astropy.io import fits
 ## WMAP9 is Hinshaw et al. 2013, H_0=69.3, Omega=0.287
 
+## set a default spectral index
 si = -0.8
 
-## read in SIMBA
+## read in Mauch & Sadler Table 5
 mauch_sadler = Table.read( paths.static / 'mauch_sadler_table5.csv', format='csv', delimiter=',' )
 ## shift using spectral index
 ms_144MHz = mauch_sadler['log10_P1p4GHz'] + np.log10( np.power( (144./1400.), si ) ) 
 
+## read in lofar data
+lotss = Table.read( paths.static / 'lockman_final_cross_match_catalogue-v1.0.fits', format='fits' )
+## flux cut
+flux_cut = 150e-6 ## 150 uJy
+valid = np.where( np.logical_or(lotss['Total_flux'] > flux_cut, lotss['Peak_flux'] > flux_cut) )[0]
+lotss = lotss[valid]
+## remove things with no redshifts
+
+## for testing
+#good_z = np.where( np.isfinite(lotss['Z_BEST']) )[0]
+good_z = np.where(lotss['Z_BEST'] < 0.5 )[0]
+lotss = lotss[good_z]
+
+## get z-max
+lotss_zmax = RLF_calculate_zmax( flux_cut, lotss['Source_Name'], lotss['Total_flux'], lotss['E_Total_flux'], lotss['Z_BEST'], np.repeat(si,len(lotss)), outfile= paths.data / 'zmaxes' )
+
+
+
+
+## Total area
+rms_image = fits.open( paths.static / 'lockman_rms_starmask_optical.fits' )
+rms = rms_image[0].data
+rms_idx = np.where(rms <= flux_cut)
+npix_total = len(rms_idx[0])
+pixscale = np.abs(rms_image[0].header['CDELT1'])
+pixarea = np.power(pixscale, 2.)
+area_deg2 = npix_total * pixarea
 
 fig = plt.figure( figsize=(5,5) )
 ## plot

### 2024-01-30 src/scripts/helper_functions.py
update helpers

@@ -1,0 +1,96 @@
+import numpy as np
+from astropy.cosmology import WMAP9 as cosmo
+from astropy.table import Table
+import astropy.units as u
+from progress.bar import Bar
+import paths
+
+def RLF_calculate_zmax( flux_limit, source_id, fluxes, flux_errors, redshifts, spectral_indices, outfile='' ):
+
+	##########################################
+	## this function has as input:
+	## flux_limit -- survey flux limit in Jy
+	## source_id -- a unique source identifier
+	## fluxes -- source flux densities in Jy
+	## flux_errors -- source flux density errors in Jy
+	## redshifts -- source redshifts
+	## spectral_indices -- source spectral indices
+	## outfile -- the outfile name
+	##
+	## the output is the maximum redshift to which the source could be seen (with errors)
+
+    ## convert from Jy to W/Hz/m^2
+    flux_limit = flux_limit * 1e-26
+    fluxes = fluxes * 1e-26
+    flux_errors = flux_errors * 1e-26
+
+    ## convert to intrinsic power 
+    tmp = cosmo.luminosity_distance(redshifts)
+    DL = tmp.to(u.m).value
+    power = 4. * np.pi * np.power( DL, 2. ) * fluxes / np.power( (1.+redshifts), (1+spectral_indices) )
+    power_error = 4. * np.pi * np.power( DL, 2. ) * flux_errors / np.power( (1.+redshifts), (1+spectral_indices) )
+
+    ## limiting distance will be when 
+    ## P_rest = 4*pi*flux_limit * D_L(z_limit)^2 / ( 1 + z_limit )
+    ## or when D_L(z_limit)^2 / ( 1 + z_limit ) = P_rest / ( 4 * pi * flux_limit )
+    galaxy_values = power / flux_limit / ( 4. * np.pi )
+    galaxy_values_err = power_error / flux_limit / ( 4. * np.pi )
+    redshift_range = np.arange( 0.001, 10.001, 0.001 )
+    redshift_range2 = np.arange( 9, 30.001, 0.001 )
+    tmp = cosmo.luminosity_distance(redshift_range)
+    V_DL = tmp.to(u.m).value
+    tmp = cosmo.luminosity_distance(redshift_range2)
+    V_DL2 = tmp.to(u.m).value
+    
+    z_limits = []
+    z_limits_up = []
+    z_limits_lo = []
+
+    pb = Bar('Processing', max=len(galaxy_values))
+    for i in np.arange(0,len(galaxy_values)):
[…]

### 2024-01-30 src/scripts/mauchsadler.py

@@ -29,7 +29,7 @@ good_z = np.where(lotss['Z_BEST'] < 0.5 )[0]
 lotss = lotss[good_z]
 
 ## get z-max
-lotss_zmax = RLF_calculate_zmax( flux_cut, lotss['Source_Name'], lotss['Total_flux'], lotss['E_Total_flux'], lotss['Z_BEST'], np.repeat(si,len(lotss)), outfile= paths.data / 'zmaxes' )
+lotss_zmax = RLF_calculate_zmax( flux_cut, lotss['Source_Name'], lotss['Total_flux'], lotss['E_Total_flux'], lotss['Z_BEST'], np.repeat(si,len(lotss)), outfile='zmaxes' )
 
 
 

### 2024-01-30 src/scripts/helper_functions.py
added lofar data

@@ -4,6 +4,15 @@ from astropy.table import Table
 import astropy.units as u
 from progress.bar import Bar
 import paths
+from astropy.stats import median_absolute_deviation as apy_mad
+
+def radio_power( obs_flux_Jy, redshift, spectral_index=-0.8):
+    flux_cgs = obs_flux_Jy * 1e-23
+    DL = cosmo.luminosity_distance(redshift).value * 3.086e24 ## convert to cm
+    power_cgs = 4. * np.pi * np.power(DL,2.) * flux_cgs / np.power( 1+redshift, 1.+spectral_index )
+    ## convert to W/Hz
+    power = power_cgs * 1e-7
+    return( power )
 
 def RLF_calculate_zmax( flux_limit, source_id, fluxes, flux_errors, redshifts, spectral_indices, outfile='' ):
 
@@ -91,6 +100,94 @@ def RLF_calculate_zmax( flux_limit, source_id, fluxes, flux_errors, redshifts, s
 
     outfits = outfile + '.fits'
 
-    t.write( paths.static / outfits , format='fits', overwrite=True )
+    t.write( paths.data / outfits , format='fits', overwrite=True )
     return( t )
+
+def RLF_from_zmax( Lum, zz, lum_bins, redshift_bins, myzmax, area_cov, area_units='deg2', error_type='rms' ):
+    ## error_type can be "data" or "rms"
+    ## "data" will use measured uncertainties to calculate upper and lower values and propagate the errors
+    ## "rms" will use the estimator from Marshall 1985 and is equal to
+    ## 1/(delta log L) * sqrt( sum( 1/vmax^2 )
+
+    ## get log10 of the values
+    log10_L_total = np.log10( Lum )
+
+    ## create an array for the RLF
+    lum_func = np.zeros( (len(lum_bins),len(redshift_bins)) )
+    lum_func_up = np.copy(lum_func)
+    lum_func_lo = np.copy(lum_func)
+    ## median luminosities
+    med_lum = np.copy( lum_func )
+    med_lum_err = np.copy( lum_func )
+    ## number of objects
+    N_obj = np.copy( lum_func )
+    
+    if area_units == 'deg2':
+        ## area is in sqare degrees, convert to sr
+        area_sr = area_cov * np.power( ( np.pi / 180. ), 2. )
+        area_val = area_sr / ( 4. * np. pi )
+        ## to avoid evaluating if statements every loop
+        area_scaling = np.repeat( area_val, len(redshift_bins) )
[…]

### 2024-01-30 src/scripts/mauchsadler.py

@@ -22,17 +22,11 @@ flux_cut = 150e-6 ## 150 uJy
 valid = np.where( np.logical_or(lotss['Total_flux'] > flux_cut, lotss['Peak_flux'] > flux_cut) )[0]
 lotss = lotss[valid]
 ## remove things with no redshifts
-
 ## for testing
 #good_z = np.where( np.isfinite(lotss['Z_BEST']) )[0]
+zmax = 0.5
 good_z = np.where(lotss['Z_BEST'] < 0.5 )[0]
 lotss = lotss[good_z]
-
-## get z-max
-lotss_zmax = RLF_calculate_zmax( flux_cut, lotss['Source_Name'], lotss['Total_flux'], lotss['E_Total_flux'], lotss['Z_BEST'], np.repeat(si,len(lotss)), outfile='zmaxes' )
-
-
-
 
 ## Total area
 rms_image = fits.open( paths.static / 'lockman_rms_starmask_optical.fits' )
@@ -43,11 +37,23 @@ pixscale = np.abs(rms_image[0].header['CDELT1'])
 pixarea = np.power(pixscale, 2.)
 area_deg2 = npix_total * pixarea
 
+## get z-max
+lotss_zmax = RLF_calculate_zmax( flux_cut, lotss['Source_Name'], lotss['Total_flux'], lotss['E_Total_flux'], lotss['Z_BEST'], np.repeat(si,len(lotss)), outfile='zmaxes' )
+
+## calculate RLFs
+redshift_bins = np.arange( 0.0, zmax+0.001,0.5 )
+lum_bins = np.arange( 19.8, 27, 0.4 ) + np.log10( np.power( (144./1400.), si ) )
+Lrad = radio_power( lotss['Total_flux'], lotss['Z_BEST'], spectral_index=si )
+RLF, RLF_up, RLF_lo, Lmed, Lmed_err, N_obj = RLF_from_zmax( Lrad, lotss['Z_BEST'], lum_bins, redshift_bins, lotss_zmax, area_deg2, area_units='deg2', error_type='rms' )
+
+
 fig = plt.figure( figsize=(5,5) )
 ## plot
 plt.plot( ms_144MHz, mauch_sadler['log10RLF_all'], color='black', linewidth=2.5, label='All' )
 plt.plot( ms_144MHz, mauch_sadler['log10RLF_SF'], color='magenta', linewidth=2.5, label='SFG' )
[…]

### 2024-01-30 src/scripts/mauchsadler.py
change to points

@@ -53,7 +53,7 @@ plt.plot( ms_144MHz, mauch_sadler['log10RLF_all'], color='black', linewidth=2.5,
 plt.plot( ms_144MHz, mauch_sadler['log10RLF_SF'], color='magenta', linewidth=2.5, label='SFG' )
 plt.plot( ms_144MHz, mauch_sadler['log10RLF_RLAGN'], color='orange', linewidth=2.5, label='AGN' )
 ## plot the lofar data
-plt.plot( Lmed, RLF[:,1], color='purple', label='data' )
+plt.plot( Lmed, RLF[:,1], 'o', color='purple', label='data' )
 plt.xlim((20,28))
 plt.ylim(-7.5,-2)
 plt.xlabel('log('+r'$L_{\mathrm{144 MHz}}$'+' W Hz'+r'$^{-1}$'+'])')

### 2024-01-30 src/scripts/mauchsadler.py
fill between

@@ -53,6 +53,7 @@ plt.plot( ms_144MHz, mauch_sadler['log10RLF_all'], color='black', linewidth=2.5,
 plt.plot( ms_144MHz, mauch_sadler['log10RLF_SF'], color='magenta', linewidth=2.5, label='SFG' )
 plt.plot( ms_144MHz, mauch_sadler['log10RLF_RLAGN'], color='orange', linewidth=2.5, label='AGN' )
 ## plot the lofar data
+plt.fill_between( Lmed, RLF_lo[:,1], RLF_up[:,1], color='purple', alpha=0.5 )
 plt.plot( Lmed, RLF[:,1], 'o', color='purple', label='data' )
 plt.xlim((20,28))
 plt.ylim(-7.5,-2)

### 2024-01-30 src/scripts/mauchsadler.py
flatten

@@ -53,7 +53,7 @@ plt.plot( ms_144MHz, mauch_sadler['log10RLF_all'], color='black', linewidth=2.5,
 plt.plot( ms_144MHz, mauch_sadler['log10RLF_SF'], color='magenta', linewidth=2.5, label='SFG' )
 plt.plot( ms_144MHz, mauch_sadler['log10RLF_RLAGN'], color='orange', linewidth=2.5, label='AGN' )
 ## plot the lofar data
-plt.fill_between( Lmed, RLF_lo[:,1], RLF_up[:,1], color='purple', alpha=0.5 )
+plt.fill_between( Lmed, RLF_lo[:,1].flatten(), RLF_up[:,1].flatten(), color='purple', alpha=0.5 )
 plt.plot( Lmed, RLF[:,1], 'o', color='purple', label='data' )
 plt.xlim((20,28))
 plt.ylim(-7.5,-2)

### 2024-02-02 src/scripts/mauchsadler.py
filter zeros

@@ -23,7 +23,9 @@ fig = plt.figure( figsize=(5,5) )
 plt.plot( ms_144MHz, mauch_sadler['log10RLF_all'], color='black', linewidth=2.5, label='All' )
 plt.plot( ms_144MHz, mauch_sadler['log10RLF_SF'], color='magenta', linewidth=2.5, label='SFG' )
 plt.plot( ms_144MHz, mauch_sadler['log10RLF_RLAGN'], color='orange', linewidth=2.5, label='AGN' )
-## plot the lofar data
+## plot the lofar data, filtering zeros
+non_zero = np.where( RLF['RLF'] != 0.0 )[0]
+RLF = RLF[non_zero]
 plt.fill_between( RLF['Lmedian'], RLF['RLF_lo'], RLF['RLF_up'], color='purple', alpha=0.5 )
 plt.plot( RLF['Lmedian'], RLF['RLF'], 'o', color='purple', label='data' )
 plt.xlim((20,28))

### 2024-02-02 src/scripts/mauchsadler.py
green for corentin

@@ -26,8 +26,8 @@ plt.plot( ms_144MHz, mauch_sadler['log10RLF_RLAGN'], color='orange', linewidth=2
 ## plot the lofar data, filtering zeros
 non_zero = np.where( RLF['RLF'] != 0.0 )[0]
 RLF = RLF[non_zero]
-plt.fill_between( RLF['Lmedian'], RLF['RLF_lo'], RLF['RLF_up'], color='purple', alpha=0.5 )
-plt.plot( RLF['Lmedian'], RLF['RLF'], 'o', color='purple', label='data' )
+plt.fill_between( RLF['Lmedian'], RLF['RLF_lo'], RLF['RLF_up'], color='green', alpha=0.5 )
+plt.plot( RLF['Lmedian'], RLF['RLF'], 'o', color='green', label='data' )
 plt.xlim((20,28))
 plt.ylim(-7.5,-2)
 plt.xlabel('log('+r'$L_{\mathrm{144 MHz}}$'+' W Hz'+r'$^{-1}$'+'])')

### 2024-02-02 src/scripts/generateZmaxes.py
lum bins

@@ -42,10 +42,10 @@ area_deg2 = npix_total * pixarea
 lotss_zmax = RLF_calculate_zmax( flux_cut, lotss['Source_Name'], lotss['Total_flux'], lotss['E_Total_flux'], lotss['Z_BEST'], np.repeat(si,len(lotss)), outfile='zmaxes' )
 
 ## calculate RLFs
-redshift_bins = np.arange( 0.0, zmax+0.001,zmax )
+#redshift_bins = np.arange( 0.0, zmax+0.001,zmax )
 
 redshift_bins = np.array([zmin,zmax])
-lum_bins = np.arange( 19.8, 27, 0.4 ) + np.log10( np.power( (144./1400.), si ) )
+lum_bins = np.arange( 19.5, 27, 0.5 ) + np.log10( np.power( (144./1400.), si ) )
 Lrad = radio_power( lotss['Total_flux'], lotss['Z_BEST'], spectral_index=si )
 RLF = RLF_from_zmax( Lrad, lotss['Z_BEST'], lum_bins, redshift_bins, lotss_zmax, area_deg2, area_units='deg2', error_type='rms' )
 

### 2024-02-02 src/scripts/mauchsadler.py
added cochrane

@@ -15,14 +15,18 @@ mauch_sadler = Table.read( paths.static / 'mauch_sadler_table5.csv', format='csv
 ## shift using spectral index
 ms_144MHz = mauch_sadler['log10_P1p4GHz'] + np.log10( np.power( (144./1400.), si ) ) 
 
+## lotss data -- star forming galaxies
+cochrane = Table.read( paths.static / 'cochrane_2023_table1.csv', format='csv', delimiter=',' )
+
 ## read in RLF
 RLF = Table.read( paths.data / 'RLF.fits' )
 
 fig = plt.figure( figsize=(5,5) )
 ## plot
-plt.plot( ms_144MHz, mauch_sadler['log10RLF_all'], color='black', linewidth=2.5, label='All' )
-plt.plot( ms_144MHz, mauch_sadler['log10RLF_SF'], color='magenta', linewidth=2.5, label='SFG' )
-plt.plot( ms_144MHz, mauch_sadler['log10RLF_RLAGN'], color='orange', linewidth=2.5, label='AGN' )
+plt.plot( ms_144MHz, mauch_sadler['log10RLF_all'], color='black', linewidth=2.5, label='MS07 All' )
+plt.plot( ms_144MHz, mauch_sadler['log10RLF_SF'], color='magenta', linewidth=2.5, label='MS07 SFG' )
+plt.plot( ms_144MHz, mauch_sadler['log10RLF_RLAGN'], color='orange', linewidth=2.5, label='MS07 AGN' )
+plt.plot( cochrane['logL150'], cochrane['logPhi'], color='blue', label='Cochrane et al. 2023, SFGs')
 ## plot the lofar data, filtering zeros
 non_zero = np.where( RLF['RLF'] != 0.0 )[0]
 RLF = RLF[non_zero]

### 2024-02-02 src/scripts/mauchsadler.py
added cochrane

@@ -15,8 +15,9 @@ mauch_sadler = Table.read( paths.static / 'mauch_sadler_table5.csv', format='csv
 ## shift using spectral index
 ms_144MHz = mauch_sadler['log10_P1p4GHz'] + np.log10( np.power( (144./1400.), si ) ) 
 
-## lotss data -- star forming galaxies
+## lotss data
 cochrane = Table.read( paths.static / 'cochrane_2023_table1.csv', format='csv', delimiter=',' )
+kondapally = Table.read( paths.static / 'kondapally_2022_table2.csv', format='csv', delimiter=',' )
 
 ## read in RLF
 RLF = Table.read( paths.data / 'RLF.fits' )
@@ -27,6 +28,7 @@ plt.plot( ms_144MHz, mauch_sadler['log10RLF_all'], color='black', linewidth=2.5,
 plt.plot( ms_144MHz, mauch_sadler['log10RLF_SF'], color='magenta', linewidth=2.5, label='MS07 SFG' )
 plt.plot( ms_144MHz, mauch_sadler['log10RLF_RLAGN'], color='orange', linewidth=2.5, label='MS07 AGN' )
 plt.plot( cochrane['logL150'], cochrane['logPhi'], color='blue', label='Cochrane et al. 2023, SFGs')
+plt.plot( kondapally['logL150'], kondapally['logPhi'], color='red', label='Kondapally et al. 2022, RLAGNs')
 ## plot the lofar data, filtering zeros
 non_zero = np.where( RLF['RLF'] != 0.0 )[0]
 RLF = RLF[non_zero]

### 2024-02-02 src/scripts/completeness.py
plot completeness

@@ -1,0 +1,36 @@
+import paths
+import numpy as np
+import matplotlib
+from matplotlib import pyplot as plt
+from astropy.table import Table, Column, join
+## WMAP9 is Hinshaw et al. 2013, H_0=69.3, Omega=0.287
+
+##############################################################
+## Plotting housekeeping
+
+matplotlib.rcParams['legend.frameon'] = False
+matplotlib.rcParams['axes.labelsize'] = 'large'
+## set up some colours
+n = 255
+mycols = plt.cm.viridis(np.linspace(0, 1,n))
+mycols_m = plt.cm.magma(np.linspace(0, 1,n))
+
+##############################################################
+
+
+## read in completeness
+cochrane = Table.read( paths.static / 'cochrane_2023_tableA1.csv', format='csv', delimiter=',' )
+kondapally = Table.read( paths.static / 'kondapally_2022_table1.csv', format='csv', delimiter=',' )
+
+fig = plt.figure( figsize=(5,5) )
+## plot simba
+plt.plot( cochrane['FluxDensity_mJy'], cochrane['Lockman'], color=mycols_m[20], linewidth=2.5, label='SFGs, Cochrane' )
+plt.plot( Simba_AGN['x'], Simba_AGN['Curve2'], color=mycols[20], linewidth=2.5, label='RLAGN, Kondapally' )
+plt.xlabel(r'$S_{\mathrm{144 MHz}}$'+' mJy')
+plt.ylabel('Completeness')
+plt.legend()
+plt.savefig(paths.figures / 'completeness.png',dpi=300)
+fig.clear()
+
+
+

### 2024-02-02 src/scripts/completeness.py
oops

@@ -25,7 +25,7 @@ kondapally = Table.read( paths.static / 'kondapally_2022_table1.csv', format='cs
 fig = plt.figure( figsize=(5,5) )
 ## plot simba
 plt.plot( cochrane['FluxDensity_mJy'], cochrane['Lockman'], color=mycols_m[20], linewidth=2.5, label='SFGs, Cochrane' )
-plt.plot( Simba_AGN['x'], Simba_AGN['Curve2'], color=mycols[20], linewidth=2.5, label='RLAGN, Kondapally' )
+plt.plot( kondapally['FluxDensity_mJy'], kondapally['Lockman'], color=mycols[20], linewidth=2.5, label='RLAGN, Kondapally' )
 plt.xlabel(r'$S_{\mathrm{144 MHz}}$'+' mJy')
 plt.ylabel('Completeness')
 plt.legend()

### 2024-02-02 src/scripts/completeness.py
ylim

@@ -24,10 +24,11 @@ kondapally = Table.read( paths.static / 'kondapally_2022_table1.csv', format='cs
 
 fig = plt.figure( figsize=(5,5) )
 ## plot simba
-plt.plot( cochrane['FluxDensity_mJy'], cochrane['Lockman'], color=mycols_m[20], linewidth=2.5, label='SFGs, Cochrane' )
+plt.plot( cochrane['FluxDensity_mJy'], cochrane['Lockman'], color=mycols_m[50], linewidth=2.5, label='SFGs, Cochrane' )
 plt.plot( kondapally['FluxDensity_mJy'], kondapally['Lockman'], color=mycols[20], linewidth=2.5, label='RLAGN, Kondapally' )
 plt.xlabel(r'$S_{\mathrm{144 MHz}}$'+' mJy')
 plt.ylabel('Completeness')
+plt.ylim((0.8,1.0))
 plt.legend()
 plt.savefig(paths.figures / 'completeness.png',dpi=300)
 fig.clear()

### 2024-04-03 src/scripts/completeness.py
add elais

@@ -22,16 +22,29 @@ mycols_m = plt.cm.magma(np.linspace(0, 1,n))
 cochrane = Table.read( paths.static / 'cochrane_2023_tableA1.csv', format='csv', delimiter=',' )
 kondapally = Table.read( paths.static / 'kondapally_2022_table1.csv', format='csv', delimiter=',' )
 
-fig = plt.figure( figsize=(5,5) )
-## plot simba
-plt.plot( cochrane['FluxDensity_mJy'], cochrane['Lockman'], color=mycols_m[250], linewidth=2.5, label='SFGs, Cochrane' )
-plt.plot( kondapally['FluxDensity_mJy'], kondapally['Lockman'], color=mycols[20], linewidth=2.5, label='RLAGN, Kondapally' )
-plt.xlabel(r'$S_{\mathrm{144 MHz}}$'+' mJy')
-plt.ylabel('Completeness')
-plt.ylim((0.8,1.0))
-plt.legend()
-plt.savefig(paths.figures / 'completeness.png',dpi=300)
+fig = plt.figure( figsize=(9,5) )
+gs = fig.add_gridspec(1,2,vspace=0)
+axs = gs.subplots(sharex=True,sharey=True)
+## plot Lockman
+axs[0].plot( cochrane['FluxDensity_mJy'], cochrane['Lockman'], color=mycols_m[250], linewidth=2.5, label='SFGs, Cochrane' )
+axs[0].plot( kondapally['FluxDensity_mJy'], kondapally['Lockman'], color=mycols[20], linewidth=2.5, label='RLAGN, Kondapally' )
+axs[0].set_xlabel(r'$S_{\mathrm{144 MHz}}$'+' mJy')
+axs[0].set_ylabel('Completeness')
+axs[0].set_ylim((0.8,1.0))
+axs[0].legend()
+axs[0].text(0.05,0.9,'Lockman',transform=axs[0].transAxes)
+## plot Elais
+axs[1].plot( cochrane['FluxDensity_mJy'], cochrane['Elais'], color=mycols_m[250], linewidth=2.5, label='SFGs, Cochrane' )
+axs[1].plot( kondapally['FluxDensity_mJy'], kondapally['Elais'], color=mycols[20], linewidth=2.5, label='RLAGN, Kondapally' )
+axs[1].set_xlabel(r'$S_{\mathrm{144 MHz}}$'+' mJy')
+#axs[1].set_ylabel('Completeness')
[…]

### 2024-04-03 src/scripts/mauchsadler.py

@@ -7,6 +7,18 @@ from helper_functions import *
 from astropy.io import fits
 ## WMAP9 is Hinshaw et al. 2013, H_0=69.3, Omega=0.287
 
+##############################################################
+## Plotting housekeeping
+
+matplotlib.rcParams['legend.frameon'] = False
+matplotlib.rcParams['axes.labelsize'] = 'large'
+## set up some colours
+n = 255
+mycols = plt.cm.viridis(np.linspace(0, 1,n))
+mycols_m = plt.cm.magma(np.linspace(0, 1,n))
+
+##############################################################
+
 ## set a default spectral index
 si = -0.8
 

### 2024-04-03 src/scripts/completeness.py
typo

@@ -23,7 +23,7 @@ cochrane = Table.read( paths.static / 'cochrane_2023_tableA1.csv', format='csv',
 kondapally = Table.read( paths.static / 'kondapally_2022_table1.csv', format='csv', delimiter=',' )
 
 fig = plt.figure( figsize=(9,5) )
-gs = fig.add_gridspec(1,2,vspace=0)
+gs = fig.add_gridspec(1,2,hspace=0)
 axs = gs.subplots(sharex=True,sharey=True)
 ## plot Lockman
 axs[0].plot( cochrane['FluxDensity_mJy'], cochrane['Lockman'], color=mycols_m[250], linewidth=2.5, label='SFGs, Cochrane' )

### 2024-04-03 src/scripts/completeness.py
hspace

@@ -40,7 +40,7 @@ axs[1].set_xlabel(r'$S_{\mathrm{144 MHz}}$'+' mJy')
 #axs[1].set_ylabel('Completeness')
 axs[1].set_ylim((0.8,1.0))
 axs[1].legend()
-axs[1].text(0.05,0.9,'ELAIS',transform=axs[0].transAxes)
+axs[1].text(0.05,0.9,'ELAIS',transform=axs[1].transAxes)
 fig.tight_layout()
 fig.savefig(paths.figures / 'completeness.png',dpi=300)
 fig.clear()

### 2024-04-03 src/scripts/completeness.py
hspace

@@ -23,7 +23,7 @@ cochrane = Table.read( paths.static / 'cochrane_2023_tableA1.csv', format='csv',
 kondapally = Table.read( paths.static / 'kondapally_2022_table1.csv', format='csv', delimiter=',' )
 
 fig = plt.figure( figsize=(9,5) )
-gs = fig.add_gridspec(1,2,hspace=0)
+gs = fig.add_gridspec(1,2,hspace=0,wspace=0)
 axs = gs.subplots(sharex=True,sharey=True)
 ## plot Lockman
 axs[0].plot( cochrane['FluxDensity_mJy'], cochrane['Lockman'], color=mycols_m[250], linewidth=2.5, label='SFGs, Cochrane' )

### 2024-04-10 src/scripts/generateVmaxes.py
using new catalogues for vmax

@@ -16,32 +16,24 @@ zmin = 0.003  ## matches Mauch & Sadler 2007
 zmax = 0.3   ## matches Mauch & Sadler 2007
 dz = 0.0001  ## matches Cochrane and Kondapally
 
+## completeness corrections
+cochrane = Table.read( paths.static / 'cochrane_2023_tableA1.csv', format='csv', delimiter=',' )
+kondapally = Table.read( paths.static / 'kondapally_2022_table1.csv', format='csv', delimiter=',' )
 
 ## Lockman
-infile = paths.static / 'lockman_final_cross_match_catalogue-v1.0_classifications.fits'
+infile = paths.static / 'lockman_final_cross_match_catalogue-v1.0_classifications_catalogue_filtered_full_SNR5_fluxscaled_withoffset_noduplicates_with_lotss_DR1_detectable.fits'
+rms_image = paths.static / '{:s}_rms_starmask_optical.fits'.format(field)
+
 field = os.path.basename(infile).split('_')[0]
-outfits = field + '_vmax.fits'
-if not os.path.exists(outfits):
-    lotss = Table.read( infile, format='fits' )
-    vmaxes = get_vmax( lotss, field, zmin=zmin, zmax=zmax, dz=dz, si=si )
-    vmaxes.write( paths.data / outfits, format='fits', overwrite=True )
-else:
-    vmaxes = Table.read( outfits )
+lotss = Table.read( infile, format='fits' )
 
-
-## calculate RLFs
-redshift_bins = np.array([zmin,zmax])
-lum_bins = np.arange( 19.5, 27, 0.5 ) + np.log10( np.power( (144./1400.), si ) )
-Lrad = radio_power( lotss['Total_flux'], lotss['Z_BEST'], spectral_index=si )
+## 6 arcsec vmaxes
+outfits = field + '_6arcsec_vmax.fits'
+vmaxes = get_vmax( lotss, field, col_suffix='_dr1' zmin=zmin, zmax=zmax, dz=dz, si=si, rms_image=rms_image, cochrane=cochrane, kondapally=kondapally )
+vmaxes.write( paths.data / outfits, format='fits', overwrite=True )
 
 
 
-RLF = RLF_from_zmax( Lrad, lotss['Z_BEST'], lum_bins, redshift_bins, lotss_zmax, area_deg2, area_units='deg2', error_type='rms' )
 
-## first  
[…]

### 2024-04-10 src/scripts/generateZmaxes.py

@@ -1,71 +1,0 @@
-import paths
-import numpy as np
-import matplotlib
-from matplotlib import pyplot as plt
-from astropy.table import Table, Column, join
-from helper_functions import *
-from astropy.io import fits
-## WMAP9 is Hinshaw et al. 2013, H_0=69.3, Omega=0.287
-
-#########################################
-## default parameters to adjust
-si = -0.8
-zmin = 0.003  ## matches Mauch & Sadler 2007
-zmax = 0.3   ## matches Mauch & Sadler 2007
-flux_cut = 150e-6
-
-## test test
-
-## read in lofar data
-lotss = Table.read( paths.static / 'lockman_final_cross_match_catalogue-v1.0_classifications.fits', format='fits' )
-## remove things with no redshifts, i.e. those which are masked
-good_z = np.where( np.logical_not(lotss['Z_BEST'].mask) )[0]
-lotss = lotss[good_z]
-## remove the mask to avoid warnings later on
-lotss['Z_BEST'] = np.ma.getdata(lotss['Z_BEST'])
-
-test = True
-if test:
-    good_z = np.where(lotss['Z_BEST'] < zmax )[0]
-    lotss = lotss[good_z]
-
-## Total area
-rms_image = fits.open( paths.static / 'lockman_rms_starmask_optical.fits' )
-rms = rms_image[0].data
-rms_idx = np.where(rms <= flux_cut)
-npix_total = len(rms_idx[0])
-pixscale = np.abs(rms_image[0].header['CDELT1'])
-pixarea = np.power(pixscale, 2.)
-area_deg2 = npix_total * pixarea
-
-## get z-max
-lotss_zmax = RLF_calculate_zmax( flux_cut, lotss['Source_Name'], lotss['Total_flux'], lotss['E_Total_flux'], lotss['Z_BEST'], np.repeat(si,len(lotss)), outfile='zmaxes' )
-
-## completeness corrections
-cochrane = Table.read( paths.static / 'cochrane_2023_tableA1.csv', format='csv', delimiter=',' )
-kondapally = Table.read( paths.static / 'kondapally_2022_table1.csv', format='csv', delimiter=',' )
-completeness = []
-for i in np.arange(0,len(lotss)):
-    if lotss['Overall_class'][i] == 'SFG':
[…]

### 2024-04-10 src/scripts/helper_functions.py

@@ -56,7 +56,6 @@ def get_source_vmax( power, power_to_flux_factor, V_DLs, completeness, fluxdens,
     if len(noise_diff) < len(flux_at_z):
         flux_at_z = power_to_flux_factor[noise_diff] * power
         V_DLs = V_DLs[noise_diff]
-    completeness = completeness.filled(0)
     solid_angle_at_z = []
     completeness_at_z = []
     for j in np.arange(0,len(flux_at_z)):
@@ -72,7 +71,10 @@ def get_source_vmax( power, power_to_flux_factor, V_DLs, completeness, fluxdens,
     vmax = np.nansum( V_DLs * theta_sz )
     return(vmax)
 
-def get_vmax( lotss, field, zmin=0.003, zmax=0.3, dz=0.0001, si=-0.8 ):
+def get_vmax( lotss, field, col_suffix='', zmin=0.003, zmax=0.3, dz=0.0001, si=-0.8, rms_image='', cochrane='', kondapally='' ):
+    fluxcol = 'Total_flux{:s}'.format(col_suffix)
+    fluxerrcol = 'E_Total_flux{:s}'.format(col_suffix)
+
     ## remove things with no redshifts, i.e. those which are masked
     good_z = np.where( np.logical_not(lotss['Z_BEST'].mask) )[0]
     lotss = lotss[good_z]
@@ -85,8 +87,6 @@ def get_vmax( lotss, field, zmin=0.003, zmax=0.3, dz=0.0001, si=-0.8 ):
         lotss = lotss[good_z]
 
     ## get solid angle
-    rms_image = paths.static / '{:s}_rms_starmask_optical.fits'.format(field)
-    print(rms_image)
     noise_bins, eff_area_deg2 = effective_area( rms_image )
     eff_area_sr = eff_area_deg2 * np.power( ( np.pi / 180. ), 2. )
     solid_angle = eff_area_sr / ( 4. * np.pi )
@@ -116,10 +116,6 @@ def get_vmax( lotss, field, zmin=0.003, zmax=0.3, dz=0.0001, si=-0.8 ):
     ## sigma cutoff
     sigma_cut = 5.
 
-    ## completeness corrections
-    cochrane = Table.read( paths.static / 'cochrane_2023_tableA1.csv', format='csv', delimiter=',' )
-    kondapally = Table.read( paths.static / 'kondapally_2022_table1.csv', format='csv', delimiter=',' )
-
[…]

### 2024-04-10 src/scripts/mauchsadler.py

@@ -31,8 +31,29 @@ ms_144MHz = mauch_sadler['log10_P1p4GHz'] + np.log10( np.power( (144./1400.), si
 cochrane = Table.read( paths.static / 'cochrane_2023_table1.csv', format='csv', delimiter=',' )
 kondapally = Table.read( paths.static / 'kondapally_2022_table2.csv', format='csv', delimiter=',' )
 
-## read in RLF
-RLF = Table.read( paths.data / 'RLF.fits' )
+## read in vmaxes
+vmaxes = Table.read( paths.data / 'lockman_vmaxes.fits', format='fits' )
+
+
+## calculate RLFs
+redshift_bins = np.array([zmin,zmax])
+lum_bins = np.arange( 19.5, 27, 0.5 ) + np.log10( np.power( (144./1400.), si ) )
+Lrad = radio_power( vmaxes['Total_flux'], vmaxes['Z_BEST'], spectral_index=si )
+log10_Lrad = np.log10(Lrad)
+
+phis = []
+for i in np.arange(1,len(lum_bins)):
+    delta_log_L = lum_bins[i] - lum_bins[i-1]
+    lum_idx = np.logical_and( log10_Lrad >= lum_bins[i-1], log10_Lrad < lum_bins[i] )
+    phi = ( 1. / np.power( 10., delta_log_L ) ) * np.sum( 1. / vmaxes['vmax'][lum_idx] )
+    phis.append(phi)
+
+## convert back to "mag" 
+phis = np.asarray(phis)/1e7
+
+lum_bin_cens = lum_bins + 0.5*(lum_bins[1]-lum_bins[0])
+
+
 
 fig = plt.figure( figsize=(5,5) )
 ## plot
@@ -42,11 +63,11 @@ fig = plt.figure( figsize=(5,5) )
 plt.plot( cochrane['logL150'], cochrane['logPhi'], color='blue', label='Cochrane et al. 2023, SFGs')
 plt.plot( kondapally['logL150'], kondapally['logPhi'], color='red', label='Kondapally et al. 2022, RLAGNs')
 ## plot the lofar data, filtering zeros
-non_zero = np.where( RLF['RLF'] != 0.0 )[0]
-RLF = RLF[non_zero]
-plt.fill_between( RLF['Lmedian'], RLF['RLF_lo'], RLF['RLF_up'], color='green', alpha=0.5 )
[…]

### 2024-04-10 src/scripts/mauchsadler.py
input for RLF plotting

@@ -32,7 +32,7 @@ cochrane = Table.read( paths.static / 'cochrane_2023_table1.csv', format='csv', 
 kondapally = Table.read( paths.static / 'kondapally_2022_table2.csv', format='csv', delimiter=',' )
 
 ## read in vmaxes
-vmaxes = Table.read( paths.data / 'lockman_vmaxes.fits', format='fits' )
+vmaxes = Table.read( paths.data / 'lockman_6arcsec_vmaxes.fits', format='fits' )
 
 
 ## calculate RLFs

### 2024-04-10 src/scripts/generateVmaxes.py
missing comma

@@ -29,7 +29,7 @@ lotss = Table.read( infile, format='fits' )
 
 ## 6 arcsec vmaxes
 outfits = field + '_6arcsec_vmax.fits'
-vmaxes = get_vmax( lotss, field, col_suffix='_dr1' zmin=zmin, zmax=zmax, dz=dz, si=si, rms_image=rms_image, cochrane=cochrane, kondapally=kondapally )
+vmaxes = get_vmax( lotss, field, col_suffix='_dr1', zmin=zmin, zmax=zmax, dz=dz, si=si, rms_image=rms_image, cochrane=cochrane, kondapally=kondapally )
 vmaxes.write( paths.data / outfits, format='fits', overwrite=True )
 
 

### 2024-04-10 src/scripts/generateVmaxes.py
field before called

@@ -22,9 +22,8 @@ kondapally = Table.read( paths.static / 'kondapally_2022_table1.csv', format='cs
 
 ## Lockman
 infile = paths.static / 'lockman_final_cross_match_catalogue-v1.0_classifications_catalogue_filtered_full_SNR5_fluxscaled_withoffset_noduplicates_with_lotss_DR1_detectable.fits'
-rms_image = paths.static / '{:s}_rms_starmask_optical.fits'.format(field)
-
 field = os.path.basename(infile).split('_')[0]
+rms_image = paths.static / '{:s}_rms_starmask_optical.fits'.format(field)
 lotss = Table.read( infile, format='fits' )
 
 ## 6 arcsec vmaxes

### 2024-05-17 src/scripts/comparison_plot.py
updates

@@ -31,8 +31,8 @@ cochrane = Table.read( paths.static / 'cochrane_2023_table1.csv', format='csv', 
 kondapally = Table.read( paths.static / 'kondapally_2022_table2.csv', format='csv', delimiter=',' )
 
 ## read in vmaxes
-lockman_vmaxes = Table.read( paths.data / 'lockman_6arcsec_vmaxes.fits', format='fits' )
-elais_vmaxes = Table.read( paths.data / 'en1_6arcsec_vmaxes.fits', format='fits' )
+lockman_vmaxes = Table.read( paths.data / 'lockman_vmaxes.fits', format='fits' )
+elais_vmaxes = Table.read( paths.data / 'en1_vmaxes.fits', format='fits' )
 
 ## combine these into a single catalogue for plotting
 keep_cols = ['Total_flux_dr','Z_BEST','vmax','agn_vmax','sf_vmax','AGN_flux','SF_flux', 'Overall_class','Mass_cons','SFR_cons']

### 2024-05-17 src/scripts/generateVmaxes.py

@@ -39,7 +39,7 @@ for field in fields:
     lotss = get_tb_information( lotss, im_weight=0.5, maj_lim=0.4, min_lim=0.3, T_e=T_e, alpha=si, ref_freqs=ref_freqs, freqs_GHz=freqs_GHz, use_z=False )
     lotss = do_SFR_AGN_separation( lotss )
     ## 6 arcsec vmaxes
-    outfits = field + '_6arcsec_vmaxes.fits'
+    outfits = field + '_vmaxes.fits'
     vmaxes = get_vmax( lotss, field, col_suffix='_dr', zmin=zmin, zmax=zmax, dz=dz, si=si, sigma_cut=sigma_cut, rms_image=rms_image, cochrane=cochrane, kondapally=kondapally, test=False )
     vmaxes.write( paths.data / outfits, format='fits', overwrite=True )
 

### 2024-05-28 src/scripts/comparison_plot.py
aesthetics

@@ -15,7 +15,7 @@ matplotlib.rcParams['axes.labelsize'] = 'large'
 ## set up some colours
 n = 255
 mycols = plt.cm.viridis(np.linspace(0, 1,n))
-mycols_m = plt.cm.magma(np.linspace(0, 1,n))
+mycols_m = plt.cm.inferno(np.linspace(0, 1,n))
 
 ##############################################################
 
@@ -124,72 +124,101 @@ gal_sf_lum_func = np.asarray(gal_sfg_rhos)
 lum_bin_cens = lum_bins[0:-1] + 0.5*(lum_bins[1]-lum_bins[0])
 
 
-fsizex = 15
+fsizex = 14
 fsizey = 5
 sbsizex = 0.8
 sbsizey = 0.8
-plxlims = (19,27)
-plylims = (-7.5,-1)
+plxlims = (20.1,27)
+plylims = (-6.5,-1)
+
+## set some colors
+kond = 'black'
+coch = 'gray'
+sfggalc = mycols[80]
+agngalc = mycols_m[120]
+sfc = mycols[180]
+agnc = mycols_m[180]
 
 fig = plt.figure( figsize=(fsizex,fsizey) )
-## start first panel -- this is Cochrane and Kondapally vs. the 6 arcsec but in the high resolution area
-p1 = plt.axes([0.1,0.1,sbsizex*fsizey/fsizex,sbsizey])
+## Left panel: Showing that we reproduce the RLFs for galaxies even when using smaller area
+p1 = plt.axes([0.05,0.1,sbsizex*fsizey/fsizex,sbsizey])
 ## plot the previous data
-p1.plot( cochrane['logL150'], cochrane['logPhi'], color='blue', label='Cochrane et al. 2023, SFGs')
-p1.plot( kondapally['logL150'], kondapally['logPhi'], color='red', label='Kondapally et al. 2022, RLAGNs')
-## plot the lofar data, filtering zeros
-#non_zero = np.where( lum_func != 0.0 )[0]
-#p1.plot( lum_bin_cens[non_zero], lum_func[non_zero], color='black', label='Total' )
-#non_zero = np.where( agn_lum_func != 0.0 )[0]
-#p1.plot( lum_bin_cens[non_zero], agn_lum_func[non_zero], color='purple', label='AGN activity' )
[…]

### 2024-05-28 src/scripts/comparison_plot.py
rename plot

@@ -214,7 +214,7 @@ p4.axes.set_xlim(plxlims)
 p4.set_xlabel('log'+r'$_{10}$'+'('+r'$L_{\mathrm{144 MHz}}$'+' W Hz'+r'$^{-1}$'+'])')
 p4.set_ylabel('Activity / Galaxy')
 p4.legend()
-fig.savefig(paths.figures / 'mauch_sadler_RLFs.png',dpi=300)
+fig.savefig(paths.figures / 'deep_fields_RLFs.png',dpi=300)
 fig.clear()
 plt.close()
 

### 2024-05-28 src/scripts/comparison_plot.py
redshift bins

@@ -218,6 +218,67 @@ fig.savefig(paths.figures / 'deep_fields_RLFs.png',dpi=300)
 fig.clear()
 plt.close()
 
+'''
+Redshift bins
+
+Rohit:
+0.5 - 1.0
+1.0 - 1.5
+1.5 - 2.0
+2.0 - 2.5
+
+
+Rachael:
+0.1 - 0.4 
+0.4 - 0.6 
+0.6 - 0.8
+0.8 - 1.0 
+1.0 - 1.3
+1.3 - 1.6
+1.6 - 2.0 
+2.0 - 2.5
+2.5 - 3.3
+3.3 - 4.6
+4.6 - 5.7
+
+'''
+
+
+## read in SIMBA
+Simba_SF = Table.read( paths.static / 'RLFS_50MYR_SF.csv', format='csv', delimiter=',' )
+Simba_AGN = Table.read( paths.static / 'RLFS_50MYR_AGN.csv', format='csv', delimiter=',' )
+## shift using spectral index
+Simba_SF['x']  = Simba_SF['x'] + np.log10( np.power( (144./1400.), si ) )
+Simba_AGN['x']  = Simba_AGN['x'] + np.log10( np.power( (144./1400.), si ) )
+
+fig = plt.figure( figsize=(5,5) )
+## plot simba
+#plt.plot( Simba_SF['x'], Simba_SF['Curve1']+np.log10(np.log(10)), color='green', linewidth=2.5, label='Simba SF' )
+#plt.plot( Simba_AGN['x'], Simba_AGN['Curve2']+np.log10(np.log(10)), color='blue', linewidth=2.5, label='Simba AGN' )
+
+plt.plot( Simba_SF['x'], Simba_SF['Curve1'], color='green', linewidth=2.5, label='Simba SF' )
+plt.plot( Simba_AGN['x'], Simba_AGN['Curve2'], color='blue', linewidth=2.5, label='Simba AGN' )
+
+
+non_zero = np.where( gal_agn_lum_func != 0.0 )[0]
+plt.plot( lum_bin_cens[non_zero], gal_agn_lum_func[non_zero], color=agngalc, label='AGN galaxies', linewidth=3, linestyle='dotted' )
+non_zero = np.where( gal_sf_lum_func != 0.0 )[0]
[…]

### 2024-05-28 src/scripts/generateVmaxes.py

@@ -14,8 +14,8 @@ import time
 #########################################
 ## default parameters to adjust
 si = -0.7
-zmin = 0.003  ## matches Mauch & Sadler 2007
-zmax = 0.3   ## matches Mauch & Sadler 2007
+#zmin = 0.003  ## matches Mauch & Sadler 2007
+#zmax = 0.3   ## matches Mauch & Sadler 2007
 dz = 0.0001  ## matches Cochrane and Kondapally
 ## for T_b
 T_e = 1e4
@@ -30,6 +30,9 @@ kondapally = Table.read( paths.static / 'kondapally_2022_table1.csv', format='cs
 
 fields = ['lockman', 'en1']
 
+zbin_starts = [ 0.003, 0.5, 1.0, 1.5, 2.0, 0.1, 0.4, 0.6, 0.8, 1.0, 1.3, 1.6, 2.5, 3.3, 4.6 ]
+zbin_ends = [ 0.3, 1.0, 1.5, 2.0, 2.5, 0.4, 0.6, 0.8, 1.0, 1.3, 1.6, 2.0, 3.3, 4.6, 5.7 ]
+
 for field in fields:
     print('Starting with field: {:s}'.format(field))
     infile = paths.static / '{:s}_03_matched_inMOC_inHR.fits'.format(field)
@@ -39,18 +42,15 @@ for field in fields:
     ## add brightness temperature information
     lotss = get_tb_information( lotss, im_weight=0.5, maj_lim=0.4, min_lim=0.3, T_e=T_e, alpha=si, ref_freqs=ref_freqs, freqs_GHz=freqs_GHz, use_z=False )
     lotss = do_SFR_AGN_separation( lotss )
-    ## 6 arcsec vmaxes
-    outfits = field + '_vmaxes.fits'
-    vmaxes = get_vmax( lotss, field, col_suffix='_dr', zmin=zmin, zmax=zmax, dz=dz, si=si, sigma_cut=sigma_cut, rms_image=rms_image, cochrane=cochrane, kondapally=kondapally, test=False )
-    vmaxes.write( paths.data / outfits, format='fits', overwrite=True )
+    ## calculate vmaxes
+    for i in np.arange(0,len(zbin_ztarts)):
+        zmin = zbin_starts[i]
+        zmax = zbin_ends[i]
+        outfits = field + '_vmaxes_zmin'+str(zmin)+'_zmax'+str(zmax)+'.fits'
[…]

### 2024-05-28 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Title
-\title[Running head]{An open source scientific article}
+\title[TBD]{TBD}
 
@@
 $^{1}$Centre for Extragalactic Astronomy, Department of Physics, Durham University, Durham DH1 3LE, UK \\
-$^{2}$Institute for Computational Cosmology, Department of Physics, University of Durham, South Road, Durham DH1 3LE, UK \\ }
+$^{2}$Institute for Computational Cosmology, Department of Physics, Durham University, South Road, Durham DH1 3LE, UK \\ }
 
@@
 \begin{abstract}
-    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-    Phasellus eu tellus sit amet tortor gravida placerat.
-    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-    Praesent eget sem vel leo ultrices bibendum.
-    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-    Curabitur auctor semper nulla donec varius orci eget risus.
-    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-    Duis eget orci sit amet orci dignissim rutrum.
+    TBD
 \end{abstract}
@@
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-Phasellus eu tellus sit amet tortor gravida placerat.
-Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-Praesent eget sem vel leo ultrices bibendum.
-Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-Curabitur auctor semper nulla donec varius orci eget risus.
-Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-Duis eget orci sit amet orci dignissim rutrum.
+Stuff and things \citep{morabito_sub-arcsecond_2022}.
 
-Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
-Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
-Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
-Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
-Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
-Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
-Pellentesque cursus luctus mauris \citep{Luger2021}.
+\section{Data and Methods}
+
+\section{Results}
 
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{figures/test.png}
+    \includegraphics[width=0.5\textwidth]{figures/completeness.png}
     \caption{Caption}
-    \label{fig:enter-label}
-    \script{testplot.py}
+    \label{fig:completeness}
+    \script{completeness.py}
 \end{figure}
 
-\bibliography{bib}
+\begin{figure*}
+    \centering
+    \includegraphics[width=0.98\textwidth]{figures/deep_fields_RLFs.png}
+    \caption{Caption}
+    \label{fig:msrlfs}
+    \script{comparison_plot.py}
+\end{figure*}
+
+\section{Discussion and conclusions}
+
+\section*{Data Availability}
+Reference showyourwork on github
+
+\section*{Acknowledgements}
+LKM is grateful for support from UKRI [MR/T042842/1]. \hl{also cosma and spider?}
+
+\bibliographystyle{mnras}
+\bibliography{references}
+
+\label{lastpage}
 

### 2024-05-28 src/scripts/generateVmaxes.py
typo

@@ -43,7 +43,7 @@ for field in fields:
     lotss = get_tb_information( lotss, im_weight=0.5, maj_lim=0.4, min_lim=0.3, T_e=T_e, alpha=si, ref_freqs=ref_freqs, freqs_GHz=freqs_GHz, use_z=False )
     lotss = do_SFR_AGN_separation( lotss )
     ## calculate vmaxes
-    for i in np.arange(0,len(zbin_ztarts)):
+    for i in np.arange(0,len(zbin_starts)):
         zmin = zbin_starts[i]
         zmax = zbin_ends[i]
         outfits = field + '_vmaxes_zmin'+str(zmin)+'_zmax'+str(zmax)+'.fits'

### 2024-06-06 src/scripts/RLF_evolution.py
move data to static

@@ -45,7 +45,7 @@ for i in np.arange(0,5):
     vmaxes = Table()
     for field in fields:
         infile = '{:s}_vmaxes_zmin{:s}_zmax{:s}.fits'.format(field,str(zmin),str(zmax))
-        tmp = Table.read( paths.data / infile, format='fits' )
+        tmp = Table.read( paths.static / infile, format='fits' )
         vmaxes = vstack([vmaxes,tmp[keep_cols]])
     lum_bins, lum_func, agn_lum_func, sf_lum_func, gal_agn_lum_func, gal_sf_lum_func = get_RLFs( vmaxes, zmin, zmax, lmin=20.5, lmax=27, dl=0.3, si=si )
     z_lum_bins.append(lum_bins)

### 2024-06-06 src/scripts/comparison_plot.py

@@ -31,8 +31,8 @@ cochrane = Table.read( paths.static / 'cochrane_2023_table1.csv', format='csv', 
 kondapally = Table.read( paths.static / 'kondapally_2022_table2.csv', format='csv', delimiter=',' )
 
 ## read in vmaxes
-lockman_vmaxes = Table.read( paths.data / 'lockman_vmaxes_zmin0.003_zmax0.3.fits', format='fits' )
-elais_vmaxes = Table.read( paths.data / 'en1_vmaxes_zmin0.003_zmax0.3.fits', format='fits' )
+lockman_vmaxes = Table.read( paths.static / 'lockman_vmaxes_zmin0.003_zmax0.3.fits', format='fits' )
+elais_vmaxes = Table.read( paths.static / 'en1_vmaxes_zmin0.003_zmax0.3.fits', format='fits' )
 
 ## combine these into a single catalogue for plotting
 keep_cols = ['Total_flux_dr','Z_BEST','vmax','agn_vmax','sf_vmax','AGN_flux','SF_flux', 'Overall_class','Mass_cons','SFR_cons']

### 2024-06-06 src/scripts/generateVmaxes.py

@@ -48,11 +48,11 @@ for field in fields:
         zmin = zbin_starts[i]
         zmax = zbin_ends[i]
         outfits = field + '_vmaxes_zmin'+str(zmin)+'_zmax'+str(zmax)+'.fits'
-        if not os.path.exists( paths.data / outfits ):
+        if not os.path.exists( paths.static / outfits ):
             vmaxes = get_vmax( lotss, field, col_suffix='_dr', zmin=zmin, zmax=zmax, dz=dz, si=si, sigma_cut=sigma_cut, rms_image=rms_image, cochrane=cochrane, kondapally=kondapally, test=False )
-            vmaxes.write( paths.data / outfits, format='fits', overwrite=True )
+            vmaxes.write( paths.static / outfits, format='fits', overwrite=True )
         else:
-            print('File {:s} already exists! If you want to regenerate, please delete output file and build again.'.format( str(paths.data / outfits) ) )
+            print('File {:s} already exists! If you want to regenerate, please delete output file and build again.'.format( str(paths.static / outfits) ) )
 
 
 ## uncertainties done by bootstrapping. From Kondapally et al.:

### 2024-06-06 src/scripts/helper_functions.py

@@ -295,7 +295,7 @@ def get_RLFs( vmaxes, zmin, zmax, lmin=20.5, lmax=27, dl=0.3, si=-0.7 ):
 
 def effective_area( rms_file ):
     tmp = os.path.basename( rms_file )
-    outfile = paths.data / tmp.replace( '.fits', '_effective_areas.fits' ) 
+    outfile = paths.static / tmp.replace( '.fits', '_effective_areas.fits' ) 
     print(outfile)
     if os.path.exists( outfile ):
         t = Table.read( outfile, format='fits' )

### 2024-06-06 src/scripts/RLF_evolution.py
add in en1 data

@@ -30,7 +30,7 @@ t = Table.read( paths.static / 'redshift_bins.csv', format='csv' )
 zbin_starts = t['zbin_starts']
 zbin_ends = t['zbin_ends']
 
-fields = ['lockman'] # ,'en1']
+fields = ['lockman','en1']
 
 z_lum_bins = []
 z_lum_func = []

### 2024-06-06 src/scripts/RLF_evolution.py
filter vmax=0

@@ -46,7 +46,10 @@ for i in np.arange(0,5):
     for field in fields:
         infile = '{:s}_vmaxes_zmin{:s}_zmax{:s}.fits'.format(field,str(zmin),str(zmax))
         tmp = Table.read( paths.static / infile, format='fits' )
-        vmaxes = vstack([vmaxes,tmp[keep_cols]])
+        vmaxes = vstack([vmaxes,tmp[keep_cols]],metadata_conflicts='silent')
+    ## here filter out anything with vmax = 0
+    ## these come from things which are always below the noise bins for the redshift range 
+    vmaxes = vmaxes[np.where(vmaxes['vmax']>0)]
     lum_bins, lum_func, agn_lum_func, sf_lum_func, gal_agn_lum_func, gal_sf_lum_func = get_RLFs( vmaxes, zmin, zmax, lmin=20.5, lmax=27, dl=0.3, si=si )
     z_lum_bins.append(lum_bins)
     z_lum_func.append(lum_func)

### 2024-06-07 src/scripts/comparison_plot.py
split plots

@@ -139,106 +139,3 @@ fig.savefig(paths.figures / 'deep_fields_RLFs.png',dpi=300)
 fig.clear()
 plt.close()
 
-'''
-Redshift bins
-
-Rohit:
-0.5 - 1.0
-1.0 - 1.5
-1.5 - 2.0
-2.0 - 2.5
-
-
-Rachael:
-0.1 - 0.4 
-0.4 - 0.6 
-0.6 - 0.8
-0.8 - 1.0 
-1.0 - 1.3
-1.3 - 1.6
-1.6 - 2.0 
-2.0 - 2.5
-2.5 - 3.3
-3.3 - 4.6
-4.6 - 5.7
-
-'''
-
-
-## read in SIMBA
-Simba_SF = Table.read( paths.static / 'RLFS_50MYR_SF.csv', format='csv', delimiter=',' )
-Simba_AGN = Table.read( paths.static / 'RLFS_50MYR_AGN.csv', format='csv', delimiter=',' )
-## shift using spectral index
-Simba_SF['x']  = Simba_SF['x'] + np.log10( np.power( (144./1400.), si ) )
-Simba_AGN['x']  = Simba_AGN['x'] + np.log10( np.power( (144./1400.), si ) )
-
-fig = plt.figure( figsize=(5,5) )
-## plot simba
-# mag = -2.5 * Lum + zeropoint
-# so Lum = mag / -2.5
-plt.plot( Simba_SF['x'], (Simba_SF['Curve1']+np.log10(2.5)), color='green', linewidth=2.5, label='Simba SF' )
-plt.plot( Simba_AGN['x'], (Simba_AGN['Curve2']+np.log10(2.5)), color='blue', linewidth=2.5, label='Simba AGN' )
-
-#plt.plot( Simba_SF['x'], Simba_SF['Curve1'], color='green', linewidth=2.5, label='Simba SF' )
-#plt.plot( Simba_AGN['x'], Simba_AGN['Curve2'], color='blue', linewidth=2.5, label='Simba AGN' )
-
-
-non_zero = np.where( gal_agn_lum_func != 0.0 )[0]
-plt.plot( lum_bin_cens[non_zero], gal_agn_lum_func[non_zero], color=agngalc, label='AGN galaxies', linewidth=3, linestyle='dotted' )
-non_zero = np.where( gal_sf_lum_func != 0.0 )[0]
[…]

### 2024-06-07 src/scripts/simba_comparison.py

@@ -1,0 +1,84 @@
+import paths
+import numpy as np
+import matplotlib
+from matplotlib import pyplot as plt
+from astropy.table import Table, Column, join, vstack
+from helper_functions import *
+from astropy.io import fits
+## WMAP9 is Hinshaw et al. 2013, H_0=69.3, Omega=0.287
+
+##############################################################
+## Plotting housekeeping
+
+matplotlib.rcParams['legend.frameon'] = False
+matplotlib.rcParams['axes.labelsize'] = 'large'
+## set up some colours
+n = 255
+mycols = plt.cm.viridis(np.linspace(0, 1,n))
+mycols_m = plt.cm.inferno(np.linspace(0, 1,n))
+
+##############################################################
+
+## set a default spectral index
+si = -0.7
+zmin = 0.003  ## matches Mauch & Sadler 2007
+zmax = 0.3   ## matches Mauch & Sadler 2007
+dz = 0.0001  ## matches Cochrane and Kondapally
+
+## read in SIMBA
+Simba_SF = Table.read( paths.static / 'RLFS_50MYR_SF.csv', format='csv', delimiter=',' )
+Simba_AGN = Table.read( paths.static / 'RLFS_50MYR_AGN.csv', format='csv', delimiter=',' )
+## shift using spectral index
+Simba_SF['x']  = Simba_SF['x'] + np.log10( np.power( (144./1400.), si ) )
+Simba_AGN['x']  = Simba_AGN['x'] + np.log10( np.power( (144./1400.), si ) )
+
+
+## read in vmaxes
+lockman_vmaxes = Table.read( paths.static / 'lockman_vmaxes_zmin0.003_zmax0.3.fits', format='fits' )
+elais_vmaxes = Table.read( paths.static / 'en1_vmaxes_zmin0.003_zmax0.3.fits', format='fits' )
+
+## combine these into a single catalogue for plotting
+keep_cols = ['Total_flux_dr','Z_BEST','vmax','agn_vmax','sf_vmax','AGN_flux','SF_flux', 'Overall_class','Mass_cons','SFR_cons']
+lockman_tmp = lockman_vmaxes[keep_cols]
[…]

### 2024-06-07 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \label{sec:intro}
+The extent to which active galactic nuclei (AGN) impact galaxy formation is a major open question in astrophysics.
+It is clear that there is an interaction, from both observational and theoretical standpoints.
+Observations have revealed tight scaling relations between the mass of a super-massive black hole and its host galaxy \citep[see, e.g.][and references therein]{kormendy_coevolution_2013}, while cosmological simulations require some form of AGN feedback \citep{bower_breaking_2006,croton_many_2006} to be able to reproduce the observed galaxy population.
+Galaxies build up their stellar mass through star formation, while super-massive black holes are expected to provide either negative or positive feedback \hl{citations!} to either suppress or stimulate galaxy growth.
+
+To understand the interplay between AGN activity and star formation (SF), it is necessary to measure both processes.
 
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{figures/completeness.png}
+    \includegraphics[width=0.5\textwidth]{figures/simba_comparison.png}
     \caption{Caption}
-    \label{fig:completeness}
-    \script{completeness.py}
+    \label{fig:simba}
+    \script{simba_comparison.py}
 \end{figure}
@@
     \script{comparison_plot.py}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=0.98\textwidth]{figures/RLF_evolution_AGN.png}
+    \caption{Caption}
+    \label{fig:agn_evolution}
+    \script{RLF_evolution.py}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=0.98\textwidth]{figures/RLF_evolution_SF.png}
+    \caption{Caption}
+    \label{fig:sf_evolution}
+    \script{RLF_evolution.py}
 \end{figure*}

### 2024-07-16 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Define document class
-\documentclass[usenatbib,fleqn]{mnras}
+\documentclass[usenatbib,fleqn,letters]{mnras}
 \usepackage{showyourwork}
@@
 \newcommand{\tbtotal}{$T_{b,\mathrm{total}}$}
+\newcommand{\muJybm}{$\mu$Jy$\,$beam$^{-1}$}
 
@@
     TBD
+    \vspace{6.75in}
 \end{abstract}
@@
 
-To understand the interplay between AGN activity and star formation (SF), it is necessary to measure both processes.
+However, if one wishes to study the cosmic history of AGN activity or star formation (SF), and by extension the interplay between them, it is necessary to measure each process separately.
+This is a difficult problem at any wavelength.
+For example \hl{mid-infrared}.
+Multi-wavelength approaches use spectral energy distribution (SED) fitting, which utilises information from a broad number of observed bands to jointly fit galaxy and AGN components \citep{pacifici_art_2022}.
+\hl{give a couple of examples} \citep{calistro_rivera_agnfitter_2016,boquien_cigale_2019}.
+However, SED fitting is costly as it relies on observations using many different instruments over the same part of the sky, which then have to be carefully cross-matched.
+Selection effects at different bands can also present a problem.
 
-Stuff and things \citep{morabito_sub-arcsecond_2022}.
+%% suggestion from Ryan: MIR community - AGN fraction
 
-\section{Data and Methods}
+An alternative method to unambiguously identify the AGN component(s) in a galaxy is through brightness temperature, $T_b$, measurements in the radio waveband.
+Radio emission has the advantage of being unobscured by dust or gas, meaning there are minimal selection effects.
+Brightness temperature is a non-physical temperature which is defined as the temperature of a blackbody which would produce the observed surface brightness.
+Models of radio emission from SF predict an upper limit to the surface brightness that can be produced, even in starburst galaxies \citep{condon_radio_1992}.
+Surface brightness which is detected above this limit therefore has to be attributed to AGN activity.
+Brightness temperature is inversely proportional to frequency and resolution, $T_b \propto \nu^{-2}\Omega^{-1}$ (where $\Omega \propto \theta_{maj}\theta_{min}$).
+Traditionally brightness temperature measurements require Very Long Baseline Interferometry (VLBI) which achieve milli-arcsec resolutions at GHz frequencies \hl{citations} to achieve enough sensitivity to make meaningful measurements.
+In \cite[][heareafter, \hl{M22}]{morabito_identifying_2022} we demonstrated using sub-arcsecond resolution (0.\sarc\ 3) at 144 MHz with the LOw Frequency ARray \citep[LOFAR;][]{van_haarlem_lofar:_2013}.
+The advantage of LOFAR is its wide field of view: $\sim$6 deg$^2$ from a single observation versus a few to a couple hundred arcmin$^2$ with VLBI at GHz frequencies, depending on whether one is imaging at the phase centre of the observation or using multi-phase centres for correlation \hl{citations}.
+For example, using an 8 hour LOFAR observation of the Lockman Hole \citep{sweijen_deep_2022} we identified 940 AGN via brightness temperature \cite{morabito_identifying_2022}, providing a sample two orders of magnitude larger than any VLBI sample of AGN identified at GHz frequencies.
+
+As part of the LOFAR Two-metre Sky Survey Deep Fields \citep{sabater_lofar_2021,tasse_lofar_2021} campaign, we have now doubled the areal coverage of 0.\sarc\ 3 resolution wide-field imaging.
+\cite{de_jong_into_2024} recently published their 0.\sarc\ 3 resolution image of the ELAIS-N1 field, which reaches a depth of 14 $\mu$Jy$\,$bm$^{-1}$ using 32 hours of observations.
+They also demonstrate the flexible resolution which can be achieved with LOFAR, by imaging at 0.\sarc\ 6 and 1.\sarc\ 2 resolution, which are both complementary to the existing 6\sarc\ \ Deep Fields Data Release 1 (DR1).
+Again this has an advantage over VLBI with instruments at GHz frequencies, which require either observing at lower frequency bands (e.g., e-MERLIN) or with other instruments (e.g., EVN, VLBA) to acquire information on the flux density of sources at other scales.
+As described in \S~5 of \citep{morabito_identifying_2022}, for sources which are unresolved in the 0.\sarc\ 3 resolution images and do not exhibit a radio excess over that predicted the star formation rate (SFR) derived from SED fitting \citep[see][for more details]{best_lofar_2023}, we can separate the AGN activity from SF using a combination of the 0.\sarc\ 3 and 6\sarc\ \ resolution images.
+
+The 2,483 sources in Lockman Hole \citep{sweijen_deep_2022} and now the 13,058 sources in ELAIS-N1 \citep{de_jong_into_2024}, represents a massive step towards robust statistical studies where the population can be broken into bins of redshift and stellar mass to investigate AGN activity and SF.\cite{cochrane_lofar_2023} and \cite{kondapally_cosmic_2022} constructed radio luminosity functions (RLFs) from the 6\sarc\ \ resolution imaging that formed the Deep Fields DR1 for the star forming galaxies and radio-excess AGN, respectively.
+Here we are able to reproduce those RLFs and further construct RLFS by activity -- either AGN or SF -- rather than by overall galaxy classification.
+This is the first time such RLFs have been constructed from observations, and provides insight into the non-radio excess population.
+These RLFs by activity also afford us the opportunity for a more direct comparison with cosmological simulations like SIMBA \hl{Nicole, citations} or EAGLE \hl{citations}.
+We can also investigate the cosmic evolution of AGN activity and SF by dividing the sample into redshift bins.
+
+In this paper, we first describe the data and methods in \S~\ref{sec:data}. We use a ...
+
+Throughout this paper, we use a $H_0 = 70$ km$\,$s$^{-1}\,$Mpc$^{-1}$, $\Omega_M = 0.3$, and $\Omega_{\Lambda}= 0.7$ cosmology.
+
+%Stuff and things \citep{morabito_sub-arcsecond_2022}.
+
+\section{Data}
+\label{sec:data}
+
+\subsection{LoTSS Deep Fields DR1}
+\label{subsec:dr1}
+The LOFAR Two-metre Sky Survey (LoTSS) has both a wide-area \citep{shimwell_lofar_2017,shimwell_lofar_2021,shimwell_lofar_2022} and Deep Fields \citep{tasse_lofar_2021,sabater_lofar_2021} component.
+The Deep Fields Data Release 1 (DR1) in 2021 included Boötes, Lockman Hole, and ELAI-N1, reaching sensitivities of 30, 23, and 20 \muJybm , using $\sim$80, $\sim$100, and $\sim$160 hours, respectively.
+Alongside these deep radio images and their associated catalogues, \cite{kondapally_lofar_2021} provided a careful compilation of the ancillary deep data at other wavebands (from far-infrared to ultraviolet) and cross-matching to the radio catalogues, followed by photometric redshifts and stellar mass estimates in \cite{duncan_lofar_2021}.
+Collectively these catalogues provide an excellent resource, and we refer the reader for more details to the individual papers cited above.
+
+Based on the cross-matched multi-wavelength catalogue \citep{kondapally_lofar_2021}, \cite{best_lofar_2023} carried out detailed SED fitting for all radio-detected sources across the three deep fields, using the multi-wavelength data.
+Full details can be found in their paper, but here we outline the relevant details.
+All sources detected above 5$\sigma$ in their respective radio images had their far-infrared (FIR) to ultraviolet (UV) data fit using four different SED fitting software packages.
+These were: MAGPHYS \citep{da_cunha_simple_2008}, BAGPIPES \citep{carnall_inferring_2018,carnall_how_2019}, CIGALE \citep{burgarella_star_2005,noll_analysis_2009,boquien_cigale_2019}, and AGN\textsc{fitter} \citep{calistro_rivera_agnfitter_2016}.
+The first two (MAGPHYS and BAGPIPES) use an energy balance approach, where optical to UV energy absorbed by dust must be transferred to thermal emission in the FIR to sub-mm.
+CIGALE also uses the same energy-balance approach, but includes an AGN component.
+The final code, AGN\textsc{fitter} is designed to work in cases where the UV light and IR emission are spatially distinct, and thus the energy balance assumption does not hold.
+Many AGN fall under this umbrella.
+Each galaxy was fit with all four codes (twice for CIGALE, using different AGN components) and a `consensus' value found for the derived galaxy properties using information from the goodness-of-fits.
+
+The derived consensus properties used in this paper are \hl{stellar masses} and SFRs.
+We also use the galaxy classifications which arise from the SED fitting.
+\cite{best_lofar_2023} use the relationship between radio luminosity and SFR to define sources as either \textit{radio excess} or not.
+The AGN components used to fit the multi-wavelength information indicate whether a source is a radiatively efficient AGN (i.e., is there evidence for a disk/torus structure) or not.
+Combining these provides the following classes:
+\begin{itemize}
+    \item Radio excess + radiatively efficient
+    \item Radio excess + not radiatively efficient
+    \item Radiatively efficient + not radio excess
+    \item Radiatively inefficient + not radio excess
+\end{itemize}
+The first three classes are AGN, and the final class is star forming galaxies (SFGs).
+The third class, radiatively efficient + not radio excess, are what are commonly called `radio-quiet' (RQ) AGN, while the first two AGN classes would be considered `radio-loud' (RL).
+We note here that this is an observational classification scheme where RL-AGN have radio emission which is not consistent with SF, but that does not mean that the RQ-AGN do not have an AGN component -- which we will return to in \S~\hl{tbd}.
+
+\subsection{0.\sarc\ 3 resolution imaging}
+\label{subsec:highres}
+
+Data for LoTSS is recorded with all available LOFAR stations, which are spread across nine different European countries, with a preponderance of \textit{core} and \textit{remote} stations at the heart of the array in the Netherlands.
+The longest baseline (currently Ireland to Latvia) is $\sim$2,000 km, providing unprecedented resolution at MHz frequencies.
+While LoTSS uses only the stations in the Netherlands for its standard data processing, it is becoming stahdard to use the full array to achieve sub-arcsecond resolution \citep[e.g.,]{morabito_sub-arcsecond_2022} on individual targets.
+We are currently post-processing LoTSS data to include the full international array where available.
+
+Post-processing of LoTSS data to achieve sub-arcsecond resolution ongoing for all Deep Fields, where we image the full field of view available to the international array.
+Due to the fact that the station layout of international stations is larger than that of stations in the Netherlands, the station beams naturally have a smaller field of view.
+Coupled with bandwidth and time smearing, this limits the field of view effectively to a radius of 1.24 deg, which still provides $\sim$5-6 deg$^2$ of coverage.
+
+The Lockman Hole was the first field to have its full field of view imaged at 0.\sarc\ 3 resolution \citep{sweijen_deep_2022}.
+This took a quarter million cpu hours to produce a single 7-billion pixel image from 8 hours of data.
+The resulting image has a median rms noise of 34 \muJybm , although this is lower in the centre and higher in the outskirts.
+It is useful to note that this image is deeper than an image at 6\sarc\ \ resolution using the exact same 8 hours of data, as adding the international stations increases the collecting area and therefore the sensitivity of the array.
+From this image, a catalogue of 2,316 5$\sigma$ sources was created, removing duplicates and grouping individual Gaussian islands together into sources.
+
+The first 0.\sarc\ 3 image of ELAIS-N1 has recently been published \citep{de_jong_into_2024}, along with companion images at intermediate resolutions of 0.\sarc\ 6 and 1.2\sarc .
+The median rms noise reaches 14 \muJybm\ in the centre of the 0.\sarc\ 3 resolution image, and increases radially.
+For consistency with the Lockman Hole sample, we do not use the intermediate resolution images here.
+In the 0.\sarc\ 3 resolution image, there were 13,058 sources catalogued above 5$\sigma$ after grouping Gaussian islands into sources and removing duplicates.
+
+Thanks to the added collecting area of the international stations, the 0.\sarc\ 3 resolution image made with 32 hours of data is now \textit{deeper} than the 6\sarc\ \ resolution image made with $\sim$160 hours of data.
+The derived galaxy properties from \cite{best_lofar_2023} (see \S~\ref{subsec:dr1}) were for LoTSS Deep Fields DR1, and thus we have to limit the source catalogue we wish to use.
+
+For both Lockman Hole and ELAIS-N1 catalogues derived from the 0.\sarc\ 3 resolution images, we first find the local rms in the 6\sarc\ \ resolution noise map\footnote{\label{dr1}Available at \href{https://lofar-surveys.org/deepfields.html}{https://lofar-surveys.org/deepfields.html}} at the position of each source.
+Using the peak brightness of the source in the catalogue from the 0.\sarc\ 3 resolution image, we calculate the peak-to-noise ratio and only select sources which would be detectable at the 5$\sigma$ limit.
+This ensures we do not have sources in the 0.\sarc\ 3 resolution catalogue which do not have a match in DR1.
+We then cross-match based on radio position, to provide the radio information from the catalogues derived from both the 0.\sarc\ 3 resolution image and the 6\sarc\ \ resolution image.
+We trim the area down for each field using their respective optical starmasks\textsuperscript{\ref{dr1}}, by selecting sources which are not contained in the null areas of these masks.
+This removes sources where the multi-wavelength information is impacted by bright stars.
+Finally, we trim the area down to the 2.5 deg$\times$2.5 deg area which is covered by the 0.\sarc\ 3 resolution image.
+
+Finally, we determine whether sources are resolved or unresolved in the same manner as \S~3.1 of \cite{shimwell_lofar_2019,shimwell_lofar_2022}.
+Truly unresolved sources in radio maps should have a peak brightness to integrated flux density ratio of unity, but LOFAR maps are impacted by smearing due to the ionosphere.
+This causes resolved sources to have a lower peak brightness to integrated flux density ratio, and they can appear slightly extended in the radio images.
+We therefore start by selecting sources which are fit by single Guassian ({\ttfamily S\_CODE = 'S'}), and making a cut to remove sources which are larger than a multiple of the beam size.
+For the Lockman Hole, we used a factor of 3, but for ELAIS-N1 we found that a factor of 4 provided a better sample.
+This may be because half of the data for ELAIS-N1 had been pre-averaged to 2 sec integration times, which will increase the time smearing \citep[see \S~2 in][]{de_jong_into_2024}.
+We fit a sigmoid function to find the upper 99.9th percentile of sources considering their integrated flux density to peak brightness ratio as a function of the local signal to noise.
+\hl{The fits and their fit parameters are shown} in Appendix~\ref{app:unresolved}.
+
+We are left with \hl{fill in} sources in Lockman Hole, and \hl{fill in} sources in ELAIS-N1.
+We use the source name from the catalogue derived from the 6\sarc\ \ resolution to cross-match to the catalogue from \cite{best_lofar_2023}.
+
+\subsection{Simulations???}
+
+\section{Methods}
+
+\subsection{Separation of AGN activity and SF}
+
+\subsection{Construction of RLFs}
 
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.5\textwidth]{figures/simba_comparison.png}
-    \caption{Caption}
-    \label{fig:simba}
-    \script{simba_comparison.py}
-\end{figure}
+- sfrd, kinetic LF
+
+%\begin{figure}
+%    \centering
+%    \includegraphics[width=0.5\textwidth]{figures/simba_comparison.png}
+%    \caption{Caption}
+%    \label{fig:simba}
+%    \script{simba_comparison.py}
+%\end{figure}
 
@@
 
+LOFAR is the Low Frequency Array designed and constructed by ASTRON.
+It has observing, data processing, and data storage facilities in several countries, which are owned by various parties (each with their own funding sources), and which are collectively operated by the ILT foundation under a joint scientific policy.
+The ILT resources have benefited from the following recent major funding sources: CNRS-INSU, Observatoire de Paris and Université d'Orléans, France; BMBF, MIWF-NRW, MPG, Germany; Science Foundation Ireland (SFI), Department of Business, Enterprise and Innovation (DBEI), Ireland; NWO, The Netherlands; The Science and Technology Facilities Council, UK; Ministry of Science and Higher Education, Poland; The Istituto Nazionale di Astrofisica (INAF), Italy.
+
+This research made use of the Dutch national e-infrastructure with support of the SURF Cooperative (e-infra 180169) and the LOFAR e-infra group.
+The Jülich LOFAR Long Term Archive and the German LOFAR network are both coordinated and operated by the Jülich Supercomputing Centre (JSC), and computing resources on the supercomputer JUWELS at JSC were provided by the Gauss Centre for Supercomputing e.V.
+(grant CHTB00) through the John von Neumann Institute for Computing (NIC).
+
+This research made use of the University of Hertfordshire high-performance computing facility and the LOFAR-UK computing facility located at the University of Hertfordshire and supported by STFC [ST/P000096/1], and of the Italian LOFAR IT computing infrastructure supported and operated by INAF, and by the Physics Department of Turin university (under an agreement with Consorzio Interuniversitario per la Fisica Spaziale) at the C3S Supercomputing Centre, Italy.
+
+\section*{Appendix A}
+
+\begin{figure*}
+    \includegraphics[width=0.49\textwidth]{figures/Lockman_Unresolved_03.png}
+    \hfill
+    \includegraphics[width=0.49\textwidth]{figures/ELAIS_Unresolved_03.png}
+    \caption{\label{fig:unres}The data and sigmoid fits to determine whether sources are resolved or unresolved as described in \S~\ref{subsec:highres}.
+The \textit{left} panel shows Lockman Hole, and the \textit{right} panel shows ELAIS-N1.
+}
+\end{figure*}
+
+Figure~\ref{fig:unres} shows the integrated flux density to peak brightness ratio as a function of the local signal to noise, parameterised by $\frac{S_{peak}}{2\sigma_{peak}} + \frac{S_{int}}{2\sigma_{int}}$.
+The data is first binned evenly in log space, and the 99.9th percentile is found in each bin.
+We then fit a sigmoid function:
+\begin{equation}
+y = A + \frac{B}{1 + (x / 96.57)^C}, 
+\end{equation}
+where 96.57 is a fixed value that has worked well for LoTSS \citep{shimwell_lofar_2022}. For Lockman Hole, we find
+
+\clearpage
+\pagebreak
+
 \bibliographystyle{mnras}

### 2024-07-16 src/scripts/RLF_evolution_v2.py
update rlf_evolution

@@ -1,0 +1,255 @@
+import paths
+import numpy as np
+import matplotlib
+from matplotlib import pyplot as plt
+from astropy.table import Table, Column, join, vstack
+from helper_functions import *
+from astropy.io import fits
+## WMAP9 is Hinshaw et al. 2013, H_0=69.3, Omega=0.287
+
+##############################################################
+## Plotting housekeeping
+
+matplotlib.rcParams['legend.frameon'] = False
+matplotlib.rcParams['axes.labelsize'] = 'large'
+## set up some colours
+n = 255
+mycols = plt.cm.viridis(np.linspace(0, 1,n))
+mycols_m = plt.cm.inferno(np.linspace(0, 1,n))
+
+##############################################################
+
+## set a default spectral index
+si = -0.7
+zmin = 0.003  ## matches Mauch & Sadler 2007
+zmax = 0.3   ## matches Mauch & Sadler 2007
+dz = 0.0001  ## matches Cochrane and Kondapally
+
+## read in SIMBA
+Simba_SF = Table.read( paths.static / 'RLFS_50MYR_SF.csv', format='csv', delimiter=',' )
+Simba_AGN = Table.read( paths.static / 'RLFS_50MYR_AGN.csv', format='csv', delimiter=',' )
+## shift using spectral index
+Simba_SF['x']  = Simba_SF['x'] + np.log10( np.power( (144./1400.), si ) )
+Simba_AGN['x']  = Simba_AGN['x'] + np.log10( np.power( (144./1400.), si ) )
+## and convert from mag to Lum
+# mag = -2.5 * Lum + zeropoint
+# so Lum = mag / -2.5
+Simba_SF['Curve1'] = Simba_SF['Curve1']+np.log10(2.5)
+Simba_AGN['Curve2'] = Simba_AGN['Curve2']+np.log10(2.5)
+
+## read in data
+t = Table.read( paths.static / 'redshift_bins.csv', format='csv' )
+zbin_starts = t['zbin_starts']
+zbin_ends = t['zbin_ends']
+
[…]

### 2024-07-16 src/scripts/RLF_evolution_v2.py
plot area

@@ -71,7 +71,7 @@ for i in np.arange(0,5):
 lum_bin_cens = z_lum_bins[0][0:-1] + 0.5*(z_lum_bins[0][1]-z_lum_bins[0][0])
 
 
-fsizex = 10
+fsizex = 9
 fsizey = 5
 sbsizex = 0.8
 sbsizey = 0.8
@@ -135,7 +135,7 @@ p3.legend()
 p3.set_title('SIMBA RLFs',fontsize=18)
 
 ## simulations (bottom panel): ratio of RLFs by galaxies and activity
-p4 = plt.axes([0.14+sbsizex*fsizey/fsizex,0.1,sbsizex*fsizey/fsizex,0.4*sbsizey])
+p4 = plt.axes([0.06+sbsizex*fsizey/fsizex,0.1,sbsizex*fsizey/fsizex,0.4*sbsizey])
 p4.plot( (19,27), (1,1), color='gray', linestyle='dashed', linewidth=1.5 )
 for i in np.arange(0,len(z_lum_bins)):
     non_zero_idx = np.where( np.logical_and( z_sf_lum_func[i] < cutoff, z_gal_sf_lum_func[i] < cutoff ) )[0]
@@ -144,7 +144,8 @@ for i in np.arange(0,len(z_lum_bins)):
 p4.axes.set_xlim(plxlims)
 p4.axes.set_ylim((0.5,1.1))
 p4.set_xlabel('log'+r'$_{10}$'+'('+r'$L_{\mathrm{144 MHz}}$'+' W Hz'+r'$^{-1}$'+'])')
-p4.set_ylabel('Activity / Galaxy')
+p4.yaxis.set_visible(False)
+#p4.set_ylabel('Activity / Galaxy')
 
 fig.savefig(paths.figures / 'RLF_evolution_SF.png',dpi=300)
 fig.clear()

### 2024-07-16 src/scripts/RLF_evolution_v2.py
fiddling

@@ -73,7 +73,7 @@ lum_bin_cens = z_lum_bins[0][0:-1] + 0.5*(z_lum_bins[0][1]-z_lum_bins[0][0])
 
 fsizex = 9
 fsizey = 5
-sbsizex = 0.8
+sbsizex = 0.6
 sbsizey = 0.8
 plxlims = (20.1,27)
 plylims = (-7.5,-1)

### 2024-07-16 src/scripts/RLF_evolution_v2.py
plot ratio

@@ -90,7 +90,7 @@ cutoff = -2
 fig = plt.figure( figsize=(fsizex,fsizey) )
 
 ## Left panel (top): galaxies and activity together
-p1 = plt.axes([0.06,0.42,sbsizex*fsizey/fsizex,0.6*sbsizey])
+p1 = plt.axes([0.06,0.42,sbsizex*fsizey/fsizex,0.7*sbsizey])
 for i in np.arange(0,len(z_lum_bins)):
     non_zero = np.where( z_gal_sf_lum_func[i] < cutoff )[0]
     p1.plot( lum_bin_cens[non_zero], z_gal_sf_lum_func[i][non_zero], color=zcols_sf[i], linewidth=3, alpha=0.75, linestyle='dotted' )
@@ -109,7 +109,7 @@ p1.add_artist(l1)
 p1.set_title('Observed RLFs',fontsize=18)
 
 ## Left panel (bottom): ratio of the RLFs by galaxies and activity
-p2 = plt.axes([0.06,0.1,sbsizex*fsizey/fsizex,0.4*sbsizey])
+p2 = plt.axes([0.06,0.1,sbsizex*fsizey/fsizex,0.3*sbsizey])
 p2.plot( (19,27), (1,1), color='gray', linestyle='dashed', linewidth=1.5 )
 for i in np.arange(0,len(z_lum_bins)):
     non_zero_idx = np.where( np.logical_and( z_sf_lum_func[i] < cutoff, z_gal_sf_lum_func[i] < cutoff ) )[0]

### 2024-07-16 src/scripts/RLF_evolution_v2.py
plot ratio

@@ -90,7 +90,7 @@ cutoff = -2
 fig = plt.figure( figsize=(fsizex,fsizey) )
 
 ## Left panel (top): galaxies and activity together
-p1 = plt.axes([0.06,0.42,sbsizex*fsizey/fsizex,0.7*sbsizey])
+p1 = plt.axes([0.06,0.32,sbsizex*fsizey/fsizex,0.7*sbsizey])
 for i in np.arange(0,len(z_lum_bins)):
     non_zero = np.where( z_gal_sf_lum_func[i] < cutoff )[0]
     p1.plot( lum_bin_cens[non_zero], z_gal_sf_lum_func[i][non_zero], color=zcols_sf[i], linewidth=3, alpha=0.75, linestyle='dotted' )

### 2024-07-16 src/scripts/RLF_evolution_v2.py
plot ratio

@@ -121,7 +121,7 @@ p2.set_xlabel('log'+r'$_{10}$'+'('+r'$L_{\mathrm{144 MHz}}$'+' W Hz'+r'$^{-1}$'+
 p2.set_ylabel('Activity / Galaxy')
 
 ## simulations (top panel): comparison
-p3 = plt.axes([0.06+sbsizex*fsizey/fsizex,0.42,sbsizex*fsizey/fsizex,0.6*sbsizey])
+p3 = plt.axes([0.06+sbsizex*fsizey/fsizex,0.32,sbsizex*fsizey/fsizex,0.7*sbsizey])
 p3.plot( Simba_SF['x'], Simba_SF['Curve1'], color='green', linewidth=2.5, label='Simba SF' )
 p3.plot( Simba_AGN['x'], Simba_AGN['Curve2'], color='blue', linewidth=2.5, label='Simba AGN' )
 
@@ -135,7 +135,7 @@ p3.legend()
 p3.set_title('SIMBA RLFs',fontsize=18)
 
 ## simulations (bottom panel): ratio of RLFs by galaxies and activity
-p4 = plt.axes([0.06+sbsizex*fsizey/fsizex,0.1,sbsizex*fsizey/fsizex,0.4*sbsizey])
+p4 = plt.axes([0.06+sbsizex*fsizey/fsizex,0.1,sbsizex*fsizey/fsizex,0.3*sbsizey])
 p4.plot( (19,27), (1,1), color='gray', linestyle='dashed', linewidth=1.5 )
 for i in np.arange(0,len(z_lum_bins)):
     non_zero_idx = np.where( np.logical_and( z_sf_lum_func[i] < cutoff, z_gal_sf_lum_func[i] < cutoff ) )[0]

### 2024-07-16 src/scripts/RLF_evolution_v2.py
plot ratio

@@ -71,9 +71,9 @@ for i in np.arange(0,5):
 lum_bin_cens = z_lum_bins[0][0:-1] + 0.5*(z_lum_bins[0][1]-z_lum_bins[0][0])
 
 
-fsizex = 9
+fsizex = 7
 fsizey = 5
-sbsizex = 0.6
+sbsizex = 0.7
 sbsizey = 0.8
 plxlims = (20.1,27)
 plylims = (-7.5,-1)
@@ -153,6 +153,13 @@ plt.close()
 
 ##########################################################
 ## AGN
+
+fsizex = 9
+fsizey = 5
+sbsizex = 0.6
+sbsizey = 0.8
+plxlims = (20.1,27)
+plylims = (-7.5,-1)
 
 zcols_agn = mycols_m[np.arange(0,len(z_lum_bins))*int(n/len(z_lum_bins))]
 

### 2024-07-16 src/scripts/RLF_evolution_v2.py
plot ratio

@@ -73,7 +73,7 @@ lum_bin_cens = z_lum_bins[0][0:-1] + 0.5*(z_lum_bins[0][1]-z_lum_bins[0][0])
 
 fsizex = 7
 fsizey = 5
-sbsizex = 0.7
+sbsizex = 0.6
 sbsizey = 0.8
 plxlims = (20.1,27)
 plylims = (-7.5,-1)

### 2024-07-16 src/scripts/RLF_evolution_v2.py
plot ratio

@@ -90,7 +90,7 @@ cutoff = -2
 fig = plt.figure( figsize=(fsizex,fsizey) )
 
 ## Left panel (top): galaxies and activity together
-p1 = plt.axes([0.06,0.32,sbsizex*fsizey/fsizex,0.7*sbsizey])
+p1 = plt.axes([0.1,0.32,sbsizex*fsizey/fsizex,0.7*sbsizey])
 for i in np.arange(0,len(z_lum_bins)):
     non_zero = np.where( z_gal_sf_lum_func[i] < cutoff )[0]
     p1.plot( lum_bin_cens[non_zero], z_gal_sf_lum_func[i][non_zero], color=zcols_sf[i], linewidth=3, alpha=0.75, linestyle='dotted' )
@@ -109,7 +109,7 @@ p1.add_artist(l1)
 p1.set_title('Observed RLFs',fontsize=18)
 
 ## Left panel (bottom): ratio of the RLFs by galaxies and activity
-p2 = plt.axes([0.06,0.1,sbsizex*fsizey/fsizex,0.3*sbsizey])
+p2 = plt.axes([0.1,0.1,sbsizex*fsizey/fsizex,0.3*sbsizey])
 p2.plot( (19,27), (1,1), color='gray', linestyle='dashed', linewidth=1.5 )
 for i in np.arange(0,len(z_lum_bins)):
     non_zero_idx = np.where( np.logical_and( z_sf_lum_func[i] < cutoff, z_gal_sf_lum_func[i] < cutoff ) )[0]
@@ -121,7 +121,7 @@ p2.set_xlabel('log'+r'$_{10}$'+'('+r'$L_{\mathrm{144 MHz}}$'+' W Hz'+r'$^{-1}$'+
 p2.set_ylabel('Activity / Galaxy')
 
 ## simulations (top panel): comparison
-p3 = plt.axes([0.06+sbsizex*fsizey/fsizex,0.32,sbsizex*fsizey/fsizex,0.7*sbsizey])
+p3 = plt.axes([0.1+sbsizex*fsizey/fsizex,0.32,sbsizex*fsizey/fsizex,0.7*sbsizey])
 p3.plot( Simba_SF['x'], Simba_SF['Curve1'], color='green', linewidth=2.5, label='Simba SF' )
 p3.plot( Simba_AGN['x'], Simba_AGN['Curve2'], color='blue', linewidth=2.5, label='Simba AGN' )
 
[…]

### 2024-08-01 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{orcidlink}
+\usepackage{comment}
 
@@
 \newcommand{\muJybm}{$\mu$Jy$\,$beam$^{-1}$}
+
+\defcitealias{morabito_identifying_2022}{M22}
+\defcitealias{best_lofar_2023}{B23}
 
@@
 Traditionally brightness temperature measurements require Very Long Baseline Interferometry (VLBI) which achieve milli-arcsec resolutions at GHz frequencies \hl{citations} to achieve enough sensitivity to make meaningful measurements.
-In \cite[][heareafter, \hl{M22}]{morabito_identifying_2022} we demonstrated using sub-arcsecond resolution (0.\sarc\ 3) at 144 MHz with the LOw Frequency ARray \citep[LOFAR;][]{van_haarlem_lofar:_2013}.
+In \cite[][hereafter, \hl{M22}]{morabito_identifying_2022} we demonstrated using sub-arcsecond resolution (0.\sarc\ 3) at 144 MHz with the LOw Frequency ARray \citep[LOFAR;][]{van_haarlem_lofar:_2013}.
 The advantage of LOFAR is its wide field of view: $\sim$6 deg$^2$ from a single observation versus a few to a couple hundred arcmin$^2$ with VLBI at GHz frequencies, depending on whether one is imaging at the phase centre of the observation or using multi-phase centres for correlation \hl{citations}.
-For example, using an 8 hour LOFAR observation of the Lockman Hole \citep{sweijen_deep_2022} we identified 940 AGN via brightness temperature \cite{morabito_identifying_2022}, providing a sample two orders of magnitude larger than any VLBI sample of AGN identified at GHz frequencies.
+For example, using an 8 hour LOFAR observation of the Lockman Hole \citep{sweijen_deep_2022} we identified 940 AGN via brightness temperature \citepalias{morabito_identifying_2022}, providing a sample two orders of magnitude larger than any VLBI sample of AGN identified at GHz frequencies.
 
@@
 Again this has an advantage over VLBI with instruments at GHz frequencies, which require either observing at lower frequency bands (e.g., e-MERLIN) or with other instruments (e.g., EVN, VLBA) to acquire information on the flux density of sources at other scales.
-As described in \S~5 of \citep{morabito_identifying_2022}, for sources which are unresolved in the 0.\sarc\ 3 resolution images and do not exhibit a radio excess over that predicted the star formation rate (SFR) derived from SED fitting \citep[see][for more details]{best_lofar_2023}, we can separate the AGN activity from SF using a combination of the 0.\sarc\ 3 and 6\sarc\ \ resolution images.
+As described in \S~5 of \citepalias{morabito_identifying_2022}, for sources which are unresolved in the 0.\sarc\ 3 resolution images and do not exhibit a radio excess over that predicted the star formation rate (SFR) derived from SED fitting \citep[see][for more details]{best_lofar_2023}, we can separate the AGN activity from SF using a combination of the 0.\sarc\ 3 and 6\sarc\ \ resolution images.
 
@@
 
-Based on the cross-matched multi-wavelength catalogue \citep{kondapally_lofar_2021}, \cite{best_lofar_2023} carried out detailed SED fitting for all radio-detected sources across the three deep fields, using the multi-wavelength data.
+Based on the cross-matched multi-wavelength catalogue \citep{kondapally_lofar_2021}, \cite[][hereafter, B23]{best_lofar_2023} carried out detailed SED fitting for all radio-detected sources across the three deep fields, using the multi-wavelength data.
 Full details can be found in their paper, but here we outline the relevant details.
@@
 We also use the galaxy classifications which arise from the SED fitting.
-\cite{best_lofar_2023} use the relationship between radio luminosity and SFR to define sources as either \textit{radio excess} or not.
+\citetalias{best_lofar_2023} use the relationship between radio luminosity and SFR to define sources as either \textit{radio excess} or not.
 The AGN components used to fit the multi-wavelength information indicate whether a source is a radiatively efficient AGN (i.e., is there evidence for a disk/torus structure) or not.
@@
 The longest baseline (currently Ireland to Latvia) is $\sim$2,000 km, providing unprecedented resolution at MHz frequencies.
-While LoTSS uses only the stations in the Netherlands for its standard data processing, it is becoming stahdard to use the full array to achieve sub-arcsecond resolution \citep[e.g.,]{morabito_sub-arcsecond_2022} on individual targets.
+While LoTSS uses only the stations in the Netherlands for its standard data processing, it is becoming stahdard to use the full array to achieve sub-arcsecond resolution \citep[e.g.,][]{morabito_sub-arcsecond_2022} on individual targets.
 We are currently post-processing LoTSS data to include the full international array where available.
@@
 Thanks to the added collecting area of the international stations, the 0.\sarc\ 3 resolution image made with 32 hours of data is now \textit{deeper} than the 6\sarc\ \ resolution image made with $\sim$160 hours of data.
-The derived galaxy properties from \cite{best_lofar_2023} (see \S~\ref{subsec:dr1}) were for LoTSS Deep Fields DR1, and thus we have to limit the source catalogue we wish to use.
+The derived galaxy properties from \citetalias{best_lofar_2023} (see \S~\ref{subsec:dr1}) were for LoTSS Deep Fields DR1, and thus we have to limit the source catalogue we wish to use.
 
@@
 This causes resolved sources to have a lower peak brightness to integrated flux density ratio, and they can appear slightly extended in the radio images.
-We therefore start by selecting sources which are fit by single Guassian ({\ttfamily S\_CODE = 'S'}), and making a cut to remove sources which are larger than a multiple of the beam size.
-For the Lockman Hole, we used a factor of 3, but for ELAIS-N1 we found that a factor of 4 provided a better sample.
+We start by selecting sources which are fit by single Guassian ({\ttfamily S\_CODE = 'S'}), and make a cut to remove sources which are larger than a multiple of the beam size.
+For the Lockman Hole, we use a factor of 3, but for ELAIS-N1 we find that a factor of 4 provided a better sample.
 This may be because half of the data for ELAIS-N1 had been pre-averaged to 2 sec integration times, which will increase the time smearing \citep[see \S~2 in][]{de_jong_into_2024}.
 We fit a sigmoid function to find the upper 99.9th percentile of sources considering their integrated flux density to peak brightness ratio as a function of the local signal to noise.
-\hl{The fits and their fit parameters are shown} in Appendix~\ref{app:unresolved}.
+\hl{The fit parameters and graphical representation of the data can be found in the supplemental online materials.}
 
 We are left with \hl{fill in} sources in Lockman Hole, and \hl{fill in} sources in ELAIS-N1.
-We use the source name from the catalogue derived from the 6\sarc\ \ resolution to cross-match to the catalogue from \cite{best_lofar_2023}.
+We use the source name from the catalogue derived from the 6\sarc\ \ resolution to cross-match to the catalogue from \citetalias{best_lofar_2023}.
 
-\subsection{Simulations???}
+\subsection{Simulations}
+\hl{Nicole}
 
@@
 
+Brightness temperature, $T_b$, is used to characterise the surface brightness of radio emission.
+It is not a physical temperature, but rather the temperature of a black body  which would produce the observed spectral radiance (assuming the Rayleigh-Jeans approximation) at long wavelengths.
+By creating a model which predicts the upper limit of $T_b$ achievable by star formation \citep[based on][]{condon_radio_1992}, if the value is \textit{above} that upper limit, the radio emission must be due to AGN activity.
+\citetalias{morabito_identifying_2022} showed that this AGN identification can be done at 150 MHz using sub-arcsecond resolution; this works because $T_b\propto\nu^{-2}\Omega^{-1}$ where $\nu$ is the observed frequency and $\Omega$ is the solid angle of the emitting region -- for unresolved sources, this corresponds to the deconvolved beam size.
+We refer the reader to \S~3 of \citetalias{morabito_identifying_2022} for more details, but note that this method yields conservative, positive AGN identifications.
+There may be lower surface brightness AGN emission which is not identified, and it is also possible \citep[although unlikely, according to][]{condon_radio_1992} that extreme, compact nuclear starbursts could produce AGN-like values of $T_b$.
+
 \subsection{Separation of AGN activity and SF}
+Once the AGN activity is identified from the 0.\sarc\ 3 resolution images, it is possible to separate it from star formation using the total flux density from the 6\sarc\ \ resolution images.
+We refer the reader to \S~5 in \citetalias{morabito_identifying_2022} for more details but outline the process here.
+Sources which are identified as radio excess according to the definition in \citetalias{best_lofar_2023} automatically have their total flux density from the 6\sarc\ resolution images added to the AGN category.
+For sources that are not radio excess, we attribute the total flux density from the 6\sarc\ resolution images to SF if in the 0.\sarc\ \ 3 resolution image they are: resolved, multi-component, or have no detection.
+Sources which do not meet these criteria (i.e., unresolved single component) are checked and if their $T_b$ is not above the limit for SF their 6\sarc\ resolution total flux density is also added to the SF category.
+The total flux density of any source in the 0.\sarc\ 3 resolution image which is $T_b$-identified AGN is assumed to be AGN flux density; this is subtracted from the total flux density measured from the 6\sarc\ \ resolution image, and what remains is assumed to be flux density from SF.
+Fig.~\ref{fig:flowchart} shows a flowchart of how flux density is assigned to either the AGN or SF category.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{figures/agn_sf_separation_workflow.png}
+    \caption{Caption}
+    \label{fig:flowchart}
+\end{figure}
 
@@
 
+We follow the same method of constructing RLFs as in \cite{kondapally_cosmic_2022} and \cite{cochrane_lofar_2023}.
+As the 0.\sarc\ \ 3 resolution image is smaller than the 6\sarc\ resolution image, we first reproduce the published RLFs (Fig.~\ref{fig:rlfs}, left panel) using the galaxy classifications from \citetalias{best_lofar_2023} to check that we have implemented the method appropriately, and that the smaller area yields the same results.
+We use the completeness corrections from the respective papers.
+The only correction we do not implement is the photo-$z$ correction used in \cite{cochrane_lofar_2023}.
+The authors note that this correction is negligible, and as it is not implemented in \cite{kondapally_cosmic_2022} we chose to keep the method for RLF construction in this paper consistent across both samples.
+We use the same redshift range, $0.003 < z < 0.3$ and see that we indeed have good agreement with the published RLFs.
+
+Next we use the same method for constructing RLFs by activity rather than overall galaxy classifcation.
+We again use the completeness corrections from \cite{cochrane_lofar_2023} and \cite{kondapally_cosmic_2022}.
+The completeness corrections are flux density dependent, and we use the overall total flux density of a galaxy, rather than the flux density of the activity (which may be smaller if there are both AGN and SF components).
+This is because completeness corrections are based on whether or not a \textit{source} will be detectable given the background rms.
+This only relies on the total flux density, not the flux density of individual components.
+The RLFs by activity rather than galaxy classification are shown in the middle panel of Fig.~\ref{fig:rlfs}, again for  $0.003 < z < 0.3$.
+
+Finally, we use RLFs constructed from SIMBA. The RLFs are constructed as in \cite{thomson_radio_2019}, but have been updated to use the same redshift bins \hl{Nicole}.
+
 \section{Results}
 
-- sfrd, kinetic LF
+To illustrate the differences in RLFs constructed by activity vs.
+by galaxy classification we divide the activity RLF by the galaxy RLF and show the ratio in the bottom right hand panel in Fig.~\ref{fig:rlfs}, for both the AGN and SF category.
+We see that above 10$^{24}\,$\wphz\ the ratio for AGN is unity, which is expected where radio excess sources dominate the population.
+Below that, the ratio for AGN climbs up to $\sim$1.25 and finally almost up to 2 at the lowest radio powers probed.
+The ratio for SF is always below unity, but it does not deviate as much as the ratio for AGN.
+This is because the SF population has larger source counts, so the total impact of moving some flux density from the SF to AGN population is less extreme.
+
+In Fig.~\ref{fig:agn_evolution} we show
+
+%- sfrd, kinetic LF
 
@@
     \caption{Caption}
-    \label{fig:msrlfs}
+    \label{fig:rlfs}
     \script{comparison_plot.py}
@@
     \centering
-    \includegraphics[width=0.98\textwidth]{figures/RLF_evolution_AGN.png}
-    \caption{Caption}
-    \label{fig:agn_evolution}
-    \script{RLF_evolution.py}
-\end{figure*}
-
-\begin{figure*}
-    \centering
-    \includegraphics[width=0.98\textwidth]{figures/RLF_evolution_SF.png}
+    \includegraphics[width=0.49\textwidth]{figures/RLF_evolution_SF.png}
+    \includegraphics[width=0.49\textwidth]{figures/RLF_evolution_SF.png}
     \caption{Caption}
     \label{fig:sf_evolution}
-    \script{RLF_evolution.py}
+    \script{RLF_evolution_v2.py}
 \end{figure*}
@@
 \section*{Acknowledgements}
-LKM is grateful for support from UKRI [MR/T042842/1]. \hl{also cosma and spider?}
+LKM is grateful for support from UKRI [MR/T042842/1].
 
@@
 
+\begin{comment}
 \section*{Appendix A}
@@
 where 96.57 is a fixed value that has worked well for LoTSS \citep{shimwell_lofar_2022}. For Lockman Hole, we find 
+\end{comment}
 

### 2024-08-01 src/scripts/RLF_evolution_v2.py
new simba data

@@ -15,6 +15,7 @@ matplotlib.rcParams['axes.labelsize'] = 'large'
 ## set up some colours
 n = 255
 mycols = plt.cm.viridis(np.linspace(0, 1,n))
+simba_mycols = plt.cm.cividis(np.linspace(0,1,n))
 mycols_m = plt.cm.inferno(np.linspace(0, 1,n))
 
 ##############################################################
@@ -26,6 +27,7 @@ zmax = 0.3   ## matches Mauch & Sadler 2007
 dz = 0.0001  ## matches Cochrane and Kondapally
 
 ## read in SIMBA
+'''
 Simba_SF = Table.read( paths.static / 'RLFS_50MYR_SF.csv', format='csv', delimiter=',' )
 Simba_AGN = Table.read( paths.static / 'RLFS_50MYR_AGN.csv', format='csv', delimiter=',' )
 ## shift using spectral index
@@ -36,6 +38,16 @@ Simba_AGN['x']  = Simba_AGN['x'] + np.log10( np.power( (144./1400.), si ) )
 # so Lum = mag / -2.5
 Simba_SF['Curve1'] = Simba_SF['Curve1']+np.log10(2.5)
 Simba_AGN['Curve2'] = Simba_AGN['Curve2']+np.log10(2.5)
+'''
+
+## read in SIMBA and convert to same values
+Simba = Table.read( paths.static / 'RLFs_flux_lim.csv', format='csv', delimiter=',' )
+Simba['logP'] = Simba['logP'] + np.log10( np.power( (144./1400.), si ) )
+for cn in ['SF','AGN','SFGs','Qs']:
+    Simba[cn] = Simba[cn]+np.log10(2.5)
+
+simba_z_bin_cens = np.unique(Simba['z'])
+simba_z_bin_cens = simba_z_bin_cens[1:]
 
 ## read in data
 t = Table.read( paths.static / 'redshift_bins.csv', format='csv' )
@@ -83,6 +95,9 @@ plylims = (-7.5,-1)
 ## Star formation
 
 zcols_sf = mycols[np.arange(0,len(z_lum_bins))*int(n/len(z_lum_bins))]
+simba_zcols = simba_mycols[np.arange(0,len(z_lum_bins))*int(n/len(z_lum_bins))]
+
+ratio_ylims = (0,1.2)
[…]

### 2024-08-01 src/scripts/simba_comparison.py

@@ -32,6 +32,18 @@ Simba_AGN = Table.read( paths.static / 'RLFS_50MYR_AGN.csv', format='csv', delim
 Simba_SF['x']  = Simba_SF['x'] + np.log10( np.power( (144./1400.), si ) )
 Simba_AGN['x']  = Simba_AGN['x'] + np.log10( np.power( (144./1400.), si ) )
 
+'''
+## new simba check
+Simba = Table.read( paths.static / 'RLFs_no_flux_lim.csv', format='csv', delimiter=',' )
+z_idx = np.where( Simba['z'] == 0.0 )[0]
+Simba = Simba[z_idx]
+Simba_SF = Table()
+Simba_SF.add_column( Column( Simba['logP']+np.log10( np.power( (144./1400.), si ) ), name='x' ) )
+Simba_SF.add_column( Column( Simba['SF'], name='Curve1' ) )
+Simba_AGN = Table()
+Simba_AGN.add_column( Column( Simba['logP']+np.log10( np.power( (144./1400.), si ) ), name='x' ) )
+Simba_AGN.add_column( Column( Simba['AGN'], name='Curve2' ) )
+'''
 
 ## read in vmaxes
 lockman_vmaxes = Table.read( paths.static / 'lockman_vmaxes_zmin0.003_zmax0.3.fits', format='fits' )

### 2024-08-01 src/scripts/simba_comparison.py
change simba files

@@ -26,15 +26,16 @@ zmax = 0.3   ## matches Mauch & Sadler 2007
 dz = 0.0001  ## matches Cochrane and Kondapally
 
 ## read in SIMBA
+'''
 Simba_SF = Table.read( paths.static / 'RLFS_50MYR_SF.csv', format='csv', delimiter=',' )
 Simba_AGN = Table.read( paths.static / 'RLFS_50MYR_AGN.csv', format='csv', delimiter=',' )
 ## shift using spectral index
 Simba_SF['x']  = Simba_SF['x'] + np.log10( np.power( (144./1400.), si ) )
 Simba_AGN['x']  = Simba_AGN['x'] + np.log10( np.power( (144./1400.), si ) )
-
 '''
+
 ## new simba check
-Simba = Table.read( paths.static / 'RLFs_no_flux_lim.csv', format='csv', delimiter=',' )
+Simba = Table.read( paths.static / 'RLFs_flux_lim.csv', format='csv', delimiter=',' )
 z_idx = np.where( Simba['z'] == 0.0 )[0]
 Simba = Simba[z_idx]
 Simba_SF = Table()
@@ -43,7 +44,7 @@ Simba_SF.add_column( Column( Simba['SF'], name='Curve1' ) )
 Simba_AGN = Table()
 Simba_AGN.add_column( Column( Simba['logP']+np.log10( np.power( (144./1400.), si ) ), name='x' ) )
 Simba_AGN.add_column( Column( Simba['AGN'], name='Curve2' ) )
-'''
+
 
 ## read in vmaxes
 lockman_vmaxes = Table.read( paths.static / 'lockman_vmaxes_zmin0.003_zmax0.3.fits', format='fits' )

### 2024-08-21 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \newcommand{\muJybm}{$\mu$Jy$\,$beam$^{-1}$}
+\newcommand{\llof}{$L_{\textrm{144MHz}}$}
 
@@
 % Title
-\title[TBD]{TBD}
+\title[TBD]{Direct comparison between simulations and observations: radio emission separated into star formation and AGN activity}
 
@@
 \subsection{Simulations}
-\hl{Nicole}
-
+\label{subsec:simba}
+\hl{Nicole}\\
+We use the SIMBA cosmological hydrodynamic simulation which accounts for an H$_{2}$ based star formation rate following the sub-grid model of \hl{Krumholtz $\&$ Gnedin (2011)}.
+Additionally, SIMBA models the growth of supermassive black holes via a two-mode prescription based on the temperature of gas within a distance, $R_{0}$, calculated over the 256 nearest particles limited to 2$h^{-1}$kpc.
+It is commonplace for cosmological simulations to employ the growth of SMBHs via the Bondi rate \hl{citations EAGLE, HORIZON AGN, Illustris, TNG}.
+In SIMBA, the Bondi rate is employed when gas within $R_{0}$ has temperatures T$>10^{5}$K, referred to as the hot mode.
+What makes SIMBA unique is the additional contribution to the accretion rate via a torque limited accretion flow when gas has temperatures T$<10^{5}$K, or in the cold mode.
+SIMBA also employs a physically motivated model for AGN feedback in the form of AGN winds \hl{cite Perna}, as well as AGN jets with velocities that increase inversely proportional to the Eddington rate in accordance with \hl{Heckman and Best 2014}.
+SIMBA therefore provides a platform to study jet-mode AGN in a way that has not been explored before. 
+Specifically, we select galaxies in SIMBA that have full-velocity jets, that is, when the Eddington rate drops below 0.02 (but non-zero) and where the black hole mass is larger than M$_{\rm BH}>10^{8}$M$_{\odot}$.
+For this population we are able to calculate the amount of radio emission due to star formation processes following \hl{Condon 1992} as well as that from accretion processes following \hl{Kording, Jester and Fender 2008}.
+Details of this method can be found in \hl{Thomas2021}.
+We can therefore calculate the contribution to the radio luminosity function from star formation and AGN accretion processes.
+We can additionally separate the galaxy population into star forming and quiescent galaxies, radio-excess, or radiatively efficient and inefficient, based on the sources black hole and host properties.
 \section{Methods}
@@
 
-In Fig.~\ref{fig:agn_evolution} we show
-
-%- sfrd, kinetic LF
-
-%\begin{figure}
-%    \centering
-%    \includegraphics[width=0.5\textwidth]{figures/simba_comparison.png}
-%    \caption{Caption}
-%    \label{fig:simba}
-%    \script{simba_comparison.py}
-%\end{figure}
-
 \begin{figure*}
@@
 
+In Fig.~\ref{fig:evolution} we show a comparison between the observed RLFs and those generated from SIMBA (see \S~\ref{subsec:simba}), and the redshift evolution for both.
+The left set of panels shows the results for SF and the right set of panels shows the results for AGN.
+\hl{We cut off the RLFs below a flux limit where they are not constrained}.
+We use the same redshift bins across all samples for consistency.
+We reproduce previous observational results that while the SF population has a strong redshift evolution, the AGN population does not.
+For the simulations, the behaviour is qualitatively similar, although the AGN population does have mild redshift evolution whereas the observations do not.
+
+The bottom panels in Fig~\ref{fig:evolution} show the ratio of activity to galaxy RLFs, the same as in the bottom right hand panel of Fig~\ref{fig:rlfs}.
+Comparing these ratios between observations and simulations, we see qualitatively the same behaviour, although the simulations reach much lower and higher values for both the SF and AGN populations, respectively.
+In the lowest redshift bin, the ratio for the simulation stabilises to around unity for \llof $<10^{23}\,$\wphz , and decreases rapidly above this, reaching much smaller values than the observations.
+Higher redshift bins probe successively higher luminosity ranges, also reaching small values for the ratio between activity and galaxies.
+This is consistent with the fact that at higher redshifts, most radio AGN in SIMBA are hosted in SFGs, and thus the \textit{galaxy} RLF for SIMBA SFGs is inflated with AGN emission.
+
+For the AGN population (Fig~\ref{fig:evolution}, right set of panels), the ratio between observed activity and galaxy RLFs  converges to unity above $\sim$\llof $10^{24}\,$\wphz\ for the lowest redshift bin.
+This convergence happens at higher luminosity values for higher redshifts.
+Below these luminosities, across all redshift bins, the ratio increases.
+The maximum value for each redshift bin approaches $\sim$2, although the fact that in the lowest redshift bin the value stabilises at $\sim1.3$ and only climbs higher at the very lowest luminosities (\llof $<10^{22}\,$\wphz ) may be due to incompleteness.
+If we disregard the lowest luminosity bins, it seems that we see a 30 - 50 percent increase in AGN activity over what is measured when using galaxy classifications alone, for all redshifts.
+The ratio for the SIMBA RLFs shows strong redshift evolution, and much higher ratio values.
+The RLFs themselves turn over, whereas the observed RLFs do not.
+This is likely a consequence of the fact that the AGN emission in SIMBA is due to high-velocity jets, whereas there is increasing observational evidence for low-luminosity jets or AGN emission \citep{macfarlane_radio_2021,yue_novel_2024}.
+The high values reached by the ratio of RLFs again is expected as many high-$z$ radio AGN in SIMBA are hosted in SFGs.
+Shifting this AGN emission to AGN \textit{activity} when it is counted in the SFG \textit{galaxy} population generates high values for the activity to galaxy ratio.
+
+%- sfrd, kinetic LF
+
 \begin{figure*}
@@
     \includegraphics[width=0.49\textwidth]{figures/RLF_evolution_SF.png}
-    \includegraphics[width=0.49\textwidth]{figures/RLF_evolution_SF.png}
+    \includegraphics[width=0.49\textwidth]{figures/RLF_evolution_AGN.png}
     \caption{Caption}
-    \label{fig:sf_evolution}
+    \label{fig:evolution}
     \script{RLF_evolution_v2.py}
@@
 
-\section{Discussion and conclusions}
+\section{Discussion}
+
+\section{Conclusions}
 

### 2024-08-21 src/scripts/RLF_evolution_v2.py
new files

@@ -184,7 +184,7 @@ plt.close()
 
 plxlims = (20.1,27)
 plylims = (-7.5,-1)
-ratio_ylims = (0.7,2.5)
+ratio_ylims = (0.7,10)
 
 zcols_agn = mycols_m[np.arange(0,len(z_lum_bins))*int(n/len(z_lum_bins))]
 
@@ -219,6 +219,7 @@ for i in np.arange(0,len(z_lum_bins)):
     p2.plot( lum_bin_cens[non_zero_idx], ratio, color=zcols_agn[i], linewidth=3 )
 p2.axes.set_xlim(plxlims)
 p2.axes.set_ylim(ratio_ylims)
+p2.axes.set_yscale('log')
 p2.set_xlabel('log'+r'$_{10}$'+'('+r'$L_{\mathrm{144 MHz}}$'+' W Hz'+r'$^{-1}$'+'])')
 p2.set_ylabel('Activity / Galaxy')
 
@@ -227,8 +228,8 @@ p3 = plt.axes([0.1+sbsizex*fsizey/fsizex,0.32,sbsizex*fsizey/fsizex,0.7*sbsizey]
 for i in np.arange(0,len(simba_z_bin_cens)):
     z_idx = np.where( Simba['z'] == simba_z_bin_cens[i] )[0]
     tmp_Simba = Simba[z_idx]
-    non_zero_idx = np.where( np.logical_and( np.isfinite(tmp_Simba['AGN']), np.isfinite(tmp_Simba['Qs']) ) )[0]
-    p3.plot( tmp_Simba['logP'][non_zero_idx], tmp_Simba['Qs'][non_zero_idx], color=simba_zcols[i], linewidth=2.5, linestyle='dotted' )
+    non_zero_idx = np.where( np.logical_and( np.isfinite(tmp_Simba['AGN']), np.isfinite(tmp_Simba['REQs']) ) )[0]
+    p3.plot( tmp_Simba['logP'][non_zero_idx], tmp_Simba['REQs'][non_zero_idx], color=simba_zcols[i], linewidth=2.5, linestyle='dotted' )
     p3.plot( tmp_Simba['logP'][non_zero_idx], tmp_Simba['AGN'][non_zero_idx], color=simba_zcols[i], linewidth=2.5, label='{:s} < z < {:s}'.format(str(zbin_starts[i]),str(zbin_ends[i])) )
[…]

### 2024-08-23 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Title
-\title[TBD]{Direct comparison between simulations and observations: radio emission separated into star formation and AGN activity}
+\title[TBD]{A hidden Active Galactic Nuclei population: the first radio luminosity functions by process}
 
@@
 Observations have revealed tight scaling relations between the mass of a super-massive black hole and its host galaxy \citep[see, e.g.][and references therein]{kormendy_coevolution_2013}, while cosmological simulations require some form of AGN feedback \citep{bower_breaking_2006,croton_many_2006} to be able to reproduce the observed galaxy population.
-Galaxies build up their stellar mass through star formation, while super-massive black holes are expected to provide either negative or positive feedback \hl{citations!} to either suppress or stimulate galaxy growth.
+Galaxies build up their stellar mass through star formation, while super-massive black holes are expected to provide either negative or positive feedback to either suppress or stimulate galaxy growth, although whether we can definitively pinpoint which type of feedback is an open question \citep{ward_cosmological_2022}.
 
@@
 \label{subsec:simba}
-\hl{Nicole}\\
-We use the SIMBA cosmological hydrodynamic simulation which accounts for an H$_{2}$ based star formation rate following the sub-grid model of \hl{Krumholtz $\&$ Gnedin (2011)}.
+
+We use the SIMBA cosmological hydrodynamic simulation which accounts for an H$_{2}$ based star formation rate following the sub-grid model of \cite{krumholz_comparison_2011}.
 Additionally, SIMBA models the growth of supermassive black holes via a two-mode prescription based on the temperature of gas within a distance, $R_{0}$, calculated over the 256 nearest particles limited to 2$h^{-1}$kpc.
@@
 What makes SIMBA unique is the additional contribution to the accretion rate via a torque limited accretion flow when gas has temperatures T$<10^{5}$K, or in the cold mode.
-SIMBA also employs a physically motivated model for AGN feedback in the form of AGN winds \hl{cite Perna}, as well as AGN jets with velocities that increase inversely proportional to the Eddington rate in accordance with \hl{Heckman and Best 2014}.
+SIMBA also employs a physically motivated model for AGN feedback in the form of AGN winds \citep{perna_x-raysdss_2017}, as well as AGN jets with velocities that increase inversely proportional to the Eddington rate in accordance with \cite{heckman_coevolution_2014}.
 SIMBA therefore provides a platform to study jet-mode AGN in a way that has not been explored before. 
 Specifically, we select galaxies in SIMBA that have full-velocity jets, that is, when the Eddington rate drops below 0.02 (but non-zero) and where the black hole mass is larger than M$_{\rm BH}>10^{8}$M$_{\odot}$.
-For this population we are able to calculate the amount of radio emission due to star formation processes following \hl{Condon 1992} as well as that from accretion processes following \hl{Kording, Jester and Fender 2008}.
-Details of this method can be found in \hl{Thomas2021}.
+For this population we are able to calculate the amount of radio emission due to star formation processes following \cite{condon_radio_1992} as well as that from accretion processes following \cite{kording_measuring_2008}.
+Details of this method can be found in \cite{thomas_radio_2021}.
 We can therefore calculate the contribution to the radio luminosity function from star formation and AGN accretion processes.
 We can additionally separate the galaxy population into star forming and quiescent galaxies, radio-excess, or radiatively efficient and inefficient, based on the sources black hole and host properties.
+
 \section{Methods}
@@
 To illustrate the differences in RLFs constructed by activity vs.
-by galaxy classification we divide the activity RLF by the galaxy RLF and show the ratio in the bottom right hand panel in Fig.~\ref{fig:rlfs}, for both the AGN and SF category.
+by galaxy classification we divide the activity RLF by the galaxy RLF for the observations and show the ratio in the bottom right hand panel in Fig.~\ref{fig:rlfs}, for both the AGN and SF category.
 We see that above 10$^{24}\,$\wphz\ the ratio for AGN is unity, which is expected where radio excess sources dominate the population.
@@
 In Fig.~\ref{fig:evolution} we show a comparison between the observed RLFs and those generated from SIMBA (see \S~\ref{subsec:simba}), and the redshift evolution for both.
+We use similar redshift bins to \cite{kondapally_cosmic_2022} and \cite{cochrane_lofar_2023} for consistency.
 The left set of panels shows the results for SF and the right set of panels shows the results for AGN.
@@
 
-The bottom panels in Fig~\ref{fig:evolution} show the ratio of activity to galaxy RLFs, the same as in the bottom right hand panel of Fig~\ref{fig:rlfs}.
+The bottom panels in Fig~\ref{fig:evolution} show the ratio of activity to galaxy RLFs, the same as in the bottom right hand panel of Fig~\ref{fig:rlfs} (although note that the y-scale is logarithmic for the AGN).
 Comparing these ratios between observations and simulations, we see qualitatively the same behaviour, although the simulations reach much lower and higher values for both the SF and AGN populations, respectively.
-In the lowest redshift bin, the ratio for the simulation stabilises to around unity for \llof $<10^{23}\,$\wphz , and decreases rapidly above this, reaching much smaller values than the observations.
+Looking first at the SF population (left set of panels), in the lowest redshift bin the ratio for the simulation stabilises to around unity for \llof $<10^{23}\,$\wphz , and decreases rapidly above this, reaching much smaller values than the observations.
 Higher redshift bins probe successively higher luminosity ranges, also reaching small values for the ratio between activity and galaxies.
-This is consistent with the fact that at higher redshifts, most radio AGN in SIMBA are hosted in SFGs, and thus the \textit{galaxy} RLF for SIMBA SFGs is inflated with AGN emission.
+This is consistent with the fact that at higher redshifts, most radio AGN in SIMBA are hosted in SFGs, and thus the \textit{galaxy} RLF for SIMBA SFGs is inflated with AGN emission; this makes the ratio of SF activity to galaxy approach zero.
 
 For the AGN population (Fig~\ref{fig:evolution}, right set of panels), the ratio between observed activity and galaxy RLFs converges to unity above $\sim$\llof $10^{24}\,$\wphz\ for the lowest redshift bin.
-This convergence happens at higher luminosity values for higher redshifts.
+This is driven by the fact that the flux density of radio excess sources is entirely attributed to AGN, and these sources tend to have higher radio luminosities.
+The convergence happens at higher luminosity values for higher redshifts.
 Below these luminosities, across all redshift bins, the ratio increases.
 The maximum value for each redshift bin approaches $\sim$2, although the fact that in the lowest redshift bin the value stabilises at $\sim1.3$ and only climbs higher at the very lowest luminosities (\llof $<10^{22}\,$\wphz ) may be due to incompleteness.
-If we disregard the lowest luminosity bins, it seems that we see a 30 - 50 percent increase in AGN activity over what is measured when using galaxy classifications alone, for all redshifts.
+If we disregard the lowest luminosity bins, it seems that we see approximately a 30 - 50 percent increase in AGN activity over what is measured when using galaxy classifications alone, for all redshifts.
 The ratio for the SIMBA RLFs shows strong redshift evolution, and much higher ratio values.

### 2024-08-23 src/scripts/RLF_evolution.py
only one plot

@@ -26,6 +26,7 @@ zmax = 0.3   ## matches Mauch & Sadler 2007
 dz = 0.0001  ## matches Cochrane and Kondapally
 
 
+## read in the data 
 t = Table.read( paths.static / 'redshift_bins.csv', format='csv' )
 zbin_starts = t['zbin_starts']
 zbin_ends = t['zbin_ends']
@@ -48,18 +49,17 @@ for i in np.arange(0,5):
         tmp = Table.read( paths.static / infile, format='fits' )
         vmaxes = vstack([vmaxes,tmp[keep_cols]],metadata_conflicts='silent')
     lum_bins, lum_func, agn_lum_func, sf_lum_func, gal_agn_lum_func, gal_sf_lum_func = get_RLFs( vmaxes, zmin, zmax, lmin=20.5, lmax=27, dl=0.3, si=si )
-    ## filter out unconstrained bins
+    ## filter out unconstrained bins - AGN
     filter_idx = np.where( np.abs(agn_lum_func - lum_func) > 4. )[0]
     agn_lum_func[filter_idx] = 0. 
     filter_idx = np.where( np.abs(gal_agn_lum_func - lum_func) > 4. )[0]
     gal_agn_lum_func[filter_idx] = 0. 
-
+    ## filter out unconstrained bins - SF
     filter_idx = np.where( np.logical_or( np.abs(sf_lum_func - lum_func) > 4., sf_lum_func > -1.8 ) )[0]
     sf_lum_func[filter_idx] = 0. 
     filter_idx = np.where( np.logical_or( np.abs(gal_sf_lum_func - lum_func) > 4., gal_sf_lum_func > -1.8 ) )[0]
     gal_sf_lum_func[filter_idx] = 0. 
-
-
+    ## add to redshift list
     z_lum_bins.append(lum_bins)
     z_lum_func.append(lum_func)
     z_agn_lum_func.append(agn_lum_func)
@@ -67,150 +67,34 @@ for i in np.arange(0,5):
     z_gal_agn_lum_func.append(gal_agn_lum_func)
     z_gal_sf_lum_func.append(gal_sf_lum_func)
 
-
+## get the luminosity bin centres
[…]

### 2024-08-23 src/scripts/flowchart_numbers.py
calc numbers

@@ -1,0 +1,47 @@
+import paths
+import numpy as np
+import matplotlib
+from matplotlib import pyplot as plt
+from astropy.table import Table, Column, join
+from helper_functions import *
+from astropy.io import fits
+import os
+## cosmology to match Kondapally and Cochrane
+from astropy.cosmology import FlatLambdaCDM
+cosmo = FlatLambdaCDM(H0=70, Om0=0.3)
+import time
+
+#########################################
+## default parameters to adjust
+si = -0.7
+#zmin = 0.003  ## matches Mauch & Sadler 2007
+#zmax = 0.3   ## matches Mauch & Sadler 2007
+dz = 0.0001  ## matches Cochrane and Kondapally
+## for T_b
+T_e = 1e4
+ref_freqs = np.array([0.003,0.01,0.03,0.1,0.3,1,3])*1e9
+freqs_GHz = np.arange( 1e-3, 1e2, 1e-3 )
+
+sigma_cut = 5.
+
+
+fields = ['lockman', 'en1']
+
+for field in fields:
+    print('Starting with field: {:s}'.format(field))
+    infile = paths.static / '{:s}_03_matched_inMOC_inHR.fits'.format(field)
+    field = os.path.basename(infile).split('_')[0]
+    rms_image = paths.static / '{:s}_DR1_rms_masked.fits'.format(field)
+    lotss = Table.read( infile, format='fits' )
+    ## add brightness temperature information
+    lotss = get_tb_information( lotss, im_weight=0.5, maj_lim=0.4, min_lim=0.3, T_e=T_e, alpha=si, ref_freqs=ref_freqs, freqs_GHz=freqs_GHz, use_z=False )
+    lotss = do_SFR_AGN_separation( lotss )
+    ## get number of detectable sources
+    ndet = len( np.where(lotss['Detectability_SNR'] > 5.)[0] )
+    with open( paths.output / '{:s}_detectable.txt'.format(field), 'w' ) as f:
+        f.write( str(ndet) )
+
+
+
+
+

### 2024-08-23 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Galaxies build up their stellar mass through star formation, while super-massive black holes are expected to provide either negative or positive feedback to either suppress or stimulate galaxy growth, although whether we can definitively pinpoint which type of feedback is an open question \citep{ward_cosmological_2022}.
+If one wishes to study the cosmic history of AGN activity or star formation (SF), and by extension the interplay between them, it is necessary to quantify the contribution of each process separately.
+Relying on overall galaxy classification can over- or under-estimate the contributions of each process; for example, by considering all of the radio emission in a galaxy classified as `star-forming' to be due to star formation.
+Quantifying this is the goal of this paper.
 
-However, if one wishes to study the cosmic history of AGN activity or star formation (SF), and by extension the interplay between them, it is necessary to measure each process separately.
-This is a difficult problem at any wavelength.
-For example \hl{mid-infrared}.
-Multi-wavelength approaches use spectral energy distribution (SED) fitting, which utilises information from a broad number of observed bands to jointly fit galaxy and AGN components \citep{pacifici_art_2022}.
-\hl{give a couple of examples} \citep{calistro_rivera_agnfitter_2016,boquien_cigale_2019}.
+Decomposing AGN activity from star formation is a difficult problem at any wavelength.
+Efforts have been made to do this via mid-infrared (MIR) spectral decomposition using template fitting \citep[e.g.,][]{laurent_mid-infrared_2000,hernan-caballero_resolving_2015,li_active_2024}.
+Multi-wavelength approaches use spectral energy distribution (SED) fitting, which utilises information from a much broader number of observed bands to jointly fit galaxy and AGN components \citep[e.g.][]{calistro_rivera_agnfitter_2016,boquien_cigale_2019,pacifici_art_2022}.
 However, SED fitting is costly as it relies on observations using many different instruments over the same part of the sky, which then have to be carefully cross-matched.
-Selection effects at different bands can also present a problem.
-
-%% suggestion from Ryan: MIR community - AGN fraction
+Varying selection effects across the different bands can also present a problem.
 
@@
 Brightness temperature is inversely proportional to frequency and resolution, $T_b \propto \nu^{-2}\Omega^{-1}$ (where $\Omega \propto \theta_{maj}\theta_{min}$).
-Traditionally brightness temperature measurements require Very Long Baseline Interferometry (VLBI) which achieve milli-arcsec resolutions at GHz frequencies \hl{citations} to achieve enough sensitivity to make meaningful measurements.
-In \cite[][hereafter, \hl{M22}]{morabito_identifying_2022} we demonstrated using sub-arcsecond resolution (0.\sarc\ 3) at 144 MHz with the LOw Frequency ARray \citep[LOFAR;][]{van_haarlem_lofar:_2013}.
-The advantage of LOFAR is its wide field of view: $\sim$6 deg$^2$ from a single observation versus a few to a couple hundred arcmin$^2$ with VLBI at GHz frequencies, depending on whether one is imaging at the phase centre of the observation or using multi-phase centres for correlation \hl{citations}.
+Traditionally brightness temperature measurements require Very Long Baseline Interferometry (VLBI) which achieve milli-arcsec resolutions at GHz frequencies \citep[e.g.,][]{middelberg_mosaiced_2013,herrera_ruiz_faint_2017,radcliffe_nowhere_2018} to achieve enough sensitivity to make meaningful measurements.
+In \cite{morabito_identifying_2022} (hereafter, \citetalias{morabito_identifying_2022}) we demonstrated using sub-arcsecond resolution (0.\sarc\ 3) at 144 MHz with the LOw Frequency ARray \citep[LOFAR;][]{van_haarlem_lofar:_2013}.
+The advantage of LOFAR is its wide field of view: $\sim$6 deg$^2$ from a single observation versus a few to a couple hundred arcmin$^2$ with VLBI at GHz frequencies, depending on whether one is imaging at the phase centre of the observation or using multi-phase centres for correlation \citep{deller_difx-2:_2011,morgan_vlbi_2011}.
 For example, using an 8 hour LOFAR observation of the Lockman Hole \citep{sweijen_deep_2022} we identified 940 AGN via brightness temperature \citepalias{morabito_identifying_2022}, providing a sample two orders of magnitude larger than any VLBI sample of AGN identified at GHz frequencies.
@@
 They also demonstrate the flexible resolution which can be achieved with LOFAR, by imaging at 0.\sarc\ 6 and 1.\sarc\ 2 resolution, which are both complementary to the existing 6\sarc\ \ Deep Fields Data Release 1 (DR1).
-Again this has an advantage over VLBI with instruments at GHz frequencies, which require either observing at lower frequency bands (e.g., e-MERLIN) or with other instruments (e.g., EVN, VLBA) to acquire information on the flux density of sources at other scales.
+Again this has an advantage over VLBI with instruments at GHz frequencies, which require either observing at different frequency bands or with other instruments to acquire information on the flux density of sources at other scales.
 As described in \S~5 of \citepalias{morabito_identifying_2022}, for sources which are unresolved in the 0.\sarc\ 3 resolution images and do not exhibit a radio excess over that predicted the star formation rate (SFR) derived from SED fitting \citep[see][for more details]{best_lofar_2023}, we can separate the AGN activity from SF using a combination of the 0.\sarc\ 3 and 6\sarc\ \ resolution images.
 
-The 2,483 sources in Lockman Hole \citep{sweijen_deep_2022} and now the 13,058 sources in ELAIS-N1 \citep{de_jong_into_2024}, represents a massive step towards robust statistical studies where the population can be broken into bins of redshift and stellar mass to investigate AGN activity and SF.\cite{cochrane_lofar_2023} and \cite{kondapally_cosmic_2022} constructed radio luminosity functions (RLFs) from the 6\sarc\ \ resolution imaging that formed the Deep Fields DR1 for the star forming galaxies and radio-excess AGN, respectively.
-Here we are able to reproduce those RLFs and further construct RLFS by activity -- either AGN or SF -- rather than by overall galaxy classification.
-This is the first time such RLFs have been constructed from observations, and provides insight into the non-radio excess population.
-These RLFs by activity also afford us the opportunity for a more direct comparison with cosmological simulations like SIMBA \hl{Nicole, citations} or EAGLE \hl{citations}.
-We can also investigate the cosmic evolution of AGN activity and SF by dividing the sample into redshift bins.
+The 2,483 sources in Lockman Hole \citep{sweijen_deep_2022} and now the 13,058 sources in ELAIS-N1 \citep{de_jong_into_2024}, represents a massive step towards robust statistical studies where the population can be broken into bins of redshift and stellar mass to investigate the AGN and SFG populations in detail.
+\cite{cochrane_lofar_2023} and \cite{kondapally_cosmic_2022} constructed radio luminosity functions (RLFs) from the 6\sarc\ \ resolution imaging that formed the Deep Fields DR1 for the star forming galaxies and radio-excess AGN, respectively.
+Here we construct RLFs, for the first time, by process rather than overall galaxy classification.
+We do this by using a combination of the 0.\sarc\ 3 and 6\sarc\ \ resolution imaging from Lockman Hole and ELAIS-N1.
+We investigate the cosmic evolution of AGN activity and SF by dividing the sample into redshift bins.
 
-In this paper, we first describe the data and methods in \S~\ref{sec:data}. We use a ...
+In this paper, we first describe the data and methods in \S~\ref{sec:data}, followed by an overview of the methods in \S~\ref{sec:methods}.
+Section~\ref{sec:results} describes the results, followed by
+% discussion and 
+conclusions in 
+%\S~\ref{sec:discussion} and
+\S~\ref{sec:conclusions}
+%, respectively
+.
 
 Throughout this paper, we use a $H_0 = 70$ km$\,$s$^{-1}\,$Mpc$^{-1}$, $\Omega_M = 0.3$, and $\Omega_{\Lambda}= 0.7$ cosmology.
+Radio flux density is defined as $S\propto \nu^{\alpha}$ where $\nu$ is frequnecy and $\alpha$ is radio spectral index.
 
@@
 The Deep Fields Data Release 1 (DR1) in 2021 included Boötes, Lockman Hole, and ELAI-N1, reaching sensitivities of 30, 23, and 20 \muJybm , using $\sim$80, $\sim$100, and $\sim$160 hours, respectively.
+The resolution of these images is all 6\sarc .
 Alongside these deep radio images and their associated catalogues, \cite{kondapally_lofar_2021} provided a careful compilation of the ancillary deep data at other wavebands (from far-infrared to ultraviolet) and cross-matching to the radio catalogues, followed by photometric redshifts and stellar mass estimates in \cite{duncan_lofar_2021}.
@@
 
-Based on the cross-matched multi-wavelength catalogue \citep{kondapally_lofar_2021}, \cite[][hereafter, B23]{best_lofar_2023} carried out detailed SED fitting for all radio-detected sources across the three deep fields, using the multi-wavelength data.
+Based on the cross-matched multi-wavelength catalogue \cite{kondapally_lofar_2021}, \cite{best_lofar_2023} (hereafter, \citetalias{best_lofar_2023}) carried out detailed SED fitting for all radio-detected sources across the three deep fields, using the multi-wavelength data.
 Full details can be found in their paper, but here we outline the relevant details.
@@
 
-The derived consensus properties used in this paper are \hl{stellar masses} and SFRs.
-We also use the galaxy classifications which arise from the SED fitting.
-\citetalias{best_lofar_2023} use the relationship between radio luminosity and SFR to define sources as either \textit{radio excess} or not.
+The derived properties used in this paper are the consensus SFRs, and the galaxy classifications which arise from the SED fitting.
+\citetalias{best_lofar_2023} use the relationship between radio luminosity and SFR to define sources as either \textit{radio excess} or not, based on the expected relationship between radio luminosity and SFR.
 The AGN components used to fit the multi-wavelength information indicate whether a source is a radiatively efficient AGN (i.e., is there evidence for a disk/torus structure) or not.
@@
 The third class, radiatively efficient + not radio excess, are what are commonly called `radio-quiet' (RQ) AGN, while the first two AGN classes would be considered `radio-loud' (RL).
-We note here that this is an observational classification scheme where RL-AGN have radio emission which is not consistent with SF, but that does not mean that the RQ-AGN do not have an AGN component -- which we will return to in \S~\hl{tbd}.
+Five percent of all sources are Unclassified, as they do not have a radio excess and the fits could not determine between radiatively efficient or inefficient.
 
@@
 The longest baseline (currently Ireland to Latvia) is $\sim$2,000 km, providing unprecedented resolution at MHz frequencies.
-While LoTSS uses only the stations in the Netherlands for its standard data processing, it is becoming stahdard to use the full array to achieve sub-arcsecond resolution \citep[e.g.,][]{morabito_sub-arcsecond_2022} on individual targets.
-We are currently post-processing LoTSS data to include the full international array where available.
-
-Post-processing of LoTSS data to achieve sub-arcsecond resolution ongoing for all Deep Fields, where we image the full field of view available to the international array.
+While LoTSS uses only the stations in the Netherlands for its standard data processing, post-processing of LoTSS data to achieve sub-arcsecond resolution is ongoing both for individual sources with $S>10\,$mJy in the wide-area survey \citep{morabito_sub-arcsecond_2022} and for the entire field of view in the Deep Fields \citep[][Escott et al., in prep; Bondi et al., in prep]{sweijen_deep_2022,de_jong_into_2024}.
+While the full field of view imaging is currently too computationally expensive for the wide area, we are focusing on the Deep Fields.
 Due to the fact that the station layout of international stations is larger than that of stations in the Netherlands, the station beams naturally have a smaller field of view.
-Coupled with bandwidth and time smearing, this limits the field of view effectively to a radius of 1.24 deg, which still provides $\sim$5-6 deg$^2$ of coverage.
+Coupled with bandwidth and time smearing, this limits the field of view effectively to a radius of 1.24 deg, which still provides 6.25 deg$^2$ of coverage\footnote{The imaged field of view is square with 2.5 deg per side.}.
 
@@
 The resulting image has a median rms noise of 34 \muJybm , although this is lower in the centre and higher in the outskirts.
-It is useful to note that this image is deeper than an image at 6\sarc\ \ resolution using the exact same 8 hours of data, as adding the international stations increases the collecting area and therefore the sensitivity of the array.
 From this image, a catalogue of 2,316 5$\sigma$ sources was created, removing duplicates and grouping individual Gaussian islands together into sources.
 
-The first 0.\sarc\ 3 image of ELAIS-N1 has recently been published \citep{de_jong_into_2024}, along with companion images at intermediate resolutions of 0.\sarc\ 6 and 1.2\sarc .
+The first 0.\sarc\ 3 image of ELAIS-N1 was recently published \citep{de_jong_into_2024}, along with companion images at intermediate resolutions of 0.\sarc\ 6 and 1.2\sarc .
 The median rms noise reaches 14 \muJybm\ in the centre of the 0.\sarc\ 3 resolution image, and increases radially.
@@
 
-Thanks to the added collecting area of the international stations, the 0.\sarc\ 3 resolution image made with 32 hours of data is now \textit{deeper} than the 6\sarc\ \ resolution image made with $\sim$160 hours of data.
-The derived galaxy properties from \citetalias{best_lofar_2023} (see \S~\ref{subsec:dr1}) were for LoTSS Deep Fields DR1, and thus we have to limit the source catalogue we wish to use.
-
-For both Lockman Hole and ELAIS-N1 catalogues derived from the 0.\sarc\ 3 resolution images, we first find the local rms in the 6\sarc\ \ resolution noise map\footnote{\label{dr1}Available at \href{https://lofar-surveys.org/deepfields.html}{https://lofar-surveys.org/deepfields.html}} at the position of each source.
+Thanks to the added collecting area of the international stations, which increases the sensitivity of the array, the 0.\sarc\ 3 resolution images are \textit{deeper} than the 6\sarc\ \ resolution images made with similar or larger amounts of data.
+The derived galaxy properties from \citetalias{best_lofar_2023} (see \S~\ref{subsec:dr1}) were for LoTSS Deep Fields DR1, and thus we have to limit the source catalogues we wish to use.
+For both Lockman Hole and ELAIS-N1 catalogues derived from the 0.\sarc\ 3 resolution images, we first find the local rms in the companion 6\sarc\ \ resolution noise maps\footnote{\label{dr1}Available at \href{https://lofar-surveys.org/deepfields.html}{https://lofar-surveys.org/deepfields.html}} at the position of each source.
 Using the peak brightness of the source in the catalogue from the 0.\sarc\ 3 resolution image, we calculate the peak-to-noise ratio and only select sources which would be detectable at the 5$\sigma$ limit.
@@
 Finally, we trim the area down to the 2.5 deg$\times$2.5 deg area which is covered by the 0.\sarc\ 3 resolution image.
+We are left with \variable{output/lockman_detectable.txt} sources in Lockman Hole, and \variable{output/en1_detectable.txt} sources in ELAIS-N1.
+We use the source name from the catalogue derived from the 6\sarc\ \ resolution to cross-match to the catalogue from \citetalias{best_lofar_2023}.
 
-Finally, we determine whether sources are resolved or unresolved in the same manner as \S~3.1 of \cite{shimwell_lofar_2019,shimwell_lofar_2022}.
-Truly unresolved sources in radio maps should have a peak brightness to integrated flux density ratio of unity, but LOFAR maps are impacted by smearing due to the ionosphere.
-This causes resolved sources to have a lower peak brightness to integrated flux density ratio, and they can appear slightly extended in the radio images.
-We start by selecting sources which are fit by single Guassian ({\ttfamily S\_CODE = 'S'}), and make a cut to remove sources which are larger than a multiple of the beam size.
+We determine whether sources are resolved or unresolved in the same manner as \S~3.1 of \cite{shimwell_lofar_2019,shimwell_lofar_2022}.
+Truly unresolved sources in radio maps should have equal peak brightness and integrated flux density, but LOFAR maps are impacted by smearing due to the ionosphere.
+This causes resolved sources to have a lower peak brightness when compared to the integrated flux density, and they can appear spatially extended.
+We start by selecting sources which are fit by a single Guassian component ({\ttfamily S\_CODE = `S'}), and make a cut to remove sources which are larger than a multiple of the beam size.
 For the Lockman Hole, we use a factor of 3, but for ELAIS-N1 we find that a factor of 4 provided a better sample.
 This may be because half of the data for ELAIS-N1 had been pre-averaged to 2 sec integration times, which will increase the time smearing \citep[see \S~2 in][]{de_jong_into_2024}.
-We fit a sigmoid function to find the upper 99.9th percentile of sources considering their integrated flux density to peak brightness ratio as a function of the local signal to noise.
-\hl{The fit parameters and graphical representation of the data can be found in the supplemental online materials.}
+We fit a Sigmoid function to find the upper 99.9th percentile of sources considering their integrated flux density to peak brightness ratio as a function of the local signal to noise.
+Anything above this boundary is considered to be resolved.
+We also treat sources which do not have {\ttfamily S\_CODE = `S'} as resolved.
 
-We are left with \hl{fill in} sources in Lockman Hole, and \hl{fill in} sources in ELAIS-N1.
-We use the source name from the catalogue derived from the 6\sarc\ \ resolution to cross-match to the catalogue from \citetalias{best_lofar_2023}.
-
+\begin{comment}
 \subsection{Simulations}
@@
 We can additionally separate the galaxy population into star forming and quiescent galaxies, radio-excess, or radiatively efficient and inefficient, based on the sources black hole and host properties.
+\end{comment}
 
 \section{Methods}
+\label{sec:methods}
 
@@
 Once the AGN activity is identified from the 0.\sarc\ 3 resolution images, it is possible to separate it from star formation using the total flux density from the 6\sarc\ \ resolution images.
-We refer the reader to \S~5 in \citetalias{morabito_identifying_2022} for more details but outline the process here.
+We refer the reader to \S~5 in \citetalias{morabito_identifying_2022} for more details but outline the process here, which is depicted in Fig.~\ref{fig:flowchart}.
 Sources which are identified as radio excess according to the definition in \citetalias{best_lofar_2023} automatically have their total flux density from the 6\sarc\ resolution images added to the AGN category.
-For sources that are not radio excess, we attribute the total flux density from the 6\sarc\ resolution images to SF if in the 0.\sarc\ \ 3 resolution image they are: resolved, multi-component, or have no detection.
-Sources which do not meet these criteria (i.e., unresolved single component) are checked and if their $T_b$ is not above the limit for SF their 6\sarc\ resolution total flux density is also added to the SF category.
-The total flux density of any source in the 0.\sarc\ 3 resolution image which is $T_b$-identified AGN is assumed to be AGN flux density; this is subtracted from the total flux density measured from the 6\sarc\ \ resolution image, and what remains is assumed to be flux density from SF.
-Fig.~\ref{fig:flowchart} shows a flowchart of how flux density is assigned to either the AGN or SF category.
+For sources that are not radio excess, we attribute the total flux density from the 6\sarc\ resolution images to SF if in the 0.\sarc\ 3 resolution image they are either resolved or have no detection.
+Sources which do not meet these criteria are checked and if their $T_b$ is not above the limit for SF their 6\sarc\ resolution total flux density is also added to the SF category.
+The total flux density of any source in the 0.\sarc\ 3 resolution image which is $T_b$-identified as an AGN is assumed to be AGN flux density; this is subtracted from the total flux density measured from the 6\sarc\ \ resolution image, and what remains is assumed to be flux density from SF.
+\hl{how many sources per category}
 
@@
     \includegraphics[width=0.5\textwidth]{figures/agn_sf_separation_workflow.png}
-    \caption{Caption}
+    \caption{The workflow for determining how to assign flux density from the 6\sarc\ and 0.\sarc\ 3 resolution images to either the AGN or SF category.
+Only sources with a $T_b$ identified AGN component have their flux density split between the two categories.}
     \label{fig:flowchart}
@@
 \section{Results}
+\label{sec:results}
 
@@
     \centering
-    \includegraphics[width=0.49\textwidth]{figures/RLF_evolution_SF.png}
-    \includegraphics[width=0.49\textwidth]{figures/RLF_evolution_AGN.png}
+    \includegraphics[width=\textwidth]{figures/RLF_evolution.png}
     \caption{Caption}
     \label{fig:evolution}
-    \script{RLF_evolution_v2.py}
+    \script{RLF_evolution.py}
 \end{figure*}
@@
 \section{Discussion}
+\label{sec:discussion}
 
 \section{Conclusions}
+\label{sec:conclusions}
 

### 2024-08-23 src/scripts/flowchart_numbers.py
calc vars

@@ -40,8 +40,3 @@ for field in fields:
     ndet = len( np.where(lotss['Detectability_SNR'] > 5.)[0] )
     with open( paths.output / '{:s}_detectable.txt'.format(field), 'w' ) as f:
         f.write( str(ndet) )
-
-
-
-
-

### 2024-08-23 src/scripts/flowchart_numbers.py
flowchart numbers

@@ -2,7 +2,7 @@ import paths
 import numpy as np
 import matplotlib
 from matplotlib import pyplot as plt
-from astropy.table import Table, Column, join
+from astropy.table import Table, Column, join, vstack
 from helper_functions import *
 from astropy.io import fits
 import os
@@ -27,6 +27,8 @@ sigma_cut = 5.
 
 fields = ['lockman', 'en1']
 
+mycat = Table()
+
 for field in fields:
     print('Starting with field: {:s}'.format(field))
     infile = paths.static / '{:s}_03_matched_inMOC_inHR.fits'.format(field)
@@ -39,4 +41,33 @@ for field in fields:
     ## get number of detectable sources
     ndet = len( np.where(lotss['Detectability_SNR'] > 5.)[0] )
     with open( paths.output / '{:s}_detectable.txt'.format(field), 'w' ) as f:
-        f.write( str(ndet) )
+        f.write( "{:,}".format(ndet) )
+    mycat = vstack([mycat,lotss],metadata_conflicts='silent')
+
+
+## first get radio detected
+detected = np.where( mycat['Detectability_SNR'] > 5. )[0]
+mycat = mycat[detected]
+
+## radio excess:
+n_radio_excess = len( np.where( mycat['Radio_excess'] >= 0.7 )[0] )
+
+## not radio excess
+n_not_excess = len( np.where( mycat['Radio_excess'] < 0.7 )[0] )
+
+## number not excess & resolved
+n_unres = len( np.where( np.logical_and( mycat['Radio_excess'] < 0.7, mycat['Resolved'] == 'U' ) )[0] )
+
+## tb-identified
+tb_idx = np.where( np.logical_and( mycat['tb_from'] > 0.0, mycat['Resolved'] == 'U' ) )[0]
+not_excess_idx = np.where( mycat['Radio_excess'] < 0.7 )[0]
+combined_idx = np.intersect1d(tb_idx,not_excess_idx)
+
+n_tbid = len( combined_idx )
+
+leftover = n_unres - n_tbid
+
+with open( paths.output / 'flowchart_numbers.txt', 'w' ) as f:
+    f.write( 'There are {:,} total sources. {:,} have a radio excess, while {:,} do not. Of those without a radio excess, there are {:,} unresolved sources for which their $T_b$ is checked. Of those, {:,} are $T_b$-identified AGN, while {:,} are not.'.format(len(detected), n_radio_excess, n_not_excess, n_unres, len(combined_idx), leftover) )
+
+

### 2024-08-23 src/scripts/flowchart_numbers.py
udpate text

@@ -68,6 +68,6 @@ n_tbid = len( combined_idx )
 leftover = n_unres - n_tbid
 
 with open( paths.output / 'flowchart_numbers.txt', 'w' ) as f:
-    f.write( 'There are {:,} total sources. {:,} have a radio excess, while {:,} do not. Of those without a radio excess, there are {:,} unresolved sources for which their $T_b$ is checked. Of those, {:,} are $T_b$-identified AGN, while {:,} are not.'.format(len(detected), n_radio_excess, n_not_excess, n_unres, len(combined_idx), leftover) )
+    f.write( 'We start with {:,} total sources. {:,} have a radio excess, while {:,} do not. Of those without a radio excess, there are {:,} unresolved sources for which their $T_b$ is checked. Of those, {:,} are $T_b$-identified AGN, while {:,} are not. In the final sample, there are {:,} sources contributing to the AGN category and {:,} sources contributing to the SF category. This is more than the total number of sources due to the $T_b$-identified AGN which some of their flux density shifted from the SF to the AGN category.'.format(len(detected), n_radio_excess, n_not_excess, n_unres, len(combined_idx), leftover, n_radio_excess+len(combined_idx), n_not_excess) )
 
 

### 2024-08-23 src/scripts/flowchart_numbers.py
tweak text

@@ -68,6 +68,6 @@ n_tbid = len( combined_idx )
 leftover = n_unres - n_tbid
 
 with open( paths.output / 'flowchart_numbers.txt', 'w' ) as f:
-    f.write( 'We start with {:,} total sources. {:,} have a radio excess, while {:,} do not. Of those without a radio excess, there are {:,} unresolved sources for which their $T_b$ is checked. Of those, {:,} are $T_b$-identified AGN, while {:,} are not. In the final sample, there are {:,} sources contributing to the AGN category and {:,} sources contributing to the SF category. This is more than the total number of sources due to the $T_b$-identified AGN which some of their flux density shifted from the SF to the AGN category.'.format(len(detected), n_radio_excess, n_not_excess, n_unres, len(combined_idx), leftover, n_radio_excess+len(combined_idx), n_not_excess) )
+    f.write( 'We start with {:,} total sources. {:,} have a radio excess, while {:,} do not. Of those without a radio excess, there are {:,} unresolved sources for which their $T_b$ is checked. Of those, {:,} are $T_b$-identified AGN, while {:,} are not. In the final sample, there are {:,} sources contributing to the AGN category and {:,} sources contributing to the SF category. This is more than the total number of sources due to the $T_b$-identified AGN which had a portion of their flux density shifted from the SF to the AGN category.'.format(len(detected), n_radio_excess, n_not_excess, n_unres, len(combined_idx), leftover, n_radio_excess+len(combined_idx), n_not_excess) )
 
 

### 2024-08-23 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Finally, we trim the area down to the 2.5 deg$\times$2.5 deg area which is covered by the 0.\sarc\ 3 resolution image.
-We are left with \variable{output/lockman_detectable.txt} sources in Lockman Hole, and \variable{output/en1_detectable.txt} sources in ELAIS-N1.
+We are left with \variable{output/lockman_detectable.txt}sources in Lockman Hole, and \variable{output/en1_detectable.txt}sources in ELAIS-N1.
 We use the source name from the catalogue derived from the 6\sarc\ \ resolution to cross-match to the catalogue from \citetalias{best_lofar_2023}.
@@
 The total flux density of any source in the 0.\sarc\ 3 resolution image which is $T_b$-identified as an AGN is assumed to be AGN flux density; this is subtracted from the total flux density measured from the 6\sarc\ \ resolution image, and what remains is assumed to be flux density from SF.
-\hl{how many sources per category}
+\variable{output/flowchart_numbers.txt}
 
@@
 We follow the same method of constructing RLFs as in \cite{kondapally_cosmic_2022} and \cite{cochrane_lofar_2023}.
-As the 0.\sarc\ \ 3 resolution image is smaller than the 6\sarc\ resolution image, we first reproduce the published RLFs (Fig.~\ref{fig:rlfs}, left panel) using the galaxy classifications from \citetalias{best_lofar_2023} to check that we have implemented the method appropriately, and that the smaller area yields the same results.
+The 0.\sarc\ 3 resolution image is smaller than the 6\sarc\ resolution image, so we first reproduce the published RLFs (Fig.~\ref{fig:rlfs}, left panel) using the galaxy classifications from \citetalias{best_lofar_2023} to check that we have implemented the method appropriately, and that the smaller area yields the same results.
 We use the completeness corrections from the respective papers.

### 2024-08-23 src/scripts/RLF_evolution.py
add table

@@ -89,26 +89,20 @@ zcols_agn = mycols_m[np.arange(0,len(z_lum_bins))*int(n/len(z_lum_bins))]
 ## start the figure
 fig = plt.figure( figsize=(fsizex,fsizey) )
 
+print('SF')
+sf_delta_int = []
+
 ## STAR FORMATION
 p1 = plt.axes([0.07,0.42,sbsizex*fsizey/fsizex,0.6*sbsizey])
 for i in np.arange(0,len(z_lum_bins)):
     non_zero = np.where( z_gal_sf_lum_func[i] < 0 )[0]
     p1.plot( lum_bin_cens[non_zero], z_gal_sf_lum_func[i][non_zero], color=zcols_sf[i], linewidth=3, alpha=0.75, linestyle='dotted' )
+    gal_trapz = np.trapz( np.power( 10., z_gal_sf_lum_func[i][non_zero] ), lum_bin_cens[non_zero] )
     non_zero = np.where( z_sf_lum_func[i] < 0  )[0]
     p1.plot( lum_bin_cens[non_zero], z_sf_lum_func[i][non_zero], color=zcols_sf[i], label='{:s} < z < {:s}'.format(str(zbin_starts[i]),str(zbin_ends[i])), linewidth=3 )
-
-
-    ## integrate under galaxies to find normalisation ?
+    trapz = np.trapz( np.power( 10., z_sf_lum_func[i][non_zero] ), lum_bin_cens[non_zero] )
+    sf_delta_int.append(trapz / gal_trapz)
     
-
-    tmp = np.trapz( z_gal_sf_lum_func[i][non_zero], lum_bin_cens[non_zero] )
-    corrfac = 1. / tmp 
-
-    test = z_gal_sf_lum_func[i][non_zero] * corrfac
-    test1 = z_sf_lum_func[i][non_zero] * corrfac
-    ratio = test / test1 
-    fred = np.trapz( ratio, lum_bin_cens[non_zero] )
-    #print(fred)
 
 l1 = p1.legend()
 dline = matplotlib.lines.Line2D([0],[0], color='black', linewidth=3 )
@@ -129,12 +123,7 @@ p2.plot( (19,27), (0.5,0.5), color='gray', linewidth=1, alpha=0.25 )
 for i in np.arange(0,len(z_lum_bins)):
[…]

### 2024-08-23 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Relying on overall galaxy classification can over- or under-estimate the contributions of each process; for example, by considering all of the radio emission in a galaxy classified as `star-forming' to be due to star formation.
-Quantifying this is the goal of this paper.
+Quantifying this is the goal of this letter.
 
@@
 Radio emission has the advantage of being unobscured by dust or gas, meaning there are minimal selection effects.
-Brightness temperature is a non-physical temperature which is defined as the temperature of a blackbody which would produce the observed surface brightness.
+%Brightness temperature is a non-physical temperature which is defined as the temperature of a blackbody which would produce the observed surface brightness.
 Models of radio emission from SF predict an upper limit to the surface brightness that can be produced, even in starburst galaxies \citep{condon_radio_1992}.
@@
 The authors note that this correction is negligible, and as it is not implemented in \cite{kondapally_cosmic_2022} we chose to keep the method for RLF construction in this paper consistent across both samples.
-We use the same redshift range, $0.003 < z < 0.3$ and see that we indeed have good agreement with the published RLFs.
+We use the same redshift range, $0.003 < z < 0.3$, to verify that we have good agreement with the published RLFs.
 
-Next we use the same method for constructing RLFs by activity rather than overall galaxy classifcation.
+Next we use the same method for constructing RLFs by physical process rather than overall galaxy classification.
 We again use the completeness corrections from \cite{cochrane_lofar_2023} and \cite{kondapally_cosmic_2022}.
-The completeness corrections are flux density dependent, and we use the overall total flux density of a galaxy, rather than the flux density of the activity (which may be smaller if there are both AGN and SF components).
-This is because completeness corrections are based on whether or not a \textit{source} will be detectable given the background rms.
+The completeness corrections are flux density dependent, and we use the overall total flux density of a galaxy, rather than the flux density of the process, which will be smaller if the radio emission from SF and AGN have been split.
+We still use the overall galaxy completeness correction because this is based on whether or not a \textit{source} will be detectable given the background rms.
 This only relies on the total flux density, not the flux density of individual components.
-The RLFs by activity rather than galaxy classification are shown in the middle panel of Fig.~\ref{fig:rlfs}, again for  $0.003 < z < 0.3$.
+Even when considering only the SF (or AGN) radio emission in a source, its overall detectability is still derived from the combination of all radio emission processes present.
+These RLFs for the physical processes of SF and AGN are shown in the middle panel of Fig.~\ref{fig:rlfs}, again for  $0.003 < z < 0.3$.
 
-Finally, we use RLFs constructed from SIMBA. The RLFs are constructed as in \cite{thomson_radio_2019}, but have been updated to use the same redshift bins \hl{Nicole}.
+% Finally, we use RLFs constructed from SIMBA. The RLFs are constructed as in \cite{thomson_radio_2019}, but have been updated to use the same redshift bins \hl{Nicole}.
 
@@
 
-To illustrate the differences in RLFs constructed by activity vs.
-by galaxy classification we divide the activity RLF by the galaxy RLF for the observations and show the ratio in the bottom right hand panel in Fig.~\ref{fig:rlfs}, for both the AGN and SF category.
+We present the very first RLFs constructed by physical process rather than galaxy classification.
+Comparing these with our reproduction of the previously published RLFs, using galaxy classifications, allows us to explore where we are over- or under-predicting the contribution of the physical processes to our statistical studies.
+To aid this comparison, we define and calculate:
+\begin{equation}
+\Delta\textrm{RLF} = \frac{\textrm{RLF(physical process)}}{\textrm{RLF(galaxy classification)}}. \notag
+\end{equation}
+The $\Delta$RLF is calculated once for the SF category/SFG classification, and once for the AGN category/AGN classification.
+We show $\Delta$RLF in the bottom right hand panel in Fig.~\ref{fig:rlfs}.
 We see that above 10$^{24}\,$\wphz\ the ratio for AGN is unity, which is expected where radio excess sources dominate the population.
-Below that, the ratio for AGN climbs up to $\sim$1.25 and finally almost up to 2 at the lowest radio powers probed.
+Below that, the ratio for AGN climbs up to $\sim$1.3 and finally almost up to 2 at the lowest radio powers probed.
 The ratio for SF is always below unity, but it does not deviate as much as the ratio for AGN.
@@
     \includegraphics[width=0.98\textwidth]{figures/deep_fields_RLFs.png}
-    \caption{Caption}
+    \caption{\textit{Left:} Validation of our re-calculated galaxy RLFs (dotted lines) for the smaller area considered here with the previously published RLFs (solid lines).
+\textit{Middle:} RLFs calculated by process rather than galaxy.
+\textit{Right top:} RLFs calculated here by galaxy classifications (dotted lines) and by physical process (solid lines; orange for AGN and green for SF).
+\textit{Bottom right:} $\Delta$RLF for both AGN and star formation.
+}
     \label{fig:rlfs}
@@
 \end{figure*}
+
+To show the integrated effect of $\Delta$RLF, we integrate each RLF separately across the luminosity bins where both RLFs are defined, and then take the ratio of the areas.
+For example, for star formation we integrate the RLF for the SF process and the RLF for SFGs, and divide the first by the second.
+These numbers are always less than unity since the SF process is over-estimated in the SFG population, and this has been corrected by moving $T_b$-identified AGN flux density contributions to the AGN process RLF.
 

### 2024-08-23 src/scripts/RLF_evolution.py
format numbers

@@ -175,12 +175,12 @@ plt.close()
 
 ## write a table
 with open( paths.output / 'integrated_differences.txt', 'w' ) as f:
-    f.write( '\\begin{table}[]\n' )
+    f.write( '\\begin{table}\n' )
     f.write( '    \\centering\n' )
     f.write( '    \\begin{tabular}{c|l|l}\n' )
     f.write( '  & SF & AGN \\\\ \\hline\n' )
     for i in np.arange(0,len(z_lum_bins)):
-        f.write( '{:s} < z < {:s} & {:2f}  & {:2f}  \\\\ \n'.format(str(zbin_starts[i]),str(zbin_ends[i]),sf_delta_int[i],agn_delta_int[i]) )
+        f.write( '{:s} \< z \< {:s} & {:1.2f}  & {:1.2f}  \\\\ \n'.format(str(zbin_starts[i]),str(zbin_ends[i]),sf_delta_int[i],agn_delta_int[i]) )
     f.write( '    \\end{tabular}\n' )
     f.write( '    \\caption{Integrated contribution of each process, calculated as the ratio of areas under the RLF curve by process to the RLF curve by galaxy classification.}\n' )
     f.write( '    \\label{tab:intvalues}\n' )

### 2024-08-23 src/scripts/RLF_evolution.py
format le signs

@@ -180,7 +180,7 @@ with open( paths.output / 'integrated_differences.txt', 'w' ) as f:
     f.write( '    \\begin{tabular}{c|l|l}\n' )
     f.write( '  & SF & AGN \\\\ \\hline\n' )
     for i in np.arange(0,len(z_lum_bins)):
-        f.write( '{:s} \< z \< {:s} & {:1.2f}  & {:1.2f}  \\\\ \n'.format(str(zbin_starts[i]),str(zbin_ends[i]),sf_delta_int[i],agn_delta_int[i]) )
+        f.write( '{:s} $<$ z $<$ {:s} & {:1.2f}  & {:1.2f}  \\\\ \n'.format(str(zbin_starts[i]),str(zbin_ends[i]),sf_delta_int[i],agn_delta_int[i]) )
     f.write( '    \\end{tabular}\n' )
     f.write( '    \\caption{Integrated contribution of each process, calculated as the ratio of areas under the RLF curve by process to the RLF curve by galaxy classification.}\n' )
     f.write( '    \\label{tab:intvalues}\n' )

### 2024-08-23 src/scripts/RLF_evolution.py
redshift title

@@ -178,7 +178,7 @@ with open( paths.output / 'integrated_differences.txt', 'w' ) as f:
     f.write( '\\begin{table}\n' )
     f.write( '    \\centering\n' )
     f.write( '    \\begin{tabular}{c|l|l}\n' )
-    f.write( '  & SF & AGN \\\\ \\hline\n' )
+    f.write( ' redshift & SF & AGN \\\\ \\hline\n' )
     for i in np.arange(0,len(z_lum_bins)):
         f.write( '{:s} $<$ z $<$ {:s} & {:1.2f}  & {:1.2f}  \\\\ \n'.format(str(zbin_starts[i]),str(zbin_ends[i]),sf_delta_int[i],agn_delta_int[i]) )
     f.write( '    \\end{tabular}\n' )

### 2024-08-23 src/scripts/RLF_evolution.py
portrait to landscape

@@ -177,6 +177,30 @@ plt.close()
 with open( paths.output / 'integrated_differences.txt', 'w' ) as f:
     f.write( '\\begin{table}\n' )
     f.write( '    \\centering\n' )
+    f.write( '    \\begin{tabular}{lccccc}\n' )
+    zstr = ' & '
+    for i in np.arange(0,len(z_lum_bins)):
+        zstr = zstr + '{:s} $<$ z $<$ {:s} & '.format(str(zbin_starts[i]),str(zbin_ends[i]))
+    zstr = zstr + '\\\\ \hline \n'
+    f.write( zstr ) 
+    sfstr = 'SF & ' 
+    for i in np.arange(0,len(sf_delta_int)):
+        sfstr = sfstr + ' & {:1.2f}'.format(sf_delta_int[i])
+    sfstr = sfstr + '\\\\ \n' )
+    sfstr = 'AGN & ' 
+    for i in np.arange(0,len(agn_delta_int)):
+        sfstr = sfstr + ' & {:1.2f}'.format(agn_delta_int[i])
+    sfstr = sfstr + '\\\\ \n' )
+    f.write( '    \\end{tabular}\n' )
+    f.write( '    \\caption{Integrated contribution of each process, calculated as the ratio of areas under the RLF curve by process to the RLF curve by galaxy classification.}\n' )
+    f.write( '    \\label{tab:intvalues}\n' )
+    f.write( '\\end{table}' )
+
+
+'''
+with open( paths.output / 'integrated_differences.txt', 'w' ) as f:
+    f.write( '\\begin{table}\n' )
+    f.write( '    \\centering\n' )
     f.write( '    \\begin{tabular}{c|l|l}\n' )
     f.write( ' redshift & SF & AGN \\\\ \\hline\n' )
     for i in np.arange(0,len(z_lum_bins)):
@@ -185,6 +209,11 @@ with open( paths.output / 'integrated_differences.txt', 'w' ) as f:
     f.write( '    \\caption{Integrated contribution of each process, calculated as the ratio of areas under the RLF curve by process to the RLF curve by galaxy classification.}\n' )
     f.write( '    \\label{tab:intvalues}\n' )
     f.write( '\\end{table}' )
+
+'''
+
+
+
 
 '''
 Redshift bins

### 2024-08-23 src/scripts/RLF_evolution.py
typos

@@ -186,11 +186,13 @@ with open( paths.output / 'integrated_differences.txt', 'w' ) as f:
     sfstr = 'SF & ' 
     for i in np.arange(0,len(sf_delta_int)):
         sfstr = sfstr + ' & {:1.2f}'.format(sf_delta_int[i])
-    sfstr = sfstr + '\\\\ \n' )
-    sfstr = 'AGN & ' 
+    sfstr = sfstr + '\\\\ \n' 
+    f.write(sfstr)
+    agnstr = 'AGN & ' 
     for i in np.arange(0,len(agn_delta_int)):
-        sfstr = sfstr + ' & {:1.2f}'.format(agn_delta_int[i])
-    sfstr = sfstr + '\\\\ \n' )
+        agnstr = agnstr + ' & {:1.2f}'.format(agn_delta_int[i])
+    agnstr = agnstr + '\\\\ \n' 
+    f.write(agnstr)
     f.write( '    \\end{tabular}\n' )
     f.write( '    \\caption{Integrated contribution of each process, calculated as the ratio of areas under the RLF curve by process to the RLF curve by galaxy classification.}\n' )
     f.write( '    \\label{tab:intvalues}\n' )

### 2024-08-23 src/scripts/RLF_evolution.py
andpersands

@@ -178,12 +178,12 @@ with open( paths.output / 'integrated_differences.txt', 'w' ) as f:
     f.write( '\\begin{table}\n' )
     f.write( '    \\centering\n' )
     f.write( '    \\begin{tabular}{lccccc}\n' )
-    zstr = ' & '
+    zstr = ' '
     for i in np.arange(0,len(z_lum_bins)):
-        zstr = zstr + '{:s} $<$ z $<$ {:s} & '.format(str(zbin_starts[i]),str(zbin_ends[i]))
+        zstr = zstr + ' & {:s} $<$ z $<$ {:s} '.format(str(zbin_starts[i]),str(zbin_ends[i]))
     zstr = zstr + '\\\\ \hline \n'
     f.write( zstr ) 
-    sfstr = 'SF & ' 
+    sfstr = 'SF ' 
     for i in np.arange(0,len(sf_delta_int)):
         sfstr = sfstr + ' & {:1.2f}'.format(sf_delta_int[i])
     sfstr = sfstr + '\\\\ \n' 

### 2024-08-23 src/scripts/RLF_evolution.py
spaces

@@ -181,17 +181,17 @@ with open( paths.output / 'integrated_differences.txt', 'w' ) as f:
     zstr = ' '
     for i in np.arange(0,len(z_lum_bins)):
         zstr = zstr + ' & {:s} $<$ z $<$ {:s} '.format(str(zbin_starts[i]),str(zbin_ends[i]))
-    zstr = zstr + '\\\\ \hline \n'
+    zstr = zstr + ' \\\\ \hline \n'
     f.write( zstr ) 
     sfstr = 'SF ' 
     for i in np.arange(0,len(sf_delta_int)):
         sfstr = sfstr + ' & {:1.2f}'.format(sf_delta_int[i])
-    sfstr = sfstr + '\\\\ \n' 
+    sfstr = sfstr + ' \\\\ \n' 
     f.write(sfstr)
-    agnstr = 'AGN & ' 
+    agnstr = 'AGN ' 
     for i in np.arange(0,len(agn_delta_int)):
         agnstr = agnstr + ' & {:1.2f}'.format(agn_delta_int[i])
-    agnstr = agnstr + '\\\\ \n' 
+    agnstr = agnstr + ' \\\\ \n' 
     f.write(agnstr)
     f.write( '    \\end{tabular}\n' )
     f.write( '    \\caption{Integrated contribution of each process, calculated as the ratio of areas under the RLF curve by process to the RLF curve by galaxy classification.}\n' )

### 2024-08-23 src/scripts/RLF_evolution.py
zbins split

@@ -178,11 +178,15 @@ with open( paths.output / 'integrated_differences.txt', 'w' ) as f:
     f.write( '\\begin{table}\n' )
     f.write( '    \\centering\n' )
     f.write( '    \\begin{tabular}{lccccc}\n' )
-    zstr = ' '
+    zmin = ' $z_{min}$ '
+    zmax = ' $z_{max}$ '
     for i in np.arange(0,len(z_lum_bins)):
-        zstr = zstr + ' & {:s} $<$ z $<$ {:s} '.format(str(zbin_starts[i]),str(zbin_ends[i]))
-    zstr = zstr + ' \\\\ \hline \n'
-    f.write( zstr ) 
+        zmin = zmin + ' & {:s}'.format(str(zbin_starts[i]))
+        zmax = zmax + ' & {:s}'.format(,str(zbin_ends[i]))
+    zmin = zmin + ' \\\\ \n'
+    zmax = zmax + ' \\\\ \hline \n'
+    f.write( zmin ) 
+    f.write( zmax )
     sfstr = 'SF ' 
     for i in np.arange(0,len(sf_delta_int)):
         sfstr = sfstr + ' & {:1.2f}'.format(sf_delta_int[i])

### 2024-08-23 src/scripts/RLF_evolution.py
remove comma

@@ -182,7 +182,7 @@ with open( paths.output / 'integrated_differences.txt', 'w' ) as f:
     zmax = ' $z_{max}$ '
     for i in np.arange(0,len(z_lum_bins)):
         zmin = zmin + ' & {:s}'.format(str(zbin_starts[i]))
-        zmax = zmax + ' & {:s}'.format(,str(zbin_ends[i]))
+        zmax = zmax + ' & {:s}'.format(str(zbin_ends[i]))
     zmin = zmin + ' \\\\ \n'
     zmax = zmax + ' \\\\ \hline \n'
     f.write( zmin ) 

### 2024-08-23 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 These numbers are always less than unity since the SF process is over-estimated in the SFG population, and this has been corrected by moving $T_b$-identified AGN flux density contributions to the AGN process RLF.
+These are reported in Tab.~\ref{tab:intvalues}, and show that while the SF process  only decreases by $\sim$10 percent in the lowest redshift bin, the AGN contribution increases by almost a factor of 2.
+It is evident from the bottom right panel of Fig.~\ref{fig:rlfs} that this `hidden' AGN population extracted from SFGs shows up in the radio AGN population at $L_{\textrm{144MHz}} < 10^{24}\,$\wphz .
 
-In Fig.~\ref{fig:evolution} we show a comparison between the observed RLFs and those generated from SIMBA (see \S~\ref{subsec:simba}), and the redshift evolution for both.
+\variable{output/integrated_differences.txt}
+
+In Fig.~\ref{fig:evolution} we show the redshift evolution for star formation (left) and AGN (right).
 We use similar redshift bins to \cite{kondapally_cosmic_2022} and \cite{cochrane_lofar_2023} for consistency.
-The left set of panels shows the results for SF and the right set of panels shows the results for AGN.
-\hl{We cut off the RLFs below a flux limit where they are not constrained}.
-We use the same redshift bins across all samples for consistency.
+Some bins become unconstrained for the RLFs by process due to the shuffling of sources between the SF and AGN populations, yielding positve y-axis values.
+We remove these by cutting out bins where there are more than 4 dex difference between the overall RLF and the process RLF.
+For the SF RLF we additionally remove points $>1.8$.
 We reproduce previous observational results that while the SF population has a strong redshift evolution, the AGN population does not.
-For the simulations, the behaviour is qualitatively similar, although the AGN population does have mild redshift evolution whereas the observations do not.
+%For the simulations, the behaviour is qualitatively similar, although the AGN population does have mild redshift evolution whereas the observations do not.
 
-The bottom panels in Fig~\ref{fig:evolution} show the ratio of activity to galaxy RLFs, the same as in the bottom right hand panel of Fig~\ref{fig:rlfs} (although note that the y-scale is logarithmic for the AGN).
-Comparing these ratios between observations and simulations, we see qualitatively the same behaviour, although the simulations reach much lower and higher values for both the SF and AGN populations, respectively.
+The bottom panels in Fig~\ref{fig:evolution} show $\Delta$RLF, the same as in the bottom right hand panel of Fig~\ref{fig:rlfs}.
+For the SF population, $\Delta$RLF is close to unity for $L_{\textrm{144MHz}} < 10^{24}\,$\wphz in the lowest redshift bin, and then this begins to drop at higher luminosities.
+The higher redshift bins show similar behaviour, perhaps reaching unity at increasingly higher values of $L_{\textrm{144MHz}}$, although deeper data is needed to confirm this.
+The lowest values reached are $\sim$50 percent.
+%Comparing these ratios between observations and simulations, we see qualitatively the same behaviour, although the simulations reach much lower and higher values for both the SF and AGN populations, respectively.
 Looking first at the SF population (left set of panels), in the lowest redshift bin the ratio for the simulation stabilises to around unity for \llof $<10^{23}\,$\wphz , and decreases rapidly above this, reaching much smaller values than the observations.
@@
 
-For the AGN population (Fig~\ref{fig:evolution}, right set of panels), the ratio between observed activity and galaxy RLFs converges to unity above $\sim$\llof $10^{24}\,$\wphz\ for the lowest redshift bin.
+For the AGN population, $\Delta$RLF has almost the opposite behaviour, converging to unity \textit{above} $\sim$\llof $10^{24}\,$\wphz\ for the lowest redshift bin.
 This is driven by the fact that the flux density of radio excess sources is entirely attributed to AGN, and these sources tend to have higher radio luminosities.
 The convergence happens at higher luminosity values for higher redshifts.
-Below these luminosities, across all redshift bins, the ratio increases.
+Below this convergence, across all redshift bins, $\Delta$RLF increases.
 The maximum value for each redshift bin approaches $\sim$2, although the fact that in the lowest redshift bin the value stabilises at $\sim1.3$ and only climbs higher at the very lowest luminosities (\llof $<10^{22}\,$\wphz ) may be due to incompleteness.
 If we disregard the lowest luminosity bins, it seems that we see approximately a 30 - 50 percent increase in AGN activity over what is measured when using galaxy classifications alone, for all redshifts.
-The ratio for the SIMBA RLFs shows strong redshift evolution, and much higher ratio values.
+%The ratio for the SIMBA RLFs shows strong redshift evolution, and much higher ratio values.
 The RLFs themselves turn over, whereas the observed RLFs do not.
@@
 Shifting this AGN emission to AGN \textit{activity} when it is counted in the SFG \textit{galaxy} population generates high values for the activity to galaxy ratio.
-
 %- sfrd, kinetic LF
@@
     \includegraphics[width=\textwidth]{figures/RLF_evolution.png}
-    \caption{Caption}
+    \caption{The redshift evolution of the RLFs (top panels) and $\Delta$RLF (bottom panels) for SF (left) and AGN (right).
+The dashed horizontal line in the bottom panels is unity, with thin solid horizontal lines at 0.5 (left) at 2.0 (right) to guide the eye.
+}
     \label{fig:evolution}
@@
 \end{figure*}
-
-\section{Discussion}
-\label{sec:discussion}
 

### 2024-08-23 src/scripts/comparison_plot.py
delta rlf

@@ -81,7 +81,7 @@ order = [3,2,0,1]
 p1.legend([handles[idx] for idx in order],[labels[idx] for idx in order])
 
 
-## Middle panel: same as left panel but new data by activity now
+## Middle panel: same as left panel but new data by process now
 p2 = plt.axes([0.05+sbsizex*fsizey/fsizex,0.1,sbsizex*fsizey/fsizex,sbsizey])
 ## plot the previous data
 p2.plot( cochrane['logL150'], cochrane['logPhi'], color=coch, label='Cochrane et al. (2023)', linewidth=2)
@@ -90,9 +90,9 @@ p2.plot( kondapally['logL150'], kondapally['logPhi'], color=kond, label='Kondapa
 #non_zero = np.where( lum_func != 0.0 )[0]
 #p2.plot( lum_bin_cens[non_zero], lum_func[non_zero], color='black', label='Total' )
 non_zero = np.where( agn_lum_func != 0.0 )[0]
-p2.plot( lum_bin_cens[non_zero], agn_lum_func[non_zero], color=agnc, label='AGN activity', linewidth=3, alpha=0.75 )
+p2.plot( lum_bin_cens[non_zero], agn_lum_func[non_zero], color=agnc, label='AGN process', linewidth=3, alpha=0.75 )
 non_zero = np.where( sf_lum_func != 0.0 )[0]
-p2.plot( lum_bin_cens[non_zero], sf_lum_func[non_zero], color=sfc, label='SF activity', linewidth=3, alpha=0.75 )
+p2.plot( lum_bin_cens[non_zero], sf_lum_func[non_zero], color=sfc, label='SF process', linewidth=3, alpha=0.75 )
 p2.axes.set_xlim(plxlims)
 p2.axes.set_ylim(plylims)
 p2.yaxis.set_visible(False)
@@ -102,16 +102,16 @@ handles, labels = p2.get_legend_handles_labels()
 order = [3,2,0,1]
 p2.legend([handles[idx] for idx in order],[labels[idx] for idx in order])
 
-## Right panel (top): RLFs using new data, galaxies and activity
+## Right panel (top): RLFs using new data, galaxies and process
 p3 = plt.axes([0.12+2*sbsizex*fsizey/fsizex,0.42,sbsizex*fsizey/fsizex,0.6*sbsizey])
[…]

### 2024-08-23 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Title
-\title[TBD]{A hidden Active Galactic Nuclei population: the first radio luminosity functions by process}
+\title[A hidden AGN population]{A hidden Active Galactic Nuclei population: the first radio luminosity functions constructed by physical process instead of galaxy classification}
 
 % Author list
-\author[L.K. Morabito]{\parbox{\textwidth}{Leah K. Morabito$^{1,2}$\thanks{E-mail: leah.k.morabito@durham.ac.uk}\orcidlink{0000-0003-0487-6651}\\}\\ 
+\author[L.K. Morabito]{\parbox{\textwidth}{Leah K. Morabito$^{1,2}$\thanks{E-mail: leah.k.morabito@durham.ac.uk}\orcidlink{0000-0003-0487-6651}
+J.M.G.H.J. de Jong$^{3}$,
+F. Sweijen$^{1}$,
+R. Kondapally$^{4}$,
+P.N. Best$^{4}$,
+B. Yue$^{4}$, and friends
+\\}\\ 
 $^{1}$Centre for Extragalactic Astronomy, Department of Physics, Durham University, Durham DH1 3LE, UK \\
-$^{2}$Institute for Computational Cosmology, Department of Physics, Durham University, South Road, Durham DH1 3LE, UK \\ }
+$^{2}$Institute for Computational Cosmology, Department of Physics, Durham University, South Road, Durham DH1 3LE, UK \\ 
+$^{3}$Leiden \\
+$^{4}$Edinburgh \\}
 
@@
 \date{}
-\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2021}
+\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2024}
 \maketitle
@@
 \begin{abstract}
-    TBD
-    \vspace{6.75in}
+Both star formation (SF) and Active galactic nuclei (AGN) play an important role in galaxy evolution.
+Quantifying their relative importance in a statistical way can be done using radio luminosity functions.
+Until now these have relied on galaxy classifications, with sources which have a mixture of radio emission from SF and AGN assigned to either a star-forming galaxy or an AGN.
+This can lead to the under-estimation of the relevance of AGN, or an over-estimation of SF.
+Brightness temperature methods were demonstrated to be effective at 144 MHz using sub-arcsecond imaging with the International LOFAR Telescope (ILT) in \cite{morabito_identifying_2022}.
+We use the combination of sub-arcsecond and arcsecond resolution imaging of 7,497 sources in the Lockman Hole and ELAIS-N1 fields to identify AGN components in the high resolution images and subtract them from the total flux density, leaving behind flux density from SF only.
+We construct, for the first time, radio luminosity functions by physical process rather than galaxy classification, revealing a hidden AGN population at \llof $<10^{24}$\wphz .
+This population is $\sim$50 percent more than expected for $0.5<z<2.0$.
+The star forming population is less impacted, with only $\sim$10 - 20 percent less star formation than expected.
+The contribution of `hidden' AGN to the overall radio population, especially at \llof $<10^{24}$\wphz , is significant, and only revealed thanks to the combination of the ILT's resolution and wide field of view.
+\vspace{3.7in}
 \end{abstract}
@@
 Observations have revealed tight scaling relations between the mass of a super-massive black hole and its host galaxy \citep[see, e.g.][and references therein]{kormendy_coevolution_2013}, while cosmological simulations require some form of AGN feedback \citep{bower_breaking_2006,croton_many_2006} to be able to reproduce the observed galaxy population.
-Galaxies build up their stellar mass through star formation, while super-massive black holes are expected to provide either negative or positive feedback to either suppress or stimulate galaxy growth, although whether we can definitively pinpoint which type of feedback is an open question \citep{ward_cosmological_2022}.
-If one wishes to study the cosmic history of AGN activity or star formation (SF), and by extension the interplay between them, it is necessary to quantify the contribution of each process separately.
-Relying on overall galaxy classification can over- or under-estimate the contributions of each process; for example, by considering all of the radio emission in a galaxy classified as `star-forming' to be due to star formation.
+%Galaxies build up their stellar mass through star formation, while super-massive black holes are expected to provide either negative or positive feedback to either suppress or stimulate galaxy growth, although whether we can definitively pinpoint which type of feedback is an open question \citep{ward_cosmological_2022}.
+The cosmic history of AGN activity and star formation (SF), and by extension the interplay between them, are interlinked.
+It is necessary to quantify the contribution of each process to understand the importance of either or both of them.
+Relying on overall galaxy classification can over- or under-estimate the contributions of each process; for example, by considering all of the radio emission in a galaxy classified as `star-forming' to be due to star formation, when some may be due to AGN.
 Quantifying this is the goal of this letter.
@@
 \begin{itemize}
-    \item Radio excess + radiatively efficient
-    \item Radio excess + not radiatively efficient
-    \item Radiatively efficient + not radio excess
+    \item Radio excess + radiatively efficient (\textit{`radio-loud'})
+    \item Radio excess + not radiatively efficient (\textit{`radio-loud'})
+    \item Radiatively efficient + not radio excess (\textit{`radio-quiet'})
     \item Radiatively inefficient + not radio excess
@@
 The first three classes are AGN, and the final class is star forming galaxies (SFGs).
-The third class, radiatively efficient + not radio excess, are what are commonly called `radio-quiet' (RQ) AGN, while the first two AGN classes would be considered `radio-loud' (RL).
 Five percent of all sources are Unclassified, as they do not have a radio excess and the fits could not determine between radiatively efficient or inefficient.
 
-\subsection{0.\sarc\ 3 resolution imaging}
+\subsection{High-resolution LOFAR imaging}
 \label{subsec:highres}
 
-Data for LoTSS is recorded with all available LOFAR stations, which are spread across nine different European countries, with a preponderance of \textit{core} and \textit{remote} stations at the heart of the array in the Netherlands.
+Data for LoTSS is recorded with all available LOFAR stations, which are spread across nine different European countries, with a preponderance of stations in the Netherlands.
 The longest baseline (currently Ireland to Latvia) is $\sim$2,000 km, providing unprecedented resolution at MHz frequencies.
 While LoTSS uses only the stations in the Netherlands for its standard data processing, post-processing of LoTSS data to achieve sub-arcsecond resolution is ongoing both for individual sources with $S>10\,$mJy in the wide-area survey \citep{morabito_sub-arcsecond_2022} and for the entire field of view in the Deep Fields \citep[][Escott et al., in prep; Bondi et al., in prep]{sweijen_deep_2022,de_jong_into_2024}.
-While the full field of view imaging is currently too computationally expensive for the wide area, we are focusing on the Deep Fields.
-Due to the fact that the station layout of international stations is larger than that of stations in the Netherlands, the station beams naturally have a smaller field of view.
+%While the full field of view imaging is currently too computationally expensive for the wide area, we are focusing on the Deep Fields.
+Due to the fact that the station layout of international stations is larger than that of stations in the Netherlands, the station beams have a smaller field of view.
 Coupled with bandwidth and time smearing, this limits the field of view effectively to a radius of 1.24 deg, which still provides 6.25 deg$^2$ of coverage\footnote{The imaged field of view is square with 2.5 deg per side.}.
 
-The Lockman Hole was the first field to have its full field of view imaged at 0.\sarc\ 3 resolution \citep{sweijen_deep_2022}.
+Both Lockman Hole and ELAIS-N1 have been imaged at 0.\sarc\ 3 resolution \citep[][respectively]{sweijen_deep_2022,de_jong_into_2024}.
+The resulting images have median rms noise of 34 \muJybm\ (Lockman Hole) and 14 \muJybm\ (ELAIS-N1), although the noise is lowest in the centre and increases radially in both images.
+Catalogues from these images were processed to group individual Gaussian islands into complete sources, and remove duplicates.
+The resulting catalogues contain 2,316 and 13,058 sources above 5$\sigma$, for Lockman Hole and ELAIS-N1 respectively.
+Although \cite{de_jong_into_2024} also imaged ELAIS-N1 at 0.\sarc\ 6 and 1.\sarc\ 2, for consistency with Lockman Hole we use only the 0.\sarc\ 3 and 6\sarc\ resolution images.
+
+%The Lockman Hole was the first field to have its full field of view imaged at 0.\sarc\ 3 resolution \citep{sweijen_deep_2022}.
 This took a quarter million cpu hours to produce a single 7-billion pixel image from 8 hours of data.
-The resulting image has a median rms noise of 34 \muJybm , although this is lower in the centre and higher in the outskirts.
+%The resulting image has a median rms noise of 34 \muJybm , although this is lower in the centre and higher in the outskirts.
 From this image, a catalogue of 2,316 5$\sigma$ sources was created, removing duplicates and grouping individual Gaussian islands together into sources.
-
-The first 0.\sarc\ 3 image of ELAIS-N1 was recently published \citep{de_jong_into_2024}, along with companion images at intermediate resolutions of 0.\sarc\ 6 and 1.2\sarc .
+%The first 0.\sarc\ 3 image of ELAIS-N1 was recently published \citep{de_jong_into_2024}, along with companion images at intermediate resolutions of 0.\sarc\ 6 and 1.2\sarc .
 The median rms noise reaches 14 \muJybm\ in the centre of the 0.\sarc\ 3 resolution image, and increases radially.
@@
 
-Thanks to the added collecting area of the international stations, which increases the sensitivity of the array, the 0.\sarc\ 3 resolution images are \textit{deeper} than the 6\sarc\ \ resolution images made with similar or larger amounts of data.
-The derived galaxy properties from \citetalias{best_lofar_2023} (see \S~\ref{subsec:dr1}) were for LoTSS Deep Fields DR1, and thus we have to limit the source catalogues we wish to use.
-For both Lockman Hole and ELAIS-N1 catalogues derived from the 0.\sarc\ 3 resolution images, we first find the local rms in the companion 6\sarc\ \ resolution noise maps\footnote{\label{dr1}Available at \href{https://lofar-surveys.org/deepfields.html}{https://lofar-surveys.org/deepfields.html}} at the position of each source.
-Using the peak brightness of the source in the catalogue from the 0.\sarc\ 3 resolution image, we calculate the peak-to-noise ratio and only select sources which would be detectable at the 5$\sigma$ limit.
-This ensures we do not have sources in the 0.\sarc\ 3 resolution catalogue which do not have a match in DR1.
+Thanks to the added collecting area of the international stations, which increases the sensitivity of the array, the 0.\sarc\ 3 resolution image for ELAIS-N1 is \textit{deeper} than the Deep Fields DR1.
+To ensure consistently built samples, and make use of DR1 information, we limit the high-resolution catalogues.
+For both fields,
+%The derived galaxy properties from \citetalias{best_lofar_2023} (see \S~\ref{subsec:dr1}) were for LoTSS Deep Fields DR1, and thus we have to limit the source catalogues we wish to use.
+For both Lockman Hole and ELAIS-N1 catalogues derived from the 0.\sarc\ 3 resolution images,
+we first find the local rms in the companion 6\sarc\ \ resolution noise maps\footnote{\label{dr1}Available at \href{https://lofar-surveys.org/deepfields.html}{https://lofar-surveys.org/deepfields.html}} at the position of each source in the 0.\sarc\ 3 resolution catalogue.
+Using the peak brightness of the source in the catalogue from the 0.\sarc\ 3 resolution image, we calculate the peak-to-noise ratio and only select sources which would be detectable at the 5$\sigma$ limit in the 6\sarc\ resolution image, effectively ensuring all sources will have a match in DR1.
 We then cross-match based on radio position, to provide the radio information from the catalogues derived from both the 0.\sarc\ 3 resolution image and the 6\sarc\ \ resolution image.
-We trim the area down for each field using their respective optical starmasks\textsuperscript{\ref{dr1}}, by selecting sources which are not contained in the null areas of these masks.
+We trim the area down for each field using their respective optical starmasks\textsuperscript{\ref{dr1}} by selecting sources which are not contained in the null areas of these masks.
 This removes sources where the multi-wavelength information is impacted by bright stars.
@@
 By creating a model which predicts the upper limit of $T_b$ achievable by star formation \citep[based on][]{condon_radio_1992}, if the value is \textit{above} that upper limit, the radio emission must be due to AGN activity.
-\citetalias{morabito_identifying_2022} showed that this AGN identification can be done at 150 MHz using sub-arcsecond resolution; this works because $T_b\propto\nu^{-2}\Omega^{-1}$ where $\nu$ is the observed frequency and $\Omega$ is the solid angle of the emitting region -- for unresolved sources, this corresponds to the deconvolved beam size.
-We refer the reader to \S~3 of \citetalias{morabito_identifying_2022} for more details, but note that this method yields conservative, positive AGN identifications.
+\citetalias{morabito_identifying_2022} showed that this AGN identification can be done at 150 MHz using sub-arcsecond resolution;
+%this works because $T_b\propto\nu^{-2}\Omega^{-1}$ where $\nu$ is the observed frequency and $\Omega$ is the solid angle of the emitting region -- for unresolved sources, this corresponds to the deconvolved beam size.
+we refer the reader to \S~3 of \citetalias{morabito_identifying_2022} for more details.
+Note that this method yields conservative, positive AGN identifications.
 There may be lower surface brightness AGN emission which is not identified, and it is also possible \citep[although unlikely, according to][]{condon_radio_1992} that extreme, compact nuclear starbursts could produce AGN-like values of $T_b$.
@@
 
+\begin{figure*}
+    \centering
+    \includegraphics[width=0.98\textwidth]{figures/deep_fields_RLFs.png}
+    \caption{\textit{Left:} Validation of our re-calculated galaxy RLFs (dotted lines) for the smaller area considered here with the previously published RLFs (solid lines).
+\textit{Middle:} RLFs calculated by process rather than galaxy.
+\textit{Right top:} RLFs calculated here by galaxy classifications (dotted lines) and by physical process (solid lines; orange for AGN and green for SF).
+\textit{Bottom right:} $\Delta$RLF for both AGN and star formation.
+}
+    \label{fig:rlfs}
+    \script{comparison_plot.py}
+\end{figure*}
+
 \section{Results}
@@
 This is because the SF population has larger source counts, so the total impact of moving some flux density from the SF to AGN population is less extreme.
-
-\begin{figure*}
-    \centering
-    \includegraphics[width=0.98\textwidth]{figures/deep_fields_RLFs.png}
-    \caption{\textit{Left:} Validation of our re-calculated galaxy RLFs (dotted lines) for the smaller area considered here with the previously published RLFs (solid lines).
-\textit{Middle:} RLFs calculated by process rather than galaxy.
-\textit{Right top:} RLFs calculated here by galaxy classifications (dotted lines) and by physical process (solid lines; orange for AGN and green for SF).
-\textit{Bottom right:} $\Delta$RLF for both AGN and star formation.
-}
-    \label{fig:rlfs}
-    \script{comparison_plot.py}
-\end{figure*}
 
@@
 If we disregard the lowest luminosity bins, it seems that we see approximately a 30 - 50 percent increase in AGN activity over what is measured when using galaxy classifications alone, for all redshifts.
+This is borne out by the values in the bottom line of Tab.~\ref{tab:intvalues}, where the difference in the integrated RLFs shows an excess of AGN contribution of $\sim$ 50 percent for the three middle redshift bins.
 %The ratio for the SIMBA RLFs shows strong redshift evolution, and much higher ratio values.
@@
 
+In this letter we have presented, for the first time, RLFs constructed by separating radio emission into either SF or AGN processes, for 7,497 sources.
+We use brightness temperature to identify AGN components in galaxies with no radio excess, and separate the radio emission for AGN and SF.
+We compare these with classical RLFs constructed using overall galaxy classifications, and find a hidden AGN population in sources with \llof $\lesssim 10^{24}$\wphz .
+The integrated impact of this is approximately a 50 percent increase in AGN activity across intermediate redshifts ($0.5 < z < 2.0$), and a doubling in the lowest redshift bin ($0.003 < z < 0.3$).
+The impact on the star forming galaxy population is less severe, as this population is much larger, amounting to an integrated decrease in the RLF of $\sim$10-20 percent.
+
+This study is only now possible thanks to LOFAR's exceptional combination of resolution and field of view.
+The sub-arcsecond resolution is sufficient for brightness temperature identification of AGN, and the field of view means that we have been able to do this for 1,022 sources.
+We have used conservative limits for AGN identification, which means that there is likely more AGN emission still remaining `uncounted' here.
+Future work to build a forward modelling method, that can also make use of the intermediate (0.\sarc\ 6 and 1.\sarc\ 2) resolution images will improve the AGN and SF separation.
+Splitting radio emission into category by physical process also opens up a pathway for more direct comparison to cosmological simulations, which we will address in a future paper.
+
+\vspace{-0.5in}
+
 \section*{Data Availability}
-Reference showyourwork on github
+All data used in this paper is either publicly available on \href{https://lofar-surveys.org}{https://lofar-surveys.org} or from the relevant publications cited herein.
+We have prepared this manuscript using using the showyourwork\footnote{ \href{https://github.com/rodluger/showyourwork}{https://github.com/rodluger/showyourwork}} open source scientific article workflow, which was first introduced in \cite{luger_mapping_2021}.
+This ensures complete transparency.
+The code is available at \hl{update link when ready to publish}.
 
@@
 The ILT resources have benefited from the following recent major funding sources: CNRS-INSU, Observatoire de Paris and Université d'Orléans, France; BMBF, MIWF-NRW, MPG, Germany; Science Foundation Ireland (SFI), Department of Business, Enterprise and Innovation (DBEI), Ireland; NWO, The Netherlands; The Science and Technology Facilities Council, UK; Ministry of Science and Higher Education, Poland; The Istituto Nazionale di Astrofisica (INAF), Italy.
-
 This research made use of the Dutch national e-infrastructure with support of the SURF Cooperative (e-infra 180169) and the LOFAR e-infra group.
@@
 (grant CHTB00) through the John von Neumann Institute for Computing (NIC).
-
 This research made use of the University of Hertfordshire high-performance computing facility and the LOFAR-UK computing facility located at the University of Hertfordshire and supported by STFC [ST/P000096/1], and of the Italian LOFAR IT computing infrastructure supported and operated by INAF, and by the Physics Department of Turin university (under an agreement with Consorzio Interuniversitario per la Fisica Spaziale) at the C3S Supercomputing Centre, Italy.
@@
 
-\clearpage
-\pagebreak
+%\clearpage
+%\pagebreak
 

### 2024-10-01 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 P.N. Best$^{4}$,
-B. Yue$^{4}$, and friends
+B.-H. Yue$^{4}$\orcidlink{0009-0009-8935-2929}, 
+T.W. Shimwell$^{3,5}$, and friends
 \\}\\ 
@@
 $^{2}$Institute for Computational Cosmology, Department of Physics, Durham University, South Road, Durham DH1 3LE, UK \\ 
-$^{3}$Leiden \\
-$^{4}$Edinburgh \\}
+$^{3}$Leiden Observatory, Leiden University, PO Box 9513, 2300 RA Leiden, The Netherlands \\
+$^{4}$Institute for Astronomy, University of Edinburgh Royal Observatory, Blackford Hill, Edinburgh, EH9 3HJ, UK \\
+$^{5}$ASTRON, Netherlands Institute for Radio Astronomy, Oude Hoogeveensedijk 4, 7991 PD, Dwingeloo, the Netherlands \\}
 
@@
 \begin{abstract}
-Both star formation (SF) and Active galactic nuclei (AGN) play an important role in galaxy evolution.
+Both star formation (SF) and Active Galactic Nuclei (AGN) play an important role in galaxy evolution.
 Quantifying their relative importance in a statistical way can be done using radio luminosity functions.
 Until now these have relied on galaxy classifications, with sources which have a mixture of radio emission from SF and AGN assigned to either a star-forming galaxy or an AGN.
-This can lead to the under-estimation of the relevance of AGN, or an over-estimation of SF.
-Brightness temperature methods were demonstrated to be effective at 144 MHz using sub-arcsecond imaging with the International LOFAR Telescope (ILT) in \cite{morabito_identifying_2022}.
+This can lead to the under- or over-estimation of the relevance of AGN.
+Brightness temperature methods were demonstrated to be effective at separating radio emission from AGN and SF at 144 MHz using sub-arcsecond imaging with the International LOFAR Telescope (ILT).
 We use the combination of sub-arcsecond and arcsecond resolution imaging of 7,497 sources in the Lockman Hole and ELAIS-N1 fields to identify AGN components in the high resolution images and subtract them from the total flux density, leaving behind flux density from SF only.
 We construct, for the first time, radio luminosity functions by physical process rather than galaxy classification, revealing a hidden AGN population at \llof $<10^{24}$\wphz .
-This population is $\sim$50 percent more than expected for $0.5<z<2.0$.
+This population is $\sim$50 percent more than expected for $0.5<z<2.0$ when comparing to RLFs by galaxy classification.
 The star forming population is less impacted, with only $\sim$10 - 20 percent less star formation than expected.
-The contribution of `hidden' AGN to the overall radio population, especially at \llof $<10^{24}$\wphz , is significant, and only revealed thanks to the combination of the ILT's resolution and wide field of view.
+The contribution of `hidden' AGN to the overall radio population, especially at \llof $<10^{24}$\wphz\ can have significant implications for the cosmic star formation rate and kinetic luminosity densities.
+This is only revealed thanks to the combination of the ILT's resolution and wide field of view.
 \vspace{3.7in}
@@
 The extent to which active galactic nuclei (AGN) impact galaxy formation is a major open question in astrophysics.
-It is clear that there is an interaction, from both observational and theoretical standpoints.
+It is clear that they influence each other, from both observational and theoretical standpoints.
 Observations have revealed tight scaling relations between the mass of a super-massive black hole and its host galaxy \citep[see, e.g.][and references therein]{kormendy_coevolution_2013}, while cosmological simulations require some form of AGN feedback \citep{bower_breaking_2006,croton_many_2006} to be able to reproduce the observed galaxy population.
 %Galaxies build up their stellar mass through star formation, while super-massive black holes are expected to provide either negative or positive feedback to either suppress or stimulate galaxy growth, although whether we can definitively pinpoint which type of feedback is an open question \citep{ward_cosmological_2022}.
-The cosmic history of AGN activity and star formation (SF), and by extension the interplay between them, are interlinked.
+The cosmic histories of AGN activity and star formation (SF), and by extension the interplay between them, are interlinked.
 It is necessary to quantify the contribution of each process to understand the importance of either or both of them.
 Relying on overall galaxy classification can over- or under-estimate the contributions of each process; for example, by considering all of the radio emission in a galaxy classified as `star-forming' to be due to star formation, when some may be due to AGN.
-Quantifying this is the goal of this letter.
+Quantifying these contributions is the goal of this letter.
 
-Decomposing AGN activity from star formation is a difficult problem at any wavelength.
+Decomposing AGN activity from star formation (SF) on a per-galaxy basis is a difficult problem at any wavelength.
 Efforts have been made to do this via mid-infrared (MIR) spectral decomposition using template fitting \citep[e.g.,][]{laurent_mid-infrared_2000,hernan-caballero_resolving_2015,li_active_2024}.
-Multi-wavelength approaches use spectral energy distribution (SED) fitting, which utilises information from a much broader number of observed bands to jointly fit galaxy and AGN components \citep[e.g.][]{calistro_rivera_agnfitter_2016,boquien_cigale_2019,pacifici_art_2022}.
-However, SED fitting is costly as it relies on observations using many different instruments over the same part of the sky, which then have to be carefully cross-matched.
+Multi-wavelength approaches use spectral energy distribution (SED) fitting, which uses information from a much broader number of observed bands to jointly fit galaxy and AGN components \citep[e.g.][]{calistro_rivera_agnfitter_2016,boquien_cigale_2019,pacifici_art_2023}.
+However, SED fitting is costly as it relies on observations using many different instruments over the same part of the sky, which then have to be carefully cross-matched, and there can be degeneracies in the fits.
 Varying selection effects across the different bands can also present a problem.
@@
 An alternative method to unambiguously identify the AGN component(s) in a galaxy is through brightness temperature, $T_b$, measurements in the radio waveband.
-Radio emission has the advantage of being unobscured by dust or gas, meaning there are minimal selection effects.
+Observations of radio emission have the advantage of being unobscured by dust or gas, so there are minimal observational biases.
 %Brightness temperature is a non-physical temperature which is defined as the temperature of a blackbody which would produce the observed surface brightness.
@@
 Surface brightness which is detected above this limit therefore has to be attributed to AGN activity.
-Brightness temperature is inversely proportional to frequency and resolution, $T_b \propto \nu^{-2}\Omega^{-1}$ (where $\Omega \propto \theta_{maj}\theta_{min}$).
-Traditionally brightness temperature measurements require Very Long Baseline Interferometry (VLBI) which achieve milli-arcsec resolutions at GHz frequencies \citep[e.g.,][]{middelberg_mosaiced_2013,herrera_ruiz_faint_2017,radcliffe_nowhere_2018} to achieve enough sensitivity to make meaningful measurements.
-In \cite{morabito_identifying_2022} (hereafter, \citetalias{morabito_identifying_2022}) we demonstrated using sub-arcsecond resolution (0.\sarc\ 3) at 144 MHz with the LOw Frequency ARray \citep[LOFAR;][]{van_haarlem_lofar:_2013}.
-The advantage of LOFAR is its wide field of view: $\sim$6 deg$^2$ from a single observation versus a few to a couple hundred arcmin$^2$ with VLBI at GHz frequencies, depending on whether one is imaging at the phase centre of the observation or using multi-phase centres for correlation \citep{deller_difx-2:_2011,morgan_vlbi_2011}.
-For example, using an 8 hour LOFAR observation of the Lockman Hole \citep{sweijen_deep_2022} we identified 940 AGN via brightness temperature \citepalias{morabito_identifying_2022}, providing a sample two orders of magnitude larger than any VLBI sample of AGN identified at GHz frequencies.
+Brightness temperature is inversely proportional to frequency and resolution, $T_b \propto \nu^{-2}\Omega^{-1}$ (where $\Omega \propto \theta_{\text{maj}}\theta_{\text{min}}$).
+Traditionally brightness temperature measurements require Very Long Baseline Interferometry (VLBI) which reach milli-arcsec resolutions at GHz frequencies \citep[e.g.,][]{middelberg_mosaiced_2013,herrera_ruiz_faint_2017,radcliffe_nowhere_2018} to achieve enough sensitivity to make meaningful measurements.
+In \cite[][hereafter, \citetalias{morabito_identifying_2022}]{morabito_identifying_2022} we demonstrated that this can be done using sub-arcsecond resolution (0.\sarc\ 3) at 144 MHz with the LOw Frequency ARray \citep[LOFAR;][]{van_haarlem_lofar:_2013}.
+The advantage of LOFAR is its wide field of view: $\sim$6 deg$^2$ from a single observation versus a few to a couple hundred arcmin$^2$ with VLBI at GHz frequencies, depending on whether one images a single phase centre or uses multi-phase centres for correlation \citep{deller_difx-2:_2011,morgan_vlbi_2011}.
+For example, using an 8 hour LOFAR observation of the Lockman Hole \citep{sweijen_deep_2022} we identified 940 AGN via brightness temperature \citepalias{morabito_identifying_2022}, providing a sample two orders of magnitude larger than any VLBI sample of AGN identified at GHz frequencies, using a fraction of the observing time.
 
-As part of the LOFAR Two-metre Sky Survey Deep Fields \citep{sabater_lofar_2021,tasse_lofar_2021} campaign, we have now doubled the areal coverage of 0.\sarc\ 3 resolution wide-field imaging.
+As part of the LOFAR Two-metre Sky Survey Deep Fields campaign \cite[][hereafter, \citetalias{best_lofar_2023}]{best_lofar_2023}, we have now doubled the areal coverage of 0.\sarc\ 3 resolution wide-field imaging.
 \cite{de_jong_into_2024} recently published their 0.\sarc\ 3 resolution image of the ELAIS-N1 field, which reaches a depth of 14 $\mu$Jy$\,$bm$^{-1}$ using 32 hours of observations.
-They also demonstrate the flexible resolution which can be achieved with LOFAR, by imaging at 0.\sarc\ 6 and 1.\sarc\ 2 resolution, which are both complementary to the existing 6\sarc\ \ Deep Fields Data Release 1 (DR1).
-Again this has an advantage over VLBI with instruments at GHz frequencies, which require either observing at different frequency bands or with other instruments to acquire information on the flux density of sources at other scales.
-As described in \S~5 of \citepalias{morabito_identifying_2022}, for sources which are unresolved in the 0.\sarc\ 3 resolution images and do not exhibit a radio excess over that predicted the star formation rate (SFR) derived from SED fitting \citep[see][for more details]{best_lofar_2023}, we can separate the AGN activity from SF using a combination of the 0.\sarc\ 3 and 6\sarc\ \ resolution images.
+They also demonstrate the flexible resolution which can be achieved with LOFAR, by imaging at 0.\sarc\ 6 and 1.\sarc\ 2 resolution, which are both complementary to the existing 6\sarc\ \ Deep Fields Data Release 1 \citep[DR1;][]{sabater_lofar_2021,tasse_lofar_2021}.
+This is another advantage over VLBI with instruments at GHz frequencies, which require either observing at different frequency bands or with other instruments to acquire information on the flux density of sources at other scales.
+As described in \S~5 of \citetalias{morabito_identifying_2022}, for sources which are unresolved in the 0.\sarc\ 3 resolution images and do not exhibit a radio excess over that predicted by the star formation rate (SFR) derived from SED fitting \citepalias[see][for more details]{best_lofar_2023}, we can separate the AGN activity from SF using a combination of the 0.\sarc\ 3 and 6\sarc\ \ resolution images.
 
-The 2,483 sources in Lockman Hole \citep{sweijen_deep_2022} and now the 13,058 sources in ELAIS-N1 \citep{de_jong_into_2024}, represents a massive step towards robust statistical studies where the population can be broken into bins of redshift and stellar mass to investigate the AGN and SFG populations in detail.
-\cite{cochrane_lofar_2023} and \cite{kondapally_cosmic_2022} constructed radio luminosity functions (RLFs) from the 6\sarc\ \ resolution imaging that formed the Deep Fields DR1 for the star forming galaxies and radio-excess AGN, respectively.
-Here we construct RLFs, for the first time, by process rather than overall galaxy classification.
-We do this by using a combination of the 0.\sarc\ 3 and 6\sarc\ \ resolution imaging from Lockman Hole and ELAIS-N1.
-We investigate the cosmic evolution of AGN activity and SF by dividing the sample into redshift bins.
+The 2,483 sources in Lockman Hole \citep{sweijen_deep_2022} and now the 13,058 sources in ELAIS-N1 \citep{de_jong_into_2024}, represent a massive step towards robust statistical studies where the population can be broken into bins of redshift and stellar mass to investigate the AGN and SFG populations in detail.
+\cite{cochrane_lofar_2023} and \cite{kondapally_cosmic_2022} constructed radio luminosity functions (RLFs) from the DR1 and SED  for the star forming galaxies and radio-excess AGN, respectively.
+Here we construct RLFs, for the first time, by physical process rather than overall galaxy classification.
+%We do this by using a combination of the 0.\sarc\ 3 and 6\sarc\ \ resolution imaging from Lockman Hole and ELAIS-N1.
+We investigate the cosmic evolution of AGN activity and SF by dividing the sample into redshift bins out to $z = 2.5$.
 
-In this paper, we first describe the data and methods in \S~\ref{sec:data}, followed by an overview of the methods in \S~\ref{sec:methods}.
-Section~\ref{sec:results} describes the results, followed by
+In this paper, we first describe the data in \S~\ref{sec:data}, followed by an overview of the methods in \S~\ref{sec:methods}. Section~\ref{sec:results} describes the results, followed by 
 % discussion and 
@@
 Throughout this paper, we use a $H_0 = 70$ km$\,$s$^{-1}\,$Mpc$^{-1}$, $\Omega_M = 0.3$, and $\Omega_{\Lambda}= 0.7$ cosmology.
-Radio flux density is defined as $S\propto \nu^{\alpha}$ where $\nu$ is frequnecy and $\alpha$ is radio spectral index.
-
-%Stuff and things \citep{morabito_sub-arcsecond_2022}.
+Radio flux density is defined as $S\propto \nu^{\alpha}$ where $\nu$ is frequency and $\alpha = -0.7$ is radio spectral index.
 
@@
 The LOFAR Two-metre Sky Survey (LoTSS) has both a wide-area \citep{shimwell_lofar_2017,shimwell_lofar_2021,shimwell_lofar_2022} and Deep Fields \citep{tasse_lofar_2021,sabater_lofar_2021} component.
-The Deep Fields Data Release 1 (DR1) in 2021 included Boötes, Lockman Hole, and ELAI-N1, reaching sensitivities of 30, 23, and 20 \muJybm , using $\sim$80, $\sim$100, and $\sim$160 hours, respectively.
+The Deep Fields Data Release 1 (DR1) included Boötes, Lockman Hole, and ELAIS-N1, reaching sensitivities of 30, 23, and 20 \muJybm , using $\sim$80, $\sim$100, and $\sim$160 hours of observations, respectively.
 The resolution of these images is all 6\sarc .
@@
 
-Based on the cross-matched multi-wavelength catalogue \cite{kondapally_lofar_2021}, \cite{best_lofar_2023} (hereafter, \citetalias{best_lofar_2023}) carried out detailed SED fitting for all radio-detected sources across the three deep fields, using the multi-wavelength data.
-Full details can be found in their paper, but here we outline the relevant details.
-All sources detected above 5$\sigma$ in their respective radio images had their far-infrared (FIR) to ultraviolet (UV) data fit using four different SED fitting software packages.
-These were: MAGPHYS \citep{da_cunha_simple_2008}, BAGPIPES \citep{carnall_inferring_2018,carnall_how_2019}, CIGALE \citep{burgarella_star_2005,noll_analysis_2009,boquien_cigale_2019}, and AGN\textsc{fitter} \citep{calistro_rivera_agnfitter_2016}.
+Using the multi-wavelength cross-matched catalogues, \citetalias{best_lofar_2023} carried out detailed SED fitting for all radio-detected sources across the three deep fields.
+Full details can be found in their paper, but here we outline the most relevant details.
+All sources detected above 5$\sigma$ in their respective radio images had their far-infrared (FIR) to ultraviolet (UV) data fit using four different SED fitting software packages, and a `consensus' value was found for the derived galaxy properties using information from the goodness-of-fits.
+%These were: MAGPHYS \citep{da_cunha_simple_2008}, BAGPIPES \citep{carnall_inferring_2018,carnall_how_2019}, CIGALE \citep{burgarella_star_2005,noll_analysis_2009,boquien_cigale_2019}, and AGN\textsc{fitter} \citep{calistro_rivera_agnfitter_2016}.
 The first two (MAGPHYS and BAGPIPES) use an energy balance approach, where optical to UV energy absorbed by dust must be transferred to thermal emission in the FIR to sub-mm.
@@
 
-The derived properties used in this paper are the consensus SFRs, and the galaxy classifications which arise from the SED fitting.
-\citetalias{best_lofar_2023} use the relationship between radio luminosity and SFR to define sources as either \textit{radio excess} or not, based on the expected relationship between radio luminosity and SFR.
+The derived properties used in this paper are the consensus SFRs and the galaxy classifications which arise from the SED fitting.
+\citetalias{best_lofar_2023} use the empirical relationship between radio luminosity and SFR to define sources as either \textit{radio excess} or not.
 The AGN components used to fit the multi-wavelength information indicate whether a source is a radiatively efficient AGN (i.e., is there evidence for a disk/torus structure) or not.
@@
 The first three classes are AGN, and the final class is star forming galaxies (SFGs).
-Five percent of all sources are Unclassified, as they do not have a radio excess and the fits could not determine between radiatively efficient or inefficient.
+Five percent of all sources are unclassified, as they do not have a radio excess and the fits could not determine between radiatively efficient or inefficient.
 
@@
 
-Data for LoTSS is recorded with all available LOFAR stations, which are spread across nine different European countries, with a preponderance of stations in the Netherlands.
-The longest baseline (currently Ireland to Latvia) is $\sim$2,000 km, providing unprecedented resolution at MHz frequencies.
+Data for LoTSS are recorded with all available LOFAR stations, which are spread across eight different European countries, with a preponderance of stations in the Netherlands.
+The longest baseline (currently Ireland to Poland) is $\sim$2,000 km, providing unprecedented resolution at MHz frequencies.
 While LoTSS uses only the stations in the Netherlands for its standard data processing, post-processing of LoTSS data to achieve sub-arcsecond resolution is ongoing both for individual sources with $S>10\,$mJy in the wide-area survey \citep{morabito_sub-arcsecond_2022} and for the entire field of view in the Deep Fields \citep[][Escott et al., in prep; Bondi et al., in prep]{sweijen_deep_2022,de_jong_into_2024}.
 %While the full field of view imaging is currently too computationally expensive for the wide area, we are focusing on the Deep Fields.
-Due to the fact that the station layout of international stations is larger than that of stations in the Netherlands, the station beams have a smaller field of view.
+Due to the fact that the size of international stations is larger than that of stations in the Netherlands, the station beams have a smaller field of view.
 Coupled with bandwidth and time smearing, this limits the field of view effectively to a radius of 1.24 deg, which still provides 6.25 deg$^2$ of coverage\footnote{The imaged field of view is square with 2.5 deg per side.}.
@@
 Both Lockman Hole and ELAIS-N1 have been imaged at 0.\sarc\ 3 resolution \citep[][respectively]{sweijen_deep_2022,de_jong_into_2024}.
-The resulting images have median rms noise of 34 \muJybm\ (Lockman Hole) and 14 \muJybm\ (ELAIS-N1), although the noise is lowest in the centre and increases radially in both images.
+The resulting images have median rms noise of 34 \muJybm\ (Lockman Hole) and 17 \muJybm\ (ELAIS-N1), although the noise is lowest in the centre and increases radially in both images.
+%(and can vary per facet).
 Catalogues from these images were processed to group individual Gaussian islands into complete sources, and remove duplicates.
@@
 
-Thanks to the added collecting area of the international stations, which increases the sensitivity of the array, the 0.\sarc\ 3 resolution image for ELAIS-N1 is \textit{deeper} than the Deep Fields DR1.
+Thanks to the added collecting area of the international stations, which increases the sensitivity of the array, the 0.\sarc\ 3 resolution image for ELAIS-N1 has a lower noise level than the Deep Fields DR1.
 To ensure consistently built samples, and make use of DR1 information, we limit the high-resolution catalogues.
@@
 we first find the local rms in the companion 6\sarc\ \ resolution noise maps\footnote{\label{dr1}Available at \href{https://lofar-surveys.org/deepfields.html}{https://lofar-surveys.org/deepfields.html}} at the position of each source in the 0.\sarc\ 3 resolution catalogue.
-Using the peak brightness of the source in the catalogue from the 0.\sarc\ 3 resolution image, we calculate the peak-to-noise ratio and only select sources which would be detectable at the 5$\sigma$ limit in the 6\sarc\ resolution image, effectively ensuring all sources will have a match in DR1.
+Using the peak brightness of the source in the catalogue from the 0.\sarc\ 3 resolution image, we calculate the peak-to-noise ratio and only select sources which would be detectable at the 5$\sigma$ limit in the 6\sarc\ resolution image, effectively ensuring all sources should have a match in DR1.
 We then cross-match based on radio position, to provide the radio information from the catalogues derived from both the 0.\sarc\ 3 resolution image and the 6\sarc\ \ resolution image.
-We trim the area down for each field using their respective optical starmasks\textsuperscript{\ref{dr1}} by selecting sources which are not contained in the null areas of these masks.
-This removes sources where the multi-wavelength information is impacted by bright stars.
-Finally, we trim the area down to the 2.5 deg$\times$2.5 deg area which is covered by the 0.\sarc\ 3 resolution image.
-We are left with \variable{output/lockman_detectable.txt}sources in Lockman Hole, and \variable{output/en1_detectable.txt}sources in ELAIS-N1.
-We use the source name from the catalogue derived from the 6\sarc\ \ resolution to cross-match to the catalogue from \citetalias{best_lofar_2023}.
+We use optical star masks\textsuperscript{\ref{dr1}} to remove sources with multi-wavelength information affected by bright stars.
+Finally, we trim the area down of the multi-wavelength catalogue to the 2.5 deg$\times$2.5 deg area which is covered by the 0.\sarc\ 3 resolution image.
+We are left with \variable{output/lockman_detectable.txt}sources in Lockman Hole, and %\variable{output/en1_detectable.txt}sources in ELAIS-N1.
+%We use the source name from the catalogue derived from the 6\sarc\ \ resolution to cross-match to the catalogue from \citetalias{best_lofar_2023}.
 
-We determine whether sources are resolved or unresolved in the same manner as \S~3.1 of \cite{shimwell_lofar_2019,shimwell_lofar_2022}.
-Truly unresolved sources in radio maps should have equal peak brightness and integrated flux density, but LOFAR maps are impacted by smearing due to the ionosphere.
-This causes resolved sources to have a lower peak brightness when compared to the integrated flux density, and they can appear spatially extended.
-We start by selecting sources which are fit by a single Guassian component ({\ttfamily S\_CODE = `S'}), and make a cut to remove sources which are larger than a multiple of the beam size.
+We determine whether sources are resolved or unresolved (relevant in the next section) in the same manner as \S~3.1 of \cite{shimwell_lofar_2019,shimwell_lofar_2022}, by finding an upper limit to the intrinsically unresolved sources.
+Theoretically, these should have equal peak brightness and total flux density, but LOFAR maps are impacted by smearing due to the ionosphere, and high resolution images have time and bandwidth smearing effects.
+Thus intrinsically unresolved sources can appear spatially extended.
+To find this upper limit, we take all sources which are fit by a single Guassian component ({\ttfamily S\_CODE = `S'}), and are smaller than a multiple of the beam size.
 For the Lockman Hole, we use a factor of 3, but for ELAIS-N1 we find that a factor of 4 provided a better sample.
-This may be because half of the data for ELAIS-N1 had been pre-averaged to 2 sec integration times, which will increase the time smearing \citep[see \S~2 in][]{de_jong_into_2024}.
-We fit a Sigmoid function to find the upper 99.9th percentile of sources considering their integrated flux density to peak brightness ratio as a function of the local signal to noise.
+This may be because the ELAIS-N1 field uses multiple nights, which have varying ionospheric conditions \citep[see \S~2 in][]{de_jong_into_2024}.
+%This may be because half of the data for ELAIS-N1 had been pre-averaged to 2 sec integration times, which will increase the time smearing \citep[see \S~2 in][]{de_jong_into_2024}. 
+Plotting the total flux density to peak brightness ratio of this sample as a function of the local signal to noise, we fit a Sigmoid function to the upper 99.9th percentile.
+Applying this cut to the entire catalogue separates sources into unresolved or resolved.
+Any source which does not have {\ttfamily S\_CODE = `S'} is also considered resolved.
+%We fit a Sigmoid function to find the upper 99.9th percentile of sources considering their total flux density to peak brightness ratio as a function of the local signal to noise.
 Anything above this boundary is considered to be resolved.
@@
 It is not a physical temperature, but rather the temperature of a black body  which would produce the observed spectral radiance (assuming the Rayleigh-Jeans approximation) at long wavelengths.
-By creating a model which predicts the upper limit of $T_b$ achievable by star formation \citep[based on][]{condon_radio_1992}, if the value is \textit{above} that upper limit, the radio emission must be due to AGN activity.
+By creating a model which predicts the upper limit of $T_b$ achievable by SF \citep[based on][]{condon_radio_1992}, if the value is \textit{above} that upper limit, the radio emission must be due to AGN activity.
 \citetalias{morabito_identifying_2022} showed that this AGN identification can be done at 150 MHz using sub-arcsecond resolution;
@@
 \subsection{Separation of AGN activity and SF}
-Once the AGN activity is identified from the 0.\sarc\ 3 resolution images, it is possible to separate it from star formation using the total flux density from the 6\sarc\ \ resolution images.
+Once the AGN activity is identified from the 0.\sarc\ 3 resolution images, it is possible to separate it from SF using the total flux density from the 6\sarc\ \ resolution images.
 We refer the reader to \S~5 in \citetalias{morabito_identifying_2022} for more details but outline the process here, which is depicted in Fig.~\ref{fig:flowchart}.
-Sources which are identified as radio excess according to the definition in \citetalias{best_lofar_2023} automatically have their total flux density from the 6\sarc\ resolution images added to the AGN category.
-For sources that are not radio excess, we attribute the total flux density from the 6\sarc\ resolution images to SF if in the 0.\sarc\ 3 resolution image they are either resolved or have no detection.
+Sources which are identified as radio excess according to the definition in \citetalias{best_lofar_2023} (and therefore by definition have $>$80 percent of flux density due to AGN) automatically have their total flux density from the 6\sarc\ \ resolution images added to the AGN category.
+For sources that are not radio excess, we attribute the total flux density from the 6\sarc\ \ resolution images to SF if in the 0.\sarc\ 3 resolution image they are either resolved or have no detection.
 Sources which do not meet these criteria are checked and if their $T_b$ is not above the limit for SF their 6\sarc\ resolution total flux density is also added to the SF category.
 The total flux density of any source in the 0.\sarc\ 3 resolution image which is $T_b$-identified as an AGN is assumed to be AGN flux density; this is subtracted from the total flux density measured from the 6\sarc\ \ resolution image, and what remains is assumed to be flux density from SF.
-\variable{output/flowchart_numbers.txt}
+%\variable{output/flowchart_numbers.txt}
 
@@
 We follow the same method of constructing RLFs as in \cite{kondapally_cosmic_2022} and \cite{cochrane_lofar_2023}.
-The 0.\sarc\ 3 resolution image is smaller than the 6\sarc\ resolution image, so we first reproduce the published RLFs (Fig.~\ref{fig:rlfs}, left panel) using the galaxy classifications from \citetalias{best_lofar_2023} to check that we have implemented the method appropriately, and that the smaller area yields the same results.
+The 0.\sarc\ 3 resolution image is smaller than the 6\sarc\ resolution image, so we first reproduce the published RLFs (Fig.~\ref{fig:rlfs}, left panel) using the galaxy classifications from \citetalias{best_lofar_2023} to check that we have implemented the method appropriately, and that the smaller area is consistent.
 We use the completeness corrections from the respective papers.
 The only correction we do not implement is the photo-$z$ correction used in \cite{cochrane_lofar_2023}.
-The authors note that this correction is negligible, and as it is not implemented in \cite{kondapally_cosmic_2022} we chose to keep the method for RLF construction in this paper consistent across both samples.
+The authors note that this correction is small, and as it is not implemented in \cite{kondapally_cosmic_2022} we chose to keep the method for RLF construction in this paper consistent across both samples.
 We use the same redshift range, $0.003 < z < 0.3$, to verify that we have good agreement with the published RLFs.
@@
 The completeness corrections are flux density dependent, and we use the overall total flux density of a galaxy, rather than the flux density of the process, which will be smaller if the radio emission from SF and AGN have been split.
-We still use the overall galaxy completeness correction because this is based on whether or not a \textit{source} will be detectable given the background rms.
-This only relies on the total flux density, not the flux density of individual components.
-Even when considering only the SF (or AGN) radio emission in a source, its overall detectability is still derived from the combination of all radio emission processes present.
+This is because the completeness correction is based on whether or not a \textit{source} will be detectable given the background rms, not any individual component.
+%This only relies on the total flux density, not the flux density of individual components. 
+%Even when considering only the SF (or AGN) radio emission in a source, its overall detectability is still derived from the combination of all radio emission processes present. 
 These RLFs for the physical processes of SF and AGN are shown in the middle panel of Fig.~\ref{fig:rlfs}, again for  $0.003 < z < 0.3$.
@@
 \textit{Right top:} RLFs calculated here by galaxy classifications (dotted lines) and by physical process (solid lines; orange for AGN and green for SF).
-\textit{Bottom right:} $\Delta$RLF for both AGN and star formation.
+\textit{Bottom right:} $\Delta$RLF for both AGN and SF.
 }
@@
 We show $\Delta$RLF in the bottom right hand panel in Fig.~\ref{fig:rlfs}.
-We see that above 10$^{24}\,$\wphz\ the ratio for AGN is unity, which is expected where radio excess sources dominate the population.
+We see that above 10$^{24}\,$\wphz\ the ratio for AGN is unity, which is expected as this is where radio excess sources dominate the population.
 Below that, the ratio for AGN climbs up to $\sim$1.3 and finally almost up to 2 at the lowest radio powers probed.
@@
 
-To show the integrated effect of $\Delta$RLF, we integrate each RLF separately across the luminosity bins where both RLFs are defined, and then take the ratio of the areas.
-For example, for star formation we integrate the RLF for the SF process and the RLF for SFGs, and divide the first by the second.
-These numbers are always less than unity since the SF process is over-estimated in the SFG population, and this has been corrected by moving $T_b$-identified AGN flux density contributions to the AGN process RLF.
-These are reported in Tab.~\ref{tab:intvalues}, and show that while the SF process  only decreases by $\sim$10 percent in the lowest redshift bin, the AGN contribution increases by almost a factor of 2.
-It is evident from the bottom right panel of Fig.~\ref{fig:rlfs} that this `hidden' AGN population extracted from SFGs shows up in the radio AGN population at $L_{\textrm{144MHz}} < 10^{24}\,$\wphz .
-
-\variable{output/integrated_differences.txt}
-
-In Fig.~\ref{fig:evolution} we show the redshift evolution for star formation (left) and AGN (right).
+In Fig.~\ref{fig:evolution} we show the redshift evolution for SF (left) and AGN (right).
 We use similar redshift bins to \cite{kondapally_cosmic_2022} and \cite{cochrane_lofar_2023} for consistency.
-Some bins become unconstrained for the RLFs by process due to the shuffling of sources between the SF and AGN populations, yielding positve y-axis values.
-We remove these by cutting out bins where there are more than 4 dex difference between the overall RLF and the process RLF.
-For the SF RLF we additionally remove points $>1.8$.
+Some bins become unconstrained for the RLFs by process due to the shuffling of sources between the SF and AGN populations, yielding positive y-axis values.
+We remove these by cutting out bins where there is more than a 4 dex difference between the overall RLF and the process RLF.
+For the SF RLF we additionally remove points $>-1.8$.
 We reproduce previous observational results that while the SF population has a strong redshift evolution, the AGN population does not.
@@
 The bottom panels in Fig~\ref{fig:evolution} show $\Delta$RLF, the same as in the bottom right hand panel of Fig~\ref{fig:rlfs}.
-For the SF population, $\Delta$RLF is close to unity for $L_{\textrm{144MHz}} < 10^{24}\,$\wphz in the lowest redshift bin, and then this begins to drop at higher luminosities.
+For the SF population, $\Delta$RLF is close to unity for $L_{\textrm{144MHz}} < 10^{24}\,$\wphz\ in the lowest redshift bin, and then this begins to drop at higher luminosities.
 The higher redshift bins show similar behaviour, perhaps reaching unity at increasingly higher values of $L_{\textrm{144MHz}}$, although deeper data is needed to confirm this.
@@
 For the AGN population, $\Delta$RLF has almost the opposite behaviour, converging to unity \textit{above} $\sim$\llof $10^{24}\,$\wphz\ for the lowest redshift bin.
-This is driven by the fact that the flux density of radio excess sources is entirely attributed to AGN, and these sources tend to have higher radio luminosities.
+This is driven by the flux density of radio excess sources being entirely attributed to AGN, which tend to have higher radio luminosities.
 The convergence happens at higher luminosity values for higher redshifts.
-Below this convergence, across all redshift bins, $\Delta$RLF increases.
+Below this convergence, $\Delta$RLF increases across all redshift bins.
 The maximum value for each redshift bin approaches $\sim$2, although the fact that in the lowest redshift bin the value stabilises at $\sim1.3$ and only climbs higher at the very lowest luminosities (\llof $<10^{22}\,$\wphz ) may be due to incompleteness.
 If we disregard the lowest luminosity bins, it seems that we see approximately a 30 - 50 percent increase in AGN activity over what is measured when using galaxy classifications alone, for all redshifts.
-This is borne out by the values in the bottom line of Tab.~\ref{tab:intvalues}, where the difference in the integrated RLFs shows an excess of AGN contribution of $\sim$ 50 percent for the three middle redshift bins.
+%This is borne out by the values in the bottom line of Tab.~\ref{tab:intvalues}, where the difference in the integrated RLFs shows an excess of AGN contribution of $\sim$ 50 percent for the three middle redshift bins.
+
 %The ratio for the SIMBA RLFs shows strong redshift evolution, and much higher ratio values.
@@
 
+To better quantify the integrated effect of $\Delta$RLF, we integrate each RLF separately across the luminosity bins where both RLFs are defined, and then take the ratio of the areas.
+For example, for SF we integrate the RLF for the SF process and the RLF for SFGs, and divide the first by the second.
+These numbers are always less than unity since the SF process is over-estimated in the SFG population, and this has been corrected by moving $T_b$-identified AGN flux density contributions to the AGN process RLF.
+These are reported in Tab.~\ref{tab:intvalues}, and show that while the SF process  only decreases by $\sim$10 percent in the lowest redshift bin, the AGN contribution increases by almost a factor of 2.
+It is evident from the bottom right panel of Fig.~\ref{fig:rlfs} that this `hidden' AGN population extracted from SFGs shows up in the radio AGN population at $L_{\textrm{144MHz}} < 10^{24}\,$\wphz .
+
+%\variable{output/integrated_differences.txt}
+
 \section{Conclusions}
@@
 In this letter we have presented, for the first time, RLFs constructed by separating radio emission into either SF or AGN processes, for 7,497 sources.
-We use brightness temperature to identify AGN components in galaxies with no radio excess, and separate the radio emission for AGN and SF.
+We use brightness temperature to identify AGN components in galaxies with no radio excess, and separate the radio emission into AGN and SF.
 We compare these with classical RLFs constructed using overall galaxy classifications, and find a hidden AGN population in sources with \llof $\lesssim 10^{24}$\wphz .
-The integrated impact of this is approximately a 50 percent increase in AGN activity across intermediate redshifts ($0.5 < z < 2.0$), and a doubling in the lowest redshift bin ($0.003 < z < 0.3$).
+The integrated impact of this is approximately a 50 percent increase in the radio luminosity output due to AGN across intermediate redshifts ($0.5 < z < 2.0$), and a doubling in the lowest redshift bin ($0.003 < z < 0.3$).
 The impact on the star forming galaxy population is less severe, as this population is much larger, amounting to an integrated decrease in the RLF of $\sim$10-20 percent.
+These impacts can have implications for the cosmic SFR density and kinetic luminosity density, which will be the topic of a future paper.
 
 This study is only now possible thanks to LOFAR's exceptional combination of resolution and field of view.
-The sub-arcsecond resolution is sufficient for brightness temperature identification of AGN, and the field of view means that we have been able to do this for 1,022 sources.
+The sub-arcsecond resolution is sufficient for brightness temperature identification of AGN, and the field of view means that we have been able to do this for 1,022 sources (out of $>$4,000).
 We have used conservative limits for AGN identification, which means that there is likely more AGN emission still remaining `uncounted' here.
@@
 LKM is grateful for support from UKRI [MR/T042842/1]. 
+FS appreciates the support of STFC [ST/Y004159/1]. 
+JMGHJdJ is grateful for support from the project CORTEX (NWA.1160.18.316) of the research programme NWA-ORC which is (partly) financed by the Dutch Research Council (NWO). 
+BY thanks the University of Edinburgh and Leiden Observatory for support through the Edinburgh–Leiden joint studentship.
+PNB is grateful for support from the UK STFC via grants [ST/V000594/1] and [ST/Y000951/1]
 
@@
 The ILT resources have benefited from the following recent major funding sources: CNRS-INSU, Observatoire de Paris and Université d'Orléans, France; BMBF, MIWF-NRW, MPG, Germany; Science Foundation Ireland (SFI), Department of Business, Enterprise and Innovation (DBEI), Ireland; NWO, The Netherlands; The Science and Technology Facilities Council, UK; Ministry of Science and Higher Education, Poland; The Istituto Nazionale di Astrofisica (INAF), Italy.
-This research made use of the Dutch national e-infrastructure with support of the SURF Cooperative (e-infra 180169) and the LOFAR e-infra group.
+
+This research made use of the Dutch national e-infrastructure with support of the SURF Cooperative (e-infra 180169 and EINF-6218) and the LOFAR e-infra group.
+
+This work is co-funded by the EGI-ACE project (Horizon 2020) under Grant number 101017567. This work was sponsored by NWO Domain Science for the use of the national computer facilities.
+
 The Jülich LOFAR Long Term Archive and the German LOFAR network are both coordinated and operated by the Jülich Supercomputing Centre (JSC), and computing resources on the supercomputer JUWELS at JSC were provided by the Gauss Centre for Supercomputing e.V.

### 2024-10-01 src/scripts/RLF_evolution.py
comment

@@ -207,7 +207,7 @@ fig.savefig(paths.figures / 'RLF_evolution.png',dpi=300)
 fig.clear()
 plt.close()
 
-## write a table
+## write an output table
 with open( paths.output / 'integrated_differences.txt', 'w' ) as f:
     f.write( '\\begin{table}\n' )
     f.write( '    \\centering\n' )

### 2024-10-02 src/scripts/comparison_plot.py
uncertainties on ratio

@@ -75,12 +75,14 @@ p1.plot( kondapally['logL150'], kondapally['logPhi'], color=kond, label='Kondapa
 x, y, dy, idx1, idx2 = get_values( lum_bin_cens, gal_agn_lum_func, e_gal_agn_lum_func )
 p1.plot( x, y, color=agngalc, label='AGN galaxies', linewidth=3, linestyle='dotted' )
 p1.fill_between(x[idx1], y[idx1]-dy[idx1], y[idx1]+dy[idx1], color=agngalc, alpha=0.4, ec=None)
+agn_idx1 = idx1
 if len(idx2) > 0:
     for idx22 in idx2:
         p1.fill_between(x[idx22], y[idx22]-dy[idx22], y[idx22]+dy[idx22], color=agngalc, alpha=0.1, hatch='xxx', ec=None)
 x, y, dy, idx1, idx2 = get_values( lum_bin_cens, gal_sf_lum_func, e_gal_sf_lum_func )
 p1.plot( x, y, color=sfggalc, label='SF galaxies', linewidth=3, linestyle='dotted' )
 p1.fill_between( x[idx1], y[idx1]-dy[idx1], y[idx1]+dy[idx1] , color=sfggalc, alpha=0.4, ec=None)
+sf_idx1 = idx1
 if len(idx2) > 0:
     for idx22 in idx2:
         p1.fill_between( x[idx22], y[idx22]-dy[idx22], y[idx22]+dy[idx22] , color=sfggalc, alpha=0.1, hatch='xxx', ec=None)
@@ -102,16 +104,18 @@ p2.plot( kondapally['logL150'], kondapally['logPhi'], color=kond, label='Kondapa
 x, y, dy, idx1, idx2 = get_values( lum_bin_cens, agn_lum_func, e_agn_lum_func )
 p2.plot( x, y, color=agnc, label='AGN process', linewidth=3, alpha=0.75 )
 p2.fill_between( x[idx1], y[idx1]-dy[idx1], y[idx1]+dy[idx1], color=agnc, alpha=0.4, ec=None)
+agn_valid = np.intersect1d( agn_idx1, idx1 )
 if len(idx2) > 0:
     for idx22 in idx2:
         p2.fill_between( x[idx22], y[idx22]-dy[idx22], y[idx22]+dy[idx22], color=agnc, alpha=0.1, hatch='xxx', ec=None)
-
[…]

### 2024-10-02 src/scripts/RLF_evolution.py
indices on evolution

@@ -115,6 +115,8 @@ print('SF')
 sf_delta_int = []
 e_sf_delta_int = []
 
+z_sf_valid = []
+
 ## STAR FORMATION
 p1 = plt.axes([0.07,0.42,sbsizex*fsizey/fsizex,0.6*sbsizey])
 for i in np.arange(0,len(z_lum_bins)):
@@ -122,6 +124,7 @@ for i in np.arange(0,len(z_lum_bins)):
     x, y, dy, idx1, idx2 = get_values( lum_bin_cens, z_gal_sf_lum_func[i], e_z_gal_sf_lum_func[i] )
     p1.plot( x, y, color=zcols_sf[i], linewidth=3, alpha=0.75, linestyle='dotted' )
     p1.fill_between( x[idx1], y[idx1]-dy[idx1], y[idx1]+dy[idx1], color=zcols_sf[i], alpha=0.4, ec=None )
+    sf_idx1 = idx1
     # use rectangular integration 
     gal_trapz = np.sum( np.power( 10., y[idx1] ) * dl )
     e_gal_trapz = np.sqrt( np.sum( np.power( dy[idx1] * np.log( 10. ) * np.power( 10., y[idx1] ), 2. ) ) ) * dl 
@@ -129,6 +132,8 @@ for i in np.arange(0,len(z_lum_bins)):
     x, y, dy, idx1, idx2 = get_values( lum_bin_cens, z_sf_lum_func[i], e_z_sf_lum_func[i] )
     p1.plot( x, y, color=zcols_sf[i], label='{:s} < z < {:s}'.format(str(zbin_starts[i]),str(zbin_ends[i])), linewidth=3 )
     p1.fill_between( x[idx1], y[idx1]-dy[idx1], y[idx1]+dy[idx1], color=zcols_sf[i], alpha=0.4, ec=None, hatch='x' )
+    sf_valid = np.intersect1d( sf_idx1, idx1 )
+    z_sf_valid.append(sf_valid)
     # use rectangular integration
     trapz = np.sum( np.power( 10., y[idx1] ) * dl )
     e_trapz = np.sqrt( np.sum( np.power( dy[idx1] * np.log( 10. ) * np.power( 10., y[idx1] ), 2. ) ) ) * dl
@@ -157,6 +162,12 @@ for i in np.arange(0,len(z_lum_bins)):
[…]

### 2024-10-02 src/scripts/RLF_evolution.py
indices on evolution

@@ -123,7 +123,7 @@ for i in np.arange(0,len(z_lum_bins)):
     ## Galaxies
     x, y, dy, idx1, idx2 = get_values( lum_bin_cens, z_gal_sf_lum_func[i], e_z_gal_sf_lum_func[i] )
     p1.plot( x, y, color=zcols_sf[i], linewidth=3, alpha=0.75, linestyle='dotted' )
-    p1.fill_between( x[idx1], y[idx1]-dy[idx1], y[idx1]+dy[idx1], color=zcols_sf[i], alpha=0.4, ec=None )
+    #p1.fill_between( x[idx1], y[idx1]-dy[idx1], y[idx1]+dy[idx1], color=zcols_sf[i], alpha=0.4, ec=None )
     sf_idx1 = idx1
     # use rectangular integration 
     gal_trapz = np.sum( np.power( 10., y[idx1] ) * dl )
@@ -131,7 +131,7 @@ for i in np.arange(0,len(z_lum_bins)):
     ## Process
     x, y, dy, idx1, idx2 = get_values( lum_bin_cens, z_sf_lum_func[i], e_z_sf_lum_func[i] )
     p1.plot( x, y, color=zcols_sf[i], label='{:s} < z < {:s}'.format(str(zbin_starts[i]),str(zbin_ends[i])), linewidth=3 )
-    p1.fill_between( x[idx1], y[idx1]-dy[idx1], y[idx1]+dy[idx1], color=zcols_sf[i], alpha=0.4, ec=None, hatch='x' )
+    #p1.fill_between( x[idx1], y[idx1]-dy[idx1], y[idx1]+dy[idx1], color=zcols_sf[i], alpha=0.4, ec=None, hatch='x' )
     sf_valid = np.intersect1d( sf_idx1, idx1 )
     z_sf_valid.append(sf_valid)
     # use rectangular integration
@@ -172,7 +172,7 @@ for i in np.arange(0,len(z_lum_bins)):
     p2.fill_between( x[idx1], y[idx1]-dy[idx1], y[idx1]+dy[idx1], alpha=0.4, color=zcols_sf[i], ec=None )
     if len(idx2) > 0:
         for idx22 in idx2:
-            p2.fill_between( x[idx22], y[idx22]-dy[idx22], y[idx22]+dy[idx22], alpha=0.1, hatch='xxx', color=zcols_sf[i], ec=None )
[…]

### 2024-10-02 src/scripts/RLF_evolution.py
typo

@@ -198,7 +198,7 @@ for i in np.arange(0,len(z_lum_bins)):
     ## Process
     x, y, dy, idx1, idx2 = get_values( lum_bin_cens, z_agn_lum_func[i], e_z_agn_lum_func[i] )
     p3.plot( x, y, color=zcols_agn[i], label='{:s} < z < {:s}'.format(str(zbin_starts[i]),str(zbin_ends[i])), linewidth=3 )
-    agn_valid = np.intsersect1d( agn_idx1, idx1 )
+    agn_valid = np.intersect1d( agn_idx1, idx1 )
     z_agn_valid.append(agn_valid)
     # use rectangular integration
     trapz = np.sum( np.power( 10., y[idx1] ) * dl )

### 2024-10-02 src/scripts/RLF_evolution.py
uncertainties on average

@@ -280,4 +280,25 @@ with open( paths.output / 'integrated_differences.txt', 'w' ) as f:
     f.write( '\\end{table}' )
 
 
+## also add up all the redshifts ...
+sf_mean = np.mean(sf_delta_int)
+## uncertainty on mean, but for independent measurements of the same variable
+e_sf_mean = np.sqrt( np.sum( np.power( (np.asarray(sf_delta_int)+np.asarray(e_sf_delta_int))-sf_mean, 2. ) ) / len(sf_delta_int) ) / len(sf_delta_int)
+## simple addition
+# e_sf_mean = np.sqrt( np.sum( np.power( np.asarray(e_sf_delta_int), 2. ) ) / len(sf_delta_int) )
+
+agn_mean = np.mean(agn_delta_int)
+## uncertainty on mean, but for independent measurements of the same variable
+e_agn_mean = np.sqrt( np.sum( np.power( (np.asarray(agn_delta_int)+np.asarray(e_agn_delta_int))-agn_mean, 2. ) ) / len(agn_delta_int) ) / len(sf_delta_int)
+## simple addition
+# e_agn_mean = np.sqrt( np.sum( np.power( np.asarray(e_agn_delta_int), 2. ) ) / len(agn_delta_int) )
+
+with open( paths.output / 'average_integrated_differences.txt', 'w' ) as f:
+    printstr = 'The mean across all redshift bins is {:1.2f}$\\pm${:1.2f} for SF and {:1.2f}$\\pm${:1.2f} for AGN.  '
+    f.write( printstr )
+
+
+
+
+
 

### 2024-10-02 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Finally, we trim the area down of the multi-wavelength catalogue to the 2.5 deg$\times$2.5 deg area which is covered by the 0.\sarc\ 3 resolution image.
-We are left with \variable{output/lockman_detectable.txt}sources in Lockman Hole, and %\variable{output/en1_detectable.txt}sources in ELAIS-N1.
+We are left with \variable{output/lockman_detectable.txt}sources in Lockman Hole, and \variable{output/en1_detectable.txt}sources in ELAIS-N1.
 %We use the source name from the catalogue derived from the 6\sarc\ \ resolution to cross-match to the catalogue from \citetalias{best_lofar_2023}.
@@
 The total flux density of any source in the 0.\sarc\ 3 resolution image which is $T_b$-identified as an AGN is assumed to be AGN flux density; this is subtracted from the total flux density measured from the 6\sarc\ \ resolution image, and what remains is assumed to be flux density from SF.
-%\variable{output/flowchart_numbers.txt}
+\variable{output/flowchart_numbers.txt}
 
@@
 The higher redshift bins show similar behaviour, perhaps reaching unity at increasingly higher values of $L_{\textrm{144MHz}}$, although deeper data is needed to confirm this.
-The lowest values reached are $\sim$50 percent.
+%The lowest values reached are $\sim$50 percent.
 %Comparing these ratios between observations and simulations, we see qualitatively the same behaviour, although the simulations reach much lower and higher values for both the SF and AGN populations, respectively.
@@
 Below this convergence, $\Delta$RLF increases across all redshift bins.
-The maximum value for each redshift bin approaches $\sim$2, although the fact that in the lowest redshift bin the value stabilises at $\sim1.3$ and only climbs higher at the very lowest luminosities (\llof $<10^{22}\,$\wphz ) may be due to incompleteness.
-If we disregard the lowest luminosity bins, it seems that we see approximately a 30 - 50 percent increase in AGN activity over what is measured when using galaxy classifications alone, for all redshifts.
+The maximum value for each redshift bin approaches $\sim$2, although the lowest redshift bin the value stabilises at $\sim1.3$ and only climbs higher at the very lowest luminosities (\llof $<10^{22}\,$\wphz ), which may be due to incompleteness.
+%If we disregard the lowest luminosity bins, it seems that we see approximately a 30 - 50 percent increase in AGN activity over what is measured when using galaxy classifications alone, for all redshifts.
 %This is borne out by the values in the bottom line of Tab.~\ref{tab:intvalues}, where the difference in the integrated RLFs shows an excess of AGN contribution of $\sim$ 50 percent for the three middle redshift bins.
@@
 
-To better quantify the integrated effect of $\Delta$RLF, we integrate each RLF separately across the luminosity bins where both RLFs are defined, and then take the ratio of the areas.
+To better quantify the integrated effect of $\Delta$RLF, we integrate each RLF separately across the luminosity bins where the uncertainties for both RLFs are well defined, and then take the ratio of the areas.
 For example, for SF we integrate the RLF for the SF process and the RLF for SFGs, and divide the first by the second.
 These numbers are always less than unity since the SF process is over-estimated in the SFG population, and this has been corrected by moving $T_b$-identified AGN flux density contributions to the AGN process RLF.
-These are reported in Tab.~\ref{tab:intvalues}, and show that while the SF process  only decreases by $\sim$10 percent in the lowest redshift bin, the AGN contribution increases by almost a factor of 2.
+These are reported in Tab.~\ref{tab:intvalues}.
+\variable{output/average_integrated_differences.txt}
+%, and show that while the SF process only decreases by $\sim$10 percent in the lowest redshift bin, the AGN contribution increases by almost a factor of 2. 
 It is evident from the bottom right panel of Fig.~\ref{fig:rlfs} that this `hidden' AGN population extracted from SFGs shows up in the radio AGN population at $L_{\textrm{144MHz}} < 10^{24}\,$\wphz .
 
-%\variable{output/integrated_differences.txt}
+\variable{output/integrated_differences.txt}
 

### 2024-10-02 src/scripts/RLF_evolution.py
formatting

@@ -294,7 +294,7 @@ e_agn_mean = np.sqrt( np.sum( np.power( (np.asarray(agn_delta_int)+np.asarray(e_
 # e_agn_mean = np.sqrt( np.sum( np.power( np.asarray(e_agn_delta_int), 2. ) ) / len(agn_delta_int) )
 
 with open( paths.output / 'average_integrated_differences.txt', 'w' ) as f:
-    printstr = 'The mean across all redshift bins is {:1.2f}$\\pm${:1.2f} for SF and {:1.2f}$\\pm${:1.2f} for AGN.  '
+    printstr = 'The mean across all redshift bins is {:1.2f}$\\pm${:1.2f} for SF and {:1.2f}$\\pm${:1.2f} for AGN. '.format( sf_mean, e_sf_mean, agn_mean, e_agn_mean )
     f.write( printstr )
 
 

### 2024-10-02 src/scripts/RLF_evolution.py
reformat table

@@ -252,28 +252,17 @@ plt.close()
 with open( paths.output / 'integrated_differences.txt', 'w' ) as f:
     f.write( '\\begin{table}\n' )
     f.write( '    \\centering\n' )
-    f.write( '    \\begin{tabular}{lccccc}\n' )
+    f.write( '    \\begin{tabular}{llcc}\n' )
     zmin = ' $z_{min}$ '
     zmax = ' $z_{max}$ '
+    hdrstr = zmin + '&' + zmax + ' SF & AGN \\\\ \\hline\n'
+    f.write( hdrstr )
     for i in np.arange(0,len(z_lum_bins)):
-        zmin = zmin + ' & {:s}'.format(str(zbin_starts[i]))
-        zmax = zmax + ' & {:s}'.format(str(zbin_ends[i]))
-    zmin = zmin + ' \\\\ \n'
-    zmax = zmax + ' \\\\ \hline \n'
-    f.write( zmin ) 
-    f.write( zmax )
-    sfstr = 'SF ' 
-    for i in np.arange(0,len(sf_delta_int)):
-        sfstr = sfstr + ' & {:1.2f}'.format(sf_delta_int[i])
-        sfstr = sfstr + '$\\pm$' + '{:1.2f}'.format(e_sf_delta_int[i]) 
-    sfstr = sfstr + ' \\\\ \n' 
-    f.write(sfstr)
-    agnstr = 'AGN ' 
-    for i in np.arange(0,len(agn_delta_int)):
-        agnstr = agnstr + ' & {:1.2f}'.format(agn_delta_int[i])
-        agnstr = agnstr + '$\\pm$' + '{:1.2f}'.format(e_agn_delta_int[i]) 
-    agnstr = agnstr + ' \\\\ \n' 
-    f.write(agnstr)
+        zstr = ' {:s} & {:s} & '.format(str(zbin_starts[i]), str(zbin_ends[i] )
+        sfstr = '{:1.2f}$\\pm${:1.2f} & '.format( sf_delta_int[i], e_sf_delta_int[i] )
+        agnstr = '{:1.2f}$\\pm${:1.2f} \\\\ \n '.format( agn_delta_int[i], e_agn_delta_int[i] )
+        linestr = zstr + sfstr + agnstr
+        f.write(linestr)
     f.write( '    \\end{tabular}\n' )
[…]

### 2024-10-02 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 B.-H. Yue$^{4}$\orcidlink{0009-0009-8935-2929}, 
-T.W. Shimwell$^{3,5}$, and friends
+T.W. Shimwell$^{3,5}$,
+Dominik J. Schwarz
+Marco Bondi
+R.J. van Weeren
 \\}\\ 
@@
 We construct, for the first time, radio luminosity functions by physical process rather than galaxy classification, revealing a hidden AGN population at \llof $<10^{24}$\wphz .
-This population is $\sim$50 percent more than expected for $0.5<z<2.0$ when comparing to RLFs by galaxy classification.
-The star forming population is less impacted, with only $\sim$10 - 20 percent less star formation than expected.
+This population is 1.56$\pm$0.06 more than expected for $0.5<z<2.0$ when comparing to RLFs by galaxy classification.
+The star forming population is less impacted, with only 0.90$\pm$0.02 of the expected star formation.
 The contribution of `hidden' AGN to the overall radio population, especially at \llof $<10^{24}$\wphz\ can have significant implications for the cosmic star formation rate and kinetic luminosity densities.
@@
 Due to the fact that the size of international stations is larger than that of stations in the Netherlands, the station beams have a smaller field of view.
-Coupled with bandwidth and time smearing, this limits the field of view effectively to a radius of 1.24 deg, which still provides 6.25 deg$^2$ of coverage\footnote{The imaged field of view is square with 2.5 deg per side.}.
+Coupled with bandwidth and time smearing, this limits the field of view effectively to a radius of 1.24 deg, which still provides 6.25 deg$^2$ of coverage (the imaged field of view is square with 2.5 deg per side).
 
@@
 For both Lockman Hole and ELAIS-N1 catalogues derived from the 0.\sarc\ 3 resolution images,
-we first find the local rms in the companion 6\sarc\ \ resolution noise maps\footnote{\label{dr1}Available at \href{https://lofar-surveys.org/deepfields.html}{https://lofar-surveys.org/deepfields.html}} at the position of each source in the 0.\sarc\ 3 resolution catalogue.
+we first find the local rms in the companion 6\sarc\ \ resolution noise maps at the position of each source in the 0.\sarc\ 3 resolution catalogue.
 Using the peak brightness of the source in the catalogue from the 0.\sarc\ 3 resolution image, we calculate the peak-to-noise ratio and only select sources which would be detectable at the 5$\sigma$ limit in the 6\sarc\ resolution image, effectively ensuring all sources should have a match in DR1.
@@
 We use optical star masks\textsuperscript{\ref{dr1}} to remove sources with multi-wavelength information affected by bright stars.
-Finally, we trim the area down of the multi-wavelength catalogue to the 2.5 deg$\times$2.5 deg area which is covered by the 0.\sarc\ 3 resolution image.
+Finally, we trim the area down of the multi-wavelength catalogue to the 2.5 deg$\times$2.5 deg area which is covered by the 0.\sarc\ 3 resolution image (although the multi-wavelength coverage in ELAIS is only 63 percent of the multi-wavelength coverage of Lockman Hole).
 We are left with \variable{output/lockman_detectable.txt}sources in Lockman Hole, and \variable{output/en1_detectable.txt}sources in ELAIS-N1.
@@
 
-\vspace{-0.5in}
+%\vspace{-0.5in}
 
 \section*{Data Availability}
-All data used in this paper is either publicly available on \href{https://lofar-surveys.org}{https://lofar-surveys.org} or from the relevant publications cited herein.
-We have prepared this manuscript using using the showyourwork\footnote{ \href{https://github.com/rodluger/showyourwork}{https://github.com/rodluger/showyourwork}} open source scientific article workflow, which was first introduced in \cite{luger_mapping_2021}.
+All data used in this paper is either publicly available on \href{https://lofar-surveys.org}{https://lofar-surveys.org} or from the relevant citations herein, including catalogues, rms images, and star masks.
+We have prepared this manuscript using using the showyourwork (\href{https://github.com/rodluger/showyourwork}{https://github.com/rodluger/showyourwork}) open source scientific article workflow, which was first introduced in \cite{luger_mapping_2021}.
 This ensures complete transparency.
@@
 PNB is grateful for support from the UK STFC via grants [ST/V000594/1] and [ST/Y000951/1]
+RK acknowledges support from the UK STFC via grant [ST/V000594/1].
+RJvW acknowledges support from the ERC Starting Grant ClusterWeb 804208. 
+MB acknowledges support from INAF under the Large Grant 2022 funding scheme (project “MeerKAT and LOFAR Team up: a Unique Radio Window on Galaxy/AGN co-Evolution”).
 
@@
 
-This research made use of the Dutch national e-infrastructure with support of the SURF Cooperative (e-infra 180169 and EINF-6218) and the LOFAR e-infra group.
+%This research made use of the Dutch national e-infrastructure with support of the SURF Cooperative (e-infra 180169 and EINF-6218) and the LOFAR e-infra group.
 
-This work is co-funded by the EGI-ACE project (Horizon 2020) under Grant number 101017567. This work was sponsored by NWO Domain Science for the use of the national computer facilities.
+%This work is co-funded by the EGI-ACE project (Horizon 2020) under Grant number 101017567. This work was sponsored by NWO Domain Science for the use of the national computer facilities.
 
-The Jülich LOFAR Long Term Archive and the German LOFAR network are both coordinated and operated by the Jülich Supercomputing Centre (JSC), and computing resources on the supercomputer JUWELS at JSC were provided by the Gauss Centre for Supercomputing e.V.
+%The Jülich LOFAR Long Term Archive and the German LOFAR network are both coordinated and operated by the Jülich Supercomputing Centre (JSC), and computing resources on the supercomputer JUWELS at JSC were provided by the Gauss Centre for Supercomputing e.V.
 (grant CHTB00) through the John von Neumann Institute for Computing (NIC).

### 2024-10-02 src/scripts/RLF_evolution.py
typo

@@ -258,7 +258,7 @@ with open( paths.output / 'integrated_differences.txt', 'w' ) as f:
     hdrstr = zmin + '&' + zmax + ' SF & AGN \\\\ \\hline\n'
     f.write( hdrstr )
     for i in np.arange(0,len(z_lum_bins)):
-        zstr = ' {:s} & {:s} & '.format(str(zbin_starts[i]), str(zbin_ends[i] )
+        zstr = ' {:s} & {:s} & '.format(str(zbin_starts[i]), str(zbin_ends[i]) )
         sfstr = '{:1.2f}$\\pm${:1.2f} & '.format( sf_delta_int[i], e_sf_delta_int[i] )
         agnstr = '{:1.2f}$\\pm${:1.2f} \\\\ \n '.format( agn_delta_int[i], e_agn_delta_int[i] )
         linestr = zstr + sfstr + agnstr

### 2024-10-02 src/scripts/RLF_evolution.py
& missing

@@ -255,7 +255,7 @@ with open( paths.output / 'integrated_differences.txt', 'w' ) as f:
     f.write( '    \\begin{tabular}{llcc}\n' )
     zmin = ' $z_{min}$ '
     zmax = ' $z_{max}$ '
-    hdrstr = zmin + '&' + zmax + ' SF & AGN \\\\ \\hline\n'
+    hdrstr = zmin + '&' + zmax + ' & SF & AGN \\\\ \\hline\n'
     f.write( hdrstr )
     for i in np.arange(0,len(z_lum_bins)):
         zstr = ' {:s} & {:s} & '.format(str(zbin_starts[i]), str(zbin_ends[i]) )

### 2024-10-30 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Title
-\title[A hidden AGN population]{A hidden Active Galactic Nuclei population: the first radio luminosity functions constructed by physical process instead of galaxy classification}
+\title[A hidden AGN population]{A hidden Active Galactic Nuclei population: the first radio luminosity functions by physical process}
 
@@
 \author[L.K. Morabito]{\parbox{\textwidth}{Leah K. Morabito$^{1,2}$\thanks{E-mail: leah.k.morabito@durham.ac.uk}\orcidlink{0000-0003-0487-6651}
-J.M.G.H.J. de Jong$^{3}$,
-F. Sweijen$^{1}$,
-R. Kondapally$^{4}$,
-P.N. Best$^{4}$,
-B.-H. Yue$^{4}$\orcidlink{0009-0009-8935-2929}, 
-T.W. Shimwell$^{3,5}$,
-Dominik J. Schwarz
-Marco Bondi
-R.J. van Weeren
+R. Kondapally$^{3}$\orcidlink{0000-0001-6127-8151},
+P.N. Best$^{3}$\orcidlink{0000-0001-5081-4801},
+B.-H. Yue$^{3}$\orcidlink{0009-0009-8935-2929}, 
+J.M.G.H.J. de Jong$^{4}$\orcidlink{0000-0001-6876-8719},
+F. Sweijen$^{1}$\orcidlink{0000-0002-6470-7967},
+Marco Bondi$^{5}$\orcidlink{0000-0002-9553-7999},
+Dominik J. Schwarz$^{6}$\orcidlink{0000-0003-2413-0881},
+D.J.B. Smith$^{7}$\orcidlink{0000-0001-9708-253X},
+R.J. van Weeren$^{4}$\orcidlink{0000-0002-0587-1660},
+H.J.A. R\"{o}ttgering$^{4}$\orcidlink{0000-0001-8887-2257},
+T.W. Shimwell$^{4,8}$\orcidlink{0000-0001-5648-9069},
+Isabella Prandoni$^{5}$\orcidlink{0000-0001-9680-7092}
 \\}\\ 
@@
 $^{2}$Institute for Computational Cosmology, Department of Physics, Durham University, South Road, Durham DH1 3LE, UK \\ 
-$^{3}$Leiden Observatory, Leiden University, PO Box 9513, 2300 RA Leiden, The Netherlands \\
-$^{4}$Institute for Astronomy, University of Edinburgh Royal Observatory, Blackford Hill, Edinburgh, EH9 3HJ, UK \\
-$^{5}$ASTRON, Netherlands Institute for Radio Astronomy, Oude Hoogeveensedijk 4, 7991 PD, Dwingeloo, the Netherlands \\}
+$^{3}$Institute for Astronomy, University of Edinburgh Royal Observatory, Blackford Hill, Edinburgh, EH9 3HJ, UK \\
+$^{4}$Leiden Observatory, Leiden University, PO Box 9513, 2300 RA Leiden, The Netherlands \\
+$^{5}$INAF - Istituto di Radioastronomia, Via Gobetti 101, 40129 Bologna, Italy  \\
+$^{6}$Fakult\"{a}t f\"{u}r Physik, Universit\"{a}t Bielefeld, Postfach 100131, 33501 Bielefeld, Germany \\
+$^{7}$Centre for Astrophysics Research, University of Hertfordshire, College Lane, Hatfield AL10 9AB, UK \\
+$^{8}$ASTRON, Netherlands Institute for Radio Astronomy, Oude Hoogeveensedijk 4, 7991 PD, Dwingeloo, the Netherlands \\}
 
-% Begin!
 \begin{document}
@@
 
-% Abstract with filler text
 \begin{abstract}
@@
 Quantifying their relative importance in a statistical way can be done using radio luminosity functions.
-Until now these have relied on galaxy classifications, with sources which have a mixture of radio emission from SF and AGN assigned to either a star-forming galaxy or an AGN.
+Until now these have relied on galaxy classifications, where sources with a mixture of radio emission from SF and AGN are labelled as either a star-forming galaxy or an AGN.
 This can lead to the under- or over-estimation of the relevance of AGN.
 Brightness temperature methods were demonstrated to be effective at separating radio emission from AGN and SF at 144 MHz using sub-arcsecond imaging with the International LOFAR Telescope (ILT).
-We use the combination of sub-arcsecond and arcsecond resolution imaging of 7,497 sources in the Lockman Hole and ELAIS-N1 fields to identify AGN components in the high resolution images and subtract them from the total flux density, leaving behind flux density from SF only.
-We construct, for the first time, radio luminosity functions by physical process rather than galaxy classification, revealing a hidden AGN population at \llof $<10^{24}$\wphz .
+We use the combination of sub-arcsecond and arcsecond resolution imaging of 7,497 sources in the Lockman Hole and ELAIS-N1 fields to identify AGN components in the high resolution images and subtract them from the total flux density, leaving flux density from SF only.
+We construct, for the first time, radio luminosity functions by physical process, either SF or AGN activity, revealing a hidden AGN population at \llof $<10^{24}$\wphz .
 This population is 1.56$\pm$0.06 more than expected for $0.5<z<2.0$ when comparing to RLFs by galaxy classification.
 The star forming population is less impacted, with only 0.90$\pm$0.02 of the expected star formation.
-The contribution of `hidden' AGN to the overall radio population, especially at \llof $<10^{24}$\wphz\ can have significant implications for the cosmic star formation rate and kinetic luminosity densities.
-This is only revealed thanks to the combination of the ILT's resolution and wide field of view.
-\vspace{3.7in}
+The contribution of `hidden' AGN to the overall radio population, especially at \llof $<10^{24}$\wphz , can have significant implications for the cosmic star formation rate and kinetic luminosity densities.
 \end{abstract}
 
-% Main body with filler text
+\begin{keywords}
+galaxies: active -- galaxies: star formation -- galaxies: evolution -- radio continuum: galaxies 
+\vspace{3.7in}
+\end{keywords}
+
 \section{Introduction}
@@
 The extent to which active galactic nuclei (AGN) impact galaxy formation is a major open question in astrophysics.
-It is clear that they influence each other, from both observational and theoretical standpoints.
+It is clear that there is an influence, from both observational and theoretical standpoints.
 Observations have revealed tight scaling relations between the mass of a super-massive black hole and its host galaxy \citep[see, e.g.][and references therein]{kormendy_coevolution_2013}, while cosmological simulations require some form of AGN feedback \citep{bower_breaking_2006,croton_many_2006} to be able to reproduce the observed galaxy population.
-%Galaxies build up their stellar mass through star formation, while super-massive black holes are expected to provide either negative or positive feedback to either suppress or stimulate galaxy growth, although whether we can definitively pinpoint which type of feedback is an open question \citep{ward_cosmological_2022}.
 The cosmic histories of AGN activity and star formation (SF), and by extension the interplay between them, are interlinked.
@@
 Efforts have been made to do this via mid-infrared (MIR) spectral decomposition using template fitting \citep[e.g.,][]{laurent_mid-infrared_2000,hernan-caballero_resolving_2015,li_active_2024}.
-Multi-wavelength approaches use spectral energy distribution (SED) fitting, which uses information from a much broader number of observed bands to jointly fit galaxy and AGN components \citep[e.g.][]{calistro_rivera_agnfitter_2016,boquien_cigale_2019,pacifici_art_2023}.
-However, SED fitting is costly as it relies on observations using many different instruments over the same part of the sky, which then have to be carefully cross-matched, and there can be degeneracies in the fits.
+Multi-wavelength approaches use spectral energy distribution (SED) fitting, which uses information from as many observed bands as possible to jointly fit galaxy and AGN components \citep[e.g.][]{calistro_rivera_agnfitter_2016,boquien_cigale_2019,pacifici_art_2023}.
+However, SED fitting is costly as it relies on observations from a variety instruments, which then have to be carefully cross-matched, and there can be degeneracy in the fits.
 Varying selection effects across the different bands can also present a problem.
@@
 Observations of radio emission have the advantage of being unobscured by dust or gas, so there are minimal observational biases.
-%Brightness temperature is a non-physical temperature which is defined as the temperature of a blackbody which would produce the observed surface brightness.
 Models of radio emission from SF predict an upper limit to the surface brightness that can be produced, even in starburst galaxies \citep{condon_radio_1992}.
 Surface brightness which is detected above this limit therefore has to be attributed to AGN activity.
-Brightness temperature is inversely proportional to frequency and resolution, $T_b \propto \nu^{-2}\Omega^{-1}$ (where $\Omega \propto \theta_{\text{maj}}\theta_{\text{min}}$).
-Traditionally brightness temperature measurements require Very Long Baseline Interferometry (VLBI) which reach milli-arcsec resolutions at GHz frequencies \citep[e.g.,][]{middelberg_mosaiced_2013,herrera_ruiz_faint_2017,radcliffe_nowhere_2018} to achieve enough sensitivity to make meaningful measurements.
-In \cite[][hereafter, \citetalias{morabito_identifying_2022}]{morabito_identifying_2022} we demonstrated that this can be done using sub-arcsecond resolution (0.\sarc\ 3) at 144 MHz with the LOw Frequency ARray \citep[LOFAR;][]{van_haarlem_lofar:_2013}.
+Brightness temperature is inversely proportional to frequency, $\nu$, and resolution, $\theta$: $T_b \propto \nu^{-2}\theta^{-2}$.
+Traditionally brightness temperature measurements require Very Long Baseline Interferometry (VLBI) which reach milli-arcsec resolutions at GHz frequencies \citep[e.g.,][]{middelberg_mosaiced_2013,herrera_ruiz_faint_2017,radcliffe_nowhere_2018} to achieve enough $T_b$ sensitivity to make meaningful measurements.
+In \citet[][hereafter, \citetalias{morabito_identifying_2022}]{morabito_identifying_2022} we demonstrated that this can be done using sub-arcsecond resolution (0.\sarc\ 3) at 144 MHz with the LOw Frequency ARray \citep[LOFAR;][]{van_haarlem_lofar:_2013}.
 The advantage of LOFAR is its wide field of view: $\sim$6 deg$^2$ from a single observation versus a few to a couple hundred arcmin$^2$ with VLBI at GHz frequencies, depending on whether one images a single phase centre or uses multi-phase centres for correlation \citep{deller_difx-2:_2011,morgan_vlbi_2011}.
-For example, using an 8 hour LOFAR observation of the Lockman Hole \citep{sweijen_deep_2022} we identified 940 AGN via brightness temperature \citepalias{morabito_identifying_2022}, providing a sample two orders of magnitude larger than any VLBI sample of AGN identified at GHz frequencies, using a fraction of the observing time.
+For example, \citetalias{morabito_identifying_2022} used an 8 hour LOFAR observation of the Lockman Hole \citep{sweijen_deep_2022} to identify 940 AGN via brightness temperature, providing a sample two orders of magnitude larger than any VLBI sample of AGN identified at GHz frequencies, using a fraction of the observing time.
 
@@
 They also demonstrate the flexible resolution which can be achieved with LOFAR, by imaging at 0.\sarc\ 6 and 1.\sarc\ 2 resolution, which are both complementary to the existing 6\sarc\ \ Deep Fields Data Release 1 \citep[DR1;][]{sabater_lofar_2021,tasse_lofar_2021}.
-This is another advantage over VLBI with instruments at GHz frequencies, which require either observing at different frequency bands or with other instruments to acquire information on the flux density of sources at other scales.
+This is another advantage over VLBI instruments operating at GHz frequencies, which require either observing at different frequency bands or with other instruments to acquire information on the flux density of sources at other scales.
 As described in \S~5 of \citetalias{morabito_identifying_2022}, for sources which are unresolved in the 0.\sarc\ 3 resolution images and do not exhibit a radio excess over that predicted by the star formation rate (SFR) derived from SED fitting \citepalias[see][for more details]{best_lofar_2023}, we can separate the AGN activity from SF using a combination of the 0.\sarc\ 3 and 6\sarc\ \ resolution images.
 
-The 2,483 sources in Lockman Hole \citep{sweijen_deep_2022} and now the 13,058 sources in ELAIS-N1 \citep{de_jong_into_2024}, represent a massive step towards robust statistical studies where the population can be broken into bins of redshift and stellar mass to investigate the AGN and SFG populations in detail.
-\cite{cochrane_lofar_2023} and \cite{kondapally_cosmic_2022} constructed radio luminosity functions (RLFs) from the DR1 and SED  for the star forming galaxies and radio-excess AGN, respectively.
+The 2,483 sources in Lockman Hole \citep{sweijen_deep_2022} and now the 13,058 sources in ELAIS-N1 \citep{de_jong_into_2024} represent a massive step towards robust statistical studies where the population can be broken into bins of redshift and stellar mass to investigate AGN and SFG populations in detail.
+\cite{cochrane_lofar_2023} and \cite{kondapally_cosmic_2022} constructed radio luminosity functions (RLFs) from DR1 for the star forming galaxies and radio-excess AGN, respectively.
 Here we construct RLFs, for the first time, by physical process rather than overall galaxy classification.
-%We do this by using a combination of the 0.\sarc\ 3 and 6\sarc\ \ resolution imaging from Lockman Hole and ELAIS-N1.
 We investigate the cosmic evolution of AGN activity and SF by dividing the sample into redshift bins out to $z = 2.5$.
 
-In this paper, we first describe the data in \S~\ref{sec:data}, followed by an overview of the methods in \S~\ref{sec:methods}. Section~\ref{sec:results} describes the results, followed by 
-% discussion and 
-conclusions in 
-%\S~\ref{sec:discussion} and
-\S~\ref{sec:conclusions}
-%, respectively
-.
-
+In this paper, we first describe the data in \S~\ref{sec:data}, followed by an overview of the methods in \S~\ref{sec:methods}.
+Section~\ref{sec:results} describes the results, followed by conclusions in \S~\ref{sec:conclusions}.
 Throughout this paper, we use a $H_0 = 70$ km$\,$s$^{-1}\,$Mpc$^{-1}$, $\Omega_M = 0.3$, and $\Omega_{\Lambda}= 0.7$ cosmology.
@@
 \label{subsec:dr1}
-The LOFAR Two-metre Sky Survey (LoTSS) has both a wide-area \citep{shimwell_lofar_2017,shimwell_lofar_2021,shimwell_lofar_2022} and Deep Fields \citep{tasse_lofar_2021,sabater_lofar_2021} component.
+The LOFAR Two-metre Sky Survey (LoTSS) has both a wide-area \citep{shimwell_lofar_2017,shimwell_lofar_2019,shimwell_lofar_2022} and Deep Fields \citep{tasse_lofar_2021,sabater_lofar_2021} component.
 The Deep Fields Data Release 1 (DR1) included Boötes, Lockman Hole, and ELAIS-N1, reaching sensitivities of 30, 23, and 20 \muJybm , using $\sim$80, $\sim$100, and $\sim$160 hours of observations, respectively.
 The resolution of these images is all 6\sarc .
-Alongside these deep radio images and their associated catalogues, \cite{kondapally_lofar_2021} provided a careful compilation of the ancillary deep data at other wavebands (from far-infrared to ultraviolet) and cross-matching to the radio catalogues, followed by photometric redshifts and stellar mass estimates in \cite{duncan_lofar_2021}.
+Alongside these deep radio images and their associated catalogues, \cite{kondapally_lofar_2021} provided a careful compilation of the ancillary data available at other wavebands (from far-infrared to ultraviolet) and cross-matching to the radio catalogues, followed by photometric redshifts in \cite{duncan_lofar_2021}.
 Collectively these catalogues provide an excellent resource, and we refer the reader for more details to the individual papers cited above.
@@
 All sources detected above 5$\sigma$ in their respective radio images had their far-infrared (FIR) to ultraviolet (UV) data fit using four different SED fitting software packages, and a `consensus' value was found for the derived galaxy properties using information from the goodness-of-fits.
-%These were: MAGPHYS \citep{da_cunha_simple_2008}, BAGPIPES \citep{carnall_inferring_2018,carnall_how_2019}, CIGALE \citep{burgarella_star_2005,noll_analysis_2009,boquien_cigale_2019}, and AGN\textsc{fitter} \citep{calistro_rivera_agnfitter_2016}.
-The first two (MAGPHYS and BAGPIPES) use an energy balance approach, where optical to UV energy absorbed by dust must be transferred to thermal emission in the FIR to sub-mm.
-CIGALE also uses the same energy-balance approach, but includes an AGN component.
-The final code, AGN\textsc{fitter} is designed to work in cases where the UV light and IR emission are spatially distinct, and thus the energy balance assumption does not hold.
-Many AGN fall under this umbrella.
-Each galaxy was fit with all four codes (twice for CIGALE, using different AGN components) and a `consensus' value found for the derived galaxy properties using information from the goodness-of-fits.
 
@@
     \item Radiatively efficient + not radio excess (\textit{`radio-quiet'})
-    \item Radiatively inefficient + not radio excess
+    \item Not radiatively efficient + not radio excess
 \end{itemize}
@@
 Five percent of all sources are unclassified, as they do not have a radio excess and the fits could not determine between radiatively efficient or inefficient.
+We note here that more recent work generated probabilistic classifications \citep{drake_lofar_2024} would have an impact on the RLFs, although the extent is a subject for future studies.
 
@@
 While LoTSS uses only the stations in the Netherlands for its standard data processing, post-processing of LoTSS data to achieve sub-arcsecond resolution is ongoing both for individual sources with $S>10\,$mJy in the wide-area survey \citep{morabito_sub-arcsecond_2022} and for the entire field of view in the Deep Fields \citep[][Escott et al., in prep; Bondi et al., in prep]{sweijen_deep_2022,de_jong_into_2024}.
-%While the full field of view imaging is currently too computationally expensive for the wide area, we are focusing on the Deep Fields.
 Due to the fact that the size of international stations is larger than that of stations in the Netherlands, the station beams have a smaller field of view.
@@
 Both Lockman Hole and ELAIS-N1 have been imaged at 0.\sarc\ 3 resolution \citep[][respectively]{sweijen_deep_2022,de_jong_into_2024}.
-The resulting images have median rms noise of 34 \muJybm\ (Lockman Hole) and 17 \muJybm\ (ELAIS-N1), although the noise is lowest in the centre and increases radially in both images.
-%(and can vary per facet).
+The resulting images have median rms noise of 34 \muJybm\ (Lockman Hole) and 17 \muJybm\ (ELAIS-N1), with the noise being lowest in the centre and increasing radially in both images.
+Although \cite{de_jong_into_2024} also imaged ELAIS-N1 at 0.\sarc\ 6 and 1.\sarc\ 2, for consistency with Lockman Hole we use only the 0.\sarc\ 3 resolution image.
 Catalogues from these images were processed to group individual Gaussian islands into complete sources, and remove duplicates.
 The resulting catalogues contain 2,316 and 13,058 sources above 5$\sigma$, for Lockman Hole and ELAIS-N1 respectively.
-Although \cite{de_jong_into_2024} also imaged ELAIS-N1 at 0.\sarc\ 6 and 1.\sarc\ 2, for consistency with Lockman Hole we use only the 0.\sarc\ 3 and 6\sarc\ resolution images.
 
-%The Lockman Hole was the first field to have its full field of view imaged at 0.\sarc\ 3 resolution \citep{sweijen_deep_2022}.
-This took a quarter million cpu hours to produce a single 7-billion pixel image from 8 hours of data.
-%The resulting image has a median rms noise of 34 \muJybm , although this is lower in the centre and higher in the outskirts.
-From this image, a catalogue of 2,316 5$\sigma$ sources was created, removing duplicates and grouping individual Gaussian islands together into sources.
-%The first 0.\sarc\ 3 image of ELAIS-N1 was recently published \citep{de_jong_into_2024}, along with companion images at intermediate resolutions of 0.\sarc\ 6 and 1.2\sarc .
-The median rms noise reaches 14 \muJybm\ in the centre of the 0.\sarc\ 3 resolution image, and increases radially.
-For consistency with the Lockman Hole sample, we do not use the intermediate resolution images here.
-In the 0.\sarc\ 3 resolution image, there were 13,058 sources catalogued above 5$\sigma$ after grouping Gaussian islands into sources and removing duplicates.
-
-Thanks to the added collecting area of the international stations, which increases the sensitivity of the array, the 0.\sarc\ 3 resolution image for ELAIS-N1 has a lower noise level than the Deep Fields DR1.
+Thanks to the added collecting area of the international stations, which increases the array's sensitivity, the 0.\sarc\ 3 resolution image for ELAIS-N1 has a lower noise level than the Deep Fields DR1.
 To ensure consistently built samples, and make use of DR1 information, we limit the high-resolution catalogues.
-For both fields,
-%The derived galaxy properties from \citetalias{best_lofar_2023} (see \S~\ref{subsec:dr1}) were for LoTSS Deep Fields DR1, and thus we have to limit the source catalogues we wish to use.
-For both Lockman Hole and ELAIS-N1 catalogues derived from the 0.\sarc\ 3 resolution images,
-we first find the local rms in the companion 6\sarc\ \ resolution noise maps at the position of each source in the 0.\sarc\ 3 resolution catalogue.
+For both fields, we first find the local rms in the companion 6\sarc\ \ resolution noise maps at the position of each source in the 0.\sarc\ 3 resolution catalogue.
 Using the peak brightness of the source in the catalogue from the 0.\sarc\ 3 resolution image, we calculate the peak-to-noise ratio and only select sources which would be detectable at the 5$\sigma$ limit in the 6\sarc\ resolution image, effectively ensuring all sources should have a match in DR1.
 We then cross-match based on radio position, to provide the radio information from the catalogues derived from both the 0.\sarc\ 3 resolution image and the 6\sarc\ \ resolution image.
-We use optical star masks\textsuperscript{\ref{dr1}} to remove sources with multi-wavelength information affected by bright stars.
-Finally, we trim the area down of the multi-wavelength catalogue to the 2.5 deg$\times$2.5 deg area which is covered by the 0.\sarc\ 3 resolution image (although the multi-wavelength coverage in ELAIS is only 63 percent of the multi-wavelength coverage of Lockman Hole).
+We use optical star masks to remove sources with multi-wavelength information affected by bright stars.
+Finally, we trim the area of the multi-wavelength catalogue to the 2.5 deg$\times$2.5 deg area which is covered by the 0.\sarc\ 3 resolution image (although the multi-wavelength coverage is smaller in ELAIS-N1 than in Lockman Hole).
 We are left with \variable{output/lockman_detectable.txt}sources in Lockman Hole, and \variable{output/en1_detectable.txt}sources in ELAIS-N1.
-%We use the source name from the catalogue derived from the 6\sarc\ \ resolution to cross-match to the catalogue from \citetalias{best_lofar_2023}.
 
-We determine whether sources are resolved or unresolved (relevant in the next section) in the same manner as \S~3.1 of \cite{shimwell_lofar_2019,shimwell_lofar_2022}, by finding an upper limit to the intrinsically unresolved sources.
+We determine whether sources are resolved or unresolved (relevant in Section~\ref{sec:methods}) in the same manner as \S~3.1 of \cite{shimwell_lofar_2019,shimwell_lofar_2022}, by finding an upper limit to the intrinsically unresolved sources.
 Theoretically, these should have equal peak brightness and total flux density, but LOFAR maps are impacted by smearing due to the ionosphere, and high resolution images have time and bandwidth smearing effects.
 Thus intrinsically unresolved sources can appear spatially extended.
-To find this upper limit, we take all sources which are fit by a single Guassian component ({\ttfamily S\_CODE = `S'}), and are smaller than a multiple of the beam size.
+To find this upper limit, we take all sources which are fit by a single Guassian component and are smaller than a multiple of the beam size.
 For the Lockman Hole, we use a factor of 3, but for ELAIS-N1 we find that a factor of 4 provided a better sample.
 This may be because the ELAIS-N1 field uses multiple nights, which have varying ionospheric conditions \citep[see \S~2 in][]{de_jong_into_2024}.
-%This may be because half of the data for ELAIS-N1 had been pre-averaged to 2 sec integration times, which will increase the time smearing \citep[see \S~2 in][]{de_jong_into_2024}. 
-Plotting the total flux density to peak brightness ratio of this sample as a function of the local signal to noise, we fit a Sigmoid function to the upper 99.9th percentile.
-Applying this cut to the entire catalogue separates sources into unresolved or resolved.
-Any source which does not have {\ttfamily S\_CODE = `S'} is also considered resolved.
-%We fit a Sigmoid function to find the upper 99.9th percentile of sources considering their total flux density to peak brightness ratio as a function of the local signal to noise.
-Anything above this boundary is considered to be resolved.
-We also treat sources which do not have {\ttfamily S\_CODE = `S'} as resolved.
-
-\begin{comment}
-\subsection{Simulations}
-\label{subsec:simba}
-
-We use the SIMBA cosmological hydrodynamic simulation which accounts for an H$_{2}$ based star formation rate following the sub-grid model of \cite{krumholz_comparison_2011}.
-Additionally, SIMBA models the growth of supermassive black holes via a two-mode prescription based on the temperature of gas within a distance, $R_{0}$, calculated over the 256 nearest particles limited to 2$h^{-1}$kpc.
-It is commonplace for cosmological simulations to employ the growth of SMBHs via the Bondi rate \hl{citations EAGLE, HORIZON AGN, Illustris, TNG}.
-In SIMBA, the Bondi rate is employed when gas within $R_{0}$ has temperatures T$>10^{5}$K, referred to as the hot mode.
-What makes SIMBA unique is the additional contribution to the accretion rate via a torque limited accretion flow when gas has temperatures T$<10^{5}$K, or in the cold mode.
-SIMBA also employs a physically motivated model for AGN feedback in the form of AGN winds \citep{perna_x-raysdss_2017}, as well as AGN jets with velocities that increase inversely proportional to the Eddington rate in accordance with \cite{heckman_coevolution_2014}.
-SIMBA therefore provides a platform to study jet-mode AGN in a way that has not been explored before. 
-Specifically, we select galaxies in SIMBA that have full-velocity jets, that is, when the Eddington rate drops below 0.02 (but non-zero) and where the black hole mass is larger than M$_{\rm BH}>10^{8}$M$_{\odot}$.
-For this population we are able to calculate the amount of radio emission due to star formation processes following \cite{condon_radio_1992} as well as that from accretion processes following \cite{kording_measuring_2008}.
-Details of this method can be found in \cite{thomas_radio_2021}.
-We can therefore calculate the contribution to the radio luminosity function from star formation and AGN accretion processes.
-We can additionally separate the galaxy population into star forming and quiescent galaxies, radio-excess, or radiatively efficient and inefficient, based on the sources black hole and host properties.
-\end{comment}
+A Sigmoid function is fit to the upper 99.9th perecentile of the total flux density to peak brightness ratio as a function of local signal to noise.
+All sources in the catalogue which lie above this function are treated as resolved, as is any source not fit by a single Gaussian component.
 
@@
 \label{sec:methods}
-
 Brightness temperature, $T_b$, is used to characterise the surface brightness of radio emission.
@@
 By creating a model which predicts the upper limit of $T_b$ achievable by SF \citep[based on][]{condon_radio_1992}, if the value is \textit{above} that upper limit, the radio emission must be due to AGN activity.
-\citetalias{morabito_identifying_2022} showed that this AGN identification can be done at 150 MHz using sub-arcsecond resolution;
-%this works because $T_b\propto\nu^{-2}\Omega^{-1}$ where $\nu$ is the observed frequency and $\Omega$ is the solid angle of the emitting region -- for unresolved sources, this corresponds to the deconvolved beam size.
-we refer the reader to \S~3 of \citetalias{morabito_identifying_2022} for more details.
+\citetalias{morabito_identifying_2022} showed that $T_b$-based AGN identification can be done at 150 MHz using sub-arcsecond resolution.
+We make the same conservative assumptions as \citetalias{morabito_identifying_2022}, i.e.
+we set $T_e=10^4\,$K and the frequency where the optical depth reaches unity, $\nu_0$, to be 3 GHz, for the star formation models.
+These equate to $T_b\sim10^6\,$K at 144 GHz (equivalent to $T_b\sim10^5\,$K at 1.4 GHz) as the upper limit for SF.
+The models directly predict flux density per solid angle, which can then be converted to $T_b$ values.
+Practically, we follow the method in \citetalias{morabito_identifying_2022} which simply uses a threshold in terms of flux density per solid angle rather than converting to $T_b$.
+We refer the reader to \S~3 of \citetalias{morabito_identifying_2022} for more details.
 Note that this method yields conservative, positive AGN identifications.
-There may be lower surface brightness AGN emission which is not identified, and it is also possible \citep[although unlikely, according to][]{condon_radio_1992} that extreme, compact nuclear starbursts could produce AGN-like values of $T_b$.
+
+There are two intermediate cases which may not be caught by this.
+For the first case take the example of Arp 220, a late-stage merger that has been shown to have a high brightness temperature which is consistent with an origin entirely from densely packed, very luminous radio supernovae \citep{smith_starburst_1998,lonsdale_vlbi_2006}.
+Sub-arcsecond resolution imaging with the ILT largely supports this \citep{varenius_subarcsecond_2016} but cannot rule out an AGN origin for the outflow in the eastern nucleus.
+In this extreme case, very high SF in a compact region can mimic AGN-like values of $T_b$.
+We note that as the models of SF predict \textit{surface brightness}, compactness of SF relative to the resolution element, which in this case is larger than for GHz VLBI, would actually \textit{decrease} the surface brightness as the measured size would be larger than the intrinsic size.
+To best imitate an AGN, either the SFR has to be extreme and/or the SF has to take place in an area which is approximately the size of the beam.
+
+The second scenario is low-surface brightness AGN emission which mimics SF.
+This could be due, for example, to restarted radio AGN with a young, prominent core and low surface-brightness extended structures from previous AGN emission \citep[e.g.][]{mingo_revisiting_2019}.
+Or it could be due to radio emission powered by shocks from AGN winds \citep[e.g.,][]{zakamska_quasar_2014,rankine_placing_2021,petley_connecting_2022}.
+In either case the radio emission due to SF will be over-estimated, as it is contaminated by low-surface brightness AGN emission.
+Based on the number of new $T_b$ identifications in \citetalias{morabito_identifying_2022}, we expect this is $\lesssim$10 percent of cases.
+
+These two scenarios act in the opposite direction, and may balance each other out in this study.
+We aim to provide better constraints based on source structure (e.g., compact starbursts) and using a wider range of spatial scales in our ILT imaging \citep[e.g.
+the 1.\sarc\ 2 resolution imaging in][]{de_jong_into_2024} in future work.
 
 \subsection{Separation of AGN activity and SF}
-Once the AGN activity is identified from the 0.\sarc\ 3 resolution images, it is possible to separate it from SF using the total flux density from the 6\sarc\ \ resolution images.
+Once the AGN activity is identified from the 0.\sarc\ 3 resolution images, it is possible to separate it from SF using the total flux density in the 6\sarc\ \ resolution images.
 We refer the reader to \S~5 in \citetalias{morabito_identifying_2022} for more details but outline the process here, which is depicted in Fig.~\ref{fig:flowchart}.
@@
 \subsection{Construction of RLFs}
+We follow the same method of constructing RLFs as in \cite{kondapally_cosmic_2022} and \cite{cochrane_lofar_2023}, which use the standard 1/$V_{\mathrm{max}}$ method \citep{schmidt_space_1968,condon_14_1989}.
+The 0.\sarc\ 3 resolution image is smaller than the 6\sarc\ resolution image, so we first reproduce the published RLFs (Fig.~\ref{fig:rlfs}, left panel) using the galaxy classifications from \citetalias{best_lofar_2023} to check that the RLFs in the smaller area are consistent.
+We use the completeness corrections from the respective papers, with the exception of the photo-$z$ correction used in \cite{cochrane_lofar_2023}.
+The authors note that this correction is small, and as it is not implemented in \cite{kondapally_cosmic_2022} we chose to keep the method in this paper consistent across both samples.
+We use the same redshift range, $0.003 < z < 0.3$, to verify that we have good agreement with the published RLFs.
+Uncertainties are bootstrapped as in \cite{kondapally_cosmic_2022}, with random sampling by replacement to generate 1000 RLF realisations from which the 1$\sigma$ uncertainties are determined.
 
-We follow the same method of constructing RLFs as in \cite{kondapally_cosmic_2022} and \cite{cochrane_lofar_2023}.
-The 0.\sarc\ 3 resolution image is smaller than the 6\sarc\ resolution image, so we first reproduce the published RLFs (Fig.~\ref{fig:rlfs}, left panel) using the galaxy classifications from \citetalias{best_lofar_2023} to check that we have implemented the method appropriately, and that the smaller area is consistent.
-We use the completeness corrections from the respective papers.
-The only correction we do not implement is the photo-$z$ correction used in \cite{cochrane_lofar_2023}.
-The authors note that this correction is small, and as it is not implemented in \cite{kondapally_cosmic_2022} we chose to keep the method for RLF construction in this paper consistent across both samples.
-We use the same redshift range, $0.003 < z < 0.3$, to verify that we have good agreement with the published RLFs.
-
-Next we use the same method for constructing RLFs by physical process rather than overall galaxy classification.
-We again use the completeness corrections from \cite{cochrane_lofar_2023} and \cite{kondapally_cosmic_2022}.
-The completeness corrections are flux density dependent, and we use the overall total flux density of a galaxy, rather than the flux density of the process, which will be smaller if the radio emission from SF and AGN have been split.
-This is because the completeness correction is based on whether or not a \textit{source} will be detectable given the background rms, not any individual component.
-%This only relies on the total flux density, not the flux density of individual components. 
-%Even when considering only the SF (or AGN) radio emission in a source, its overall detectability is still derived from the combination of all radio emission processes present. 
-These RLFs for the physical processes of SF and AGN are shown in the middle panel of Fig.~\ref{fig:rlfs}, again for  $0.003 < z < 0.3$.
-
-% Finally, we use RLFs constructed from SIMBA. The RLFs are constructed as in \cite{thomson_radio_2019}, but have been updated to use the same redshift bins \hl{Nicole}.
+Next we use the same 1/$V_{\mathrm{max}}$ method for constructing RLFs by physical process rather than overall galaxy classification, using the flux density of the relevant process, AGN activity or SF, rather than the overall total flux density (where they are not the same).
+We directly use the completeness corrections from \cite{cochrane_lofar_2023} and \cite{kondapally_cosmic_2022}, based on a galaxy's total flux density.
+These corrections are flux density dependent, but they are based on
+%we use the total flux density of a galaxy, rather than the flux density of the process, which will be smaller if the radio emission from SF and AGN have been split.
+This is because the completeness correction is based on
+whether or not a \textit{source} will be detectable given the background rms, not any individual component, although the component flux density can be smaller than the total flux density.
 
@@
     \includegraphics[width=0.98\textwidth]{figures/deep_fields_RLFs.png}
-    \caption{\textit{Left:} Validation of our re-calculated galaxy RLFs (dotted lines) for the smaller area considered here with the previously published RLFs (solid lines).
+    \caption{\textit{Left:} The re-calculated galaxy RLFs (dotted lines) for the smaller area considered here, compared with previously published RLFs (solid lines).
 \textit{Middle:} RLFs calculated by process rather than galaxy.
@@
 \textit{Bottom right:} $\Delta$RLF for both AGN and SF.
-}
+Hatched regions show where the uncertainties are large and values should be treated with caution.}
     \label{fig:rlfs}
@@
 \label{sec:results}
-
-We present the very first RLFs constructed by physical process rather than galaxy classification.
-Comparing these with our reproduction of the previously published RLFs, using galaxy classifications, allows us to explore where we are over- or under-predicting the contribution of the physical processes to our statistical studies.
+The RLFs for the physical processes of SF and AGN are shown in the middle panel of Fig.~\ref{fig:rlfs}, for  $0.003 < z < 0.3$. 
+%We present the first RLFs constructed by physical process rather than galaxy classification in Fig~\ref{fig:rlfs}. 
+Comparing these with RLFs using galaxy classifications allows us to explore where we are over- or under-predicting the contribution of the physical processes to our statistical studies.
 To aid this comparison, we define and calculate:
@@
 \end{equation}
-The $\Delta$RLF is calculated once for the SF category/SFG classification, and once for the AGN category/AGN classification.
+$\Delta$RLF is calculated once for the SF process/SFG classification, and once for the AGN process/AGN classification.
 We show $\Delta$RLF in the bottom right hand panel in Fig.~\ref{fig:rlfs}.
-We see that above 10$^{24}\,$\wphz\ the ratio for AGN is unity, which is expected as this is where radio excess sources dominate the population.
+Above 10$^{24}\,$\wphz\ the ratio for AGN is unity, which is expected as this is where radio excess sources dominate the population.
 Below that, the ratio for AGN climbs up to $\sim$1.3 and finally almost up to 2 at the lowest radio powers probed.
@@
 We reproduce previous observational results that while the SF population has a strong redshift evolution, the AGN population does not.
-%For the simulations, the behaviour is qualitatively similar, although the AGN population does have mild redshift evolution whereas the observations do not.
 
-The bottom panels in Fig~\ref{fig:evolution} show $\Delta$RLF, the same as in the bottom right hand panel of Fig~\ref{fig:rlfs}.
+The bottom panels in Fig.~\ref{fig:evolution} show $\Delta$RLF, the same as in the bottom right hand panel of Fig~\ref{fig:rlfs}.
 For the SF population, $\Delta$RLF is close to unity for $L_{\textrm{144MHz}} < 10^{24}\,$\wphz\ in the lowest redshift bin, and then this begins to drop at higher luminosities.
 The higher redshift bins show similar behaviour, perhaps reaching unity at increasingly higher values of $L_{\textrm{144MHz}}$, although deeper data is needed to confirm this.
-%The lowest values reached are $\sim$50 percent.
-%Comparing these ratios between observations and simulations, we see qualitatively the same behaviour, although the simulations reach much lower and higher values for both the SF and AGN populations, respectively.
-Looking first at the SF population (left set of panels), in the lowest redshift bin the ratio for the simulation stabilises to around unity for \llof $<10^{23}\,$\wphz , and decreases rapidly above this, reaching much smaller values than the observations.
-Higher redshift bins probe successively higher luminosity ranges, also reaching small values for the ratio between activity and galaxies.
-This is consistent with the fact that at higher redshifts, most radio AGN in SIMBA are hosted in SFGs, and thus the \textit{galaxy} RLF for SIMBA SFGs is inflated with AGN emission; this makes the ratio of SF activity to galaxy approach zero.
 
-For the AGN population, $\Delta$RLF has almost the opposite behaviour, converging to unity \textit{above} $\sim$\llof $10^{24}\,$\wphz\ for the lowest redshift bin.
-This is driven by the flux density of radio excess sources being entirely attributed to AGN, which tend to have higher radio luminosities.
-The convergence happens at higher luminosity values for higher redshifts.
+For the AGN population, $\Delta$RLF has almost the opposite behaviour, converging to unity \textit{above} \llof $\sim10^{24}\,$\wphz\ for $0.003 < z < 0.3$.
+This is driven by the flux density of radio excess sources, which tend to have higher radio luminosity, being entirely attributed to AGN.
+The convergence happens at higher luminosity for higher redshift.
 Below this convergence, $\Delta$RLF increases across all redshift bins.
-The maximum value for each redshift bin approaches $\sim$2, although the lowest redshift bin the value stabilises at $\sim1.3$ and only climbs higher at the very lowest luminosities (\llof $<10^{22}\,$\wphz ), which may be due to incompleteness.
-%If we disregard the lowest luminosity bins, it seems that we see approximately a 30 - 50 percent increase in AGN activity over what is measured when using galaxy classifications alone, for all redshifts.
-%This is borne out by the values in the bottom line of Tab.~\ref{tab:intvalues}, where the difference in the integrated RLFs shows an excess of AGN contribution of $\sim$ 50 percent for the three middle redshift bins.
-
-%The ratio for the SIMBA RLFs shows strong redshift evolution, and much higher ratio values.
-The RLFs themselves turn over, whereas the observed RLFs do not.
-This is likely a consequence of the fact that the AGN emission in SIMBA is due to high-velocity jets, whereas there is increasing observational evidence for low-luminosity jets or AGN emission \citep{macfarlane_radio_2021,yue_novel_2024}.
-The high values reached by the ratio of RLFs again is expected as many high-$z$ radio AGN in SIMBA are hosted in SFGs.
-Shifting this AGN emission to AGN \textit{activity} when it is counted in the SFG \textit{galaxy} population generates high values for the activity to galaxy ratio.
-%- sfrd, kinetic LF
+The maximum value for each redshift bin approaches $\sim$2, although in the lowest redshift bin this stabilises at $\sim1.3$ and only climbs higher at the very lowest luminosities (\llof $<10^{22}\,$\wphz ), which may be due to incompleteness.
 
@@
 The dashed horizontal line in the bottom panels is unity, with thin solid horizontal lines at 0.5 (left) at 2.0 (right) to guide the eye.
+To avoid overcrowding, uncertainties are only plotted for $\Delta$RLF, in the same manner as Fig.~\ref{fig:rlfs}.
 }
@@
 For example, for SF we integrate the RLF for the SF process and the RLF for SFGs, and divide the first by the second.
-These numbers are always less than unity since the SF process is over-estimated in the SFG population, and this has been corrected by moving $T_b$-identified AGN flux density contributions to the AGN process RLF.
-These are reported in Tab.~\ref{tab:intvalues}.
+These numbers are always less than unity since the SF process is over-estimated in the SFG population, which has been corrected by moving $T_b$-identified AGN flux density contributions to the AGN process RLF.
+The integrated values are reported in Tab.~\ref{tab:intvalues}.
 \variable{output/average_integrated_differences.txt} 
-%, and show that while the SF process only decreases by $\sim$10 percent in the lowest redshift bin, the AGN contribution increases by almost a factor of 2. 
 It is evident from the bottom right panel of Fig.~\ref{fig:rlfs} that this `hidden' AGN population extracted from SFGs shows up in the radio AGN population at $L_{\textrm{144MHz}} < 10^{24}\,$\wphz .
@@
 \label{sec:conclusions}
-
-In this letter we have presented, for the first time, RLFs constructed by separating radio emission into either SF or AGN processes, for 7,497 sources.
+In this letter we presented, for the first time, RLFs constructed by separating radio emission into either SF or AGN processes, for 7,497 sources.
 We use brightness temperature to identify AGN components in galaxies with no radio excess, and separate the radio emission into AGN and SF.
-We compare these with classical RLFs constructed using overall galaxy classifications, and find a hidden AGN population in sources with \llof $\lesssim 10^{24}$\wphz .
-The integrated impact of this is approximately a 50 percent increase in the radio luminosity output due to AGN across intermediate redshifts ($0.5 < z < 2.0$), and a doubling in the lowest redshift bin ($0.003 < z < 0.3$).
-The impact on the star forming galaxy population is less severe, as this population is much larger, amounting to an integrated decrease in the RLF of $\sim$10-20 percent.
-These impacts can have implications for the cosmic SFR density and kinetic luminosity density, which will be the topic of a future paper.
+We compare these with classical RLFs constructed using galaxy classifications, and find a hidden AGN population in sources with \llof $\lesssim 10^{24}$\wphz .
+The mean integrated impact of this is a 1.52$\pm$0.06 increase in the radio luminosity output due to AGN across intermediate redshifts ($0.5 < z < 2.0$), and a doubling in the lowest redshift bin ($0.003 < z < 0.3$).
+The impact on the SFG population is less severe, amounting to a mean integrated decrease in the RLF to 0.90$\pm$0.02 .
+The results can have implications for the cosmic SFR density and kinetic luminosity density, which we will explore in a future paper.
 
@@
 The sub-arcsecond resolution is sufficient for brightness temperature identification of AGN, and the field of view means that we have been able to do this for 1,022 sources (out of $>$4,000).
-We have used conservative limits for AGN identification, which means that there is likely more AGN emission still remaining `uncounted' here.
-Future work to build a forward modelling method, that can also make use of the intermediate (0.\sarc\ 6 and 1.\sarc\ 2) resolution images will improve the AGN and SF separation.
-Splitting radio emission into category by physical process also opens up a pathway for more direct comparison to cosmological simulations, which we will address in a future paper.
-
-%\vspace{-0.5in}
-
-\section*{Data Availability}
-All data used in this paper is either publicly available on \href{https://lofar-surveys.org}{https://lofar-surveys.org} or from the relevant citations herein, including catalogues, rms images, and star masks.
-We have prepared this manuscript using using the showyourwork (\href{https://github.com/rodluger/showyourwork}{https://github.com/rodluger/showyourwork}) open source scientific article workflow, which was first introduced in \cite{luger_mapping_2021}.
-This ensures complete transparency.
-The code is available at \hl{update link when ready to publish}.
-
-\section*{Acknowledgements}
-LKM is grateful for support from UKRI [MR/T042842/1]. 
-FS appreciates the support of STFC [ST/Y004159/1]. 
-JMGHJdJ is grateful for support from the project CORTEX (NWA.1160.18.316) of the research programme NWA-ORC which is (partly) financed by the Dutch Research Council (NWO). 
-BY thanks the University of Edinburgh and Leiden Observatory for support through the Edinburgh–Leiden joint studentship.
-PNB is grateful for support from the UK STFC via grants [ST/V000594/1] and [ST/Y000951/1]
-RK acknowledges support from the UK STFC via grant [ST/V000594/1].
-RJvW acknowledges support from the ERC Starting Grant ClusterWeb 804208. 
-MB acknowledges support from INAF under the Large Grant 2022 funding scheme (project “MeerKAT and LOFAR Team up: a Unique Radio Window on Galaxy/AGN co-Evolution”).
-
-LOFAR is the Low Frequency Array designed and constructed by ASTRON.
-It has observing, data processing, and data storage facilities in several countries, which are owned by various parties (each with their own funding sources), and which are collectively operated by the ILT foundation under a joint scientific policy.
-The ILT resources have benefited from the following recent major funding sources: CNRS-INSU, Observatoire de Paris and Université d'Orléans, France; BMBF, MIWF-NRW, MPG, Germany; Science Foundation Ireland (SFI), Department of Business, Enterprise and Innovation (DBEI), Ireland; NWO, The Netherlands; The Science and Technology Facilities Council, UK; Ministry of Science and Higher Education, Poland; The Istituto Nazionale di Astrofisica (INAF), Italy.
-
-%This research made use of the Dutch national e-infrastructure with support of the SURF Cooperative (e-infra 180169 and EINF-6218) and the LOFAR e-infra group.
-
-%This work is co-funded by the EGI-ACE project (Horizon 2020) under Grant number 101017567. This work was sponsored by NWO Domain Science for the use of the national computer facilities.
-
-%The Jülich LOFAR Long Term Archive and the German LOFAR network are both coordinated and operated by the Jülich Supercomputing Centre (JSC), and computing resources on the supercomputer JUWELS at JSC were provided by the Gauss Centre for Supercomputing e.V.
-(grant CHTB00) through the John von Neumann Institute for Computing (NIC).
-This research made use of the University of Hertfordshire high-performance computing facility and the LOFAR-UK computing facility located at the University of Hertfordshire and supported by STFC [ST/P000096/1], and of the Italian LOFAR IT computing infrastructure supported and operated by INAF, and by the Physics Department of Turin university (under an agreement with Consorzio Interuniversitario per la Fisica Spaziale) at the C3S Supercomputing Centre, Italy.
-
-\begin{comment}
-\section*{Appendix A}
-
-\begin{figure*}
-    \includegraphics[width=0.49\textwidth]{figures/Lockman_Unresolved_03.png}
-    \hfill
-    \includegraphics[width=0.49\textwidth]{figures/ELAIS_Unresolved_03.png}
-    \caption{\label{fig:unres}The data and sigmoid fits to determine whether sources are resolved or unresolved as described in \S~\ref{subsec:highres}.
-The \textit{left} panel shows Lockman Hole, and the \textit{right} panel shows ELAIS-N1.
-}
-\end{figure*}
-
-Figure~\ref{fig:unres} shows the integrated flux density to peak brightness ratio as a function of the local signal to noise, parameterised by $\frac{S_{peak}}{2\sigma_{peak}} + \frac{S_{int}}{2\sigma_{int}}$.
-The data is first binned evenly in log space, and the 99.9th percentile is found in each bin.
-We then fit a sigmoid function:
-\begin{equation}
-y = A + \frac{B}{1 + (x / 96.57)^C}, 
-\end{equation}
-where 96.57 is a fixed value that has worked well for LoTSS \citep{shimwell_lofar_2022}. For Lockman Hole, we find 
-\end{comment}
+We have used conservative limits for AGN identification, which means that there is likely more AGN emission still remaining `uncounted'.
+Future work to build a forward modelling method, which will also use intermediate resolution images to improve the AGN and SF separation.
+Splitting radio emission into category by physical process also opens a pathway for direct comparison to cosmological simulations, which we will address in a future paper.
 
@@
 %\pagebreak
+
+\section*{Data Availability}
+All data used in this paper is either publicly available on \href{https://lofar-surveys.org}{https://lofar-surveys.org} or from relevant citations herein, including catalogues, rms images, and star masks.
+We have prepared this manuscript using using the showyourwork (\href{https://github.com/rodluger/showyourwork}{https://github.com/rodluger/showyourwork}) open source scientific article workflow, first introduced in \cite{luger_mapping_2021}.
+This ensures complete transparency.
+The code is available at \hl{update link when ready to publish}.
+\section*{Acknowledgements}
+LKM recognizes support from UKRI [MR/T042842/1]. 
+RK recognizes support from UK STFC via grant [ST/V000594/1].
+PNB recognizes support from UK STFC via grants [ST/V000594/1] and [ST/Y000951/1].
+BY thanks the University of Edinburgh and Leiden Observatory for support through the Edinburgh–Leiden joint studentship.
+JMGHJdJ recognizes support from project CORTEX (NWA.1160.18.316) of research programme NWA-ORC which is (partly) financed by the Dutch Research Council (NWO). 
+FS recognizes support from STFC [ST/Y004159/1]. 
+MB and IP recognize support from INAF under the Large Grant 2022 funding scheme (project “MeerKAT and LOFAR Team up: a Unique Radio Window on Galaxy/AGN co-Evolution”).
+DJBS recognizes support from UK STFC via grants [ST/V000624/1] and  [ST/Y001028/1].
+RJvW recognizes support from ERC Starting Grant ClusterWeb 804208. 
+This work was supported by SURF Cooperative grant EINF-6218 and EGI-ACE project (Horizon 2020 grant 101017567).
+
+The Low Frequency Array was designed and constructed by ASTRON.
+It has observing, data processing, and data storage facilities in several countries, which are owned by various parties (each with their own funding sources), and collectively operated by the ILT foundation under a joint scientific policy.
+The ILT resources have benefited from the following recent major funding sources: CNRS-INSU, Observatoire de Paris and Université d'Orléans, France; BMBF, MIWF-NRW, MPG, Germany; Science Foundation Ireland (SFI), Department of Business, Enterprise and Innovation (DBEI), Ireland; NWO, The Netherlands; The Science and Technology Facilities Council, UK; Ministry of Science and Higher Education, Poland; The Istituto Nazionale di Astrofisica (INAF), Italy.
 

### 2024-10-30 src/scripts/RLF_evolution_v2.py
removed obsolete

@@ -1,304 +1,0 @@
-import paths
-import numpy as np
-import matplotlib
-from matplotlib import pyplot as plt
-from astropy.table import Table, Column, join, vstack
-from helper_functions import *
-from astropy.io import fits
-## WMAP9 is Hinshaw et al. 2013, H_0=69.3, Omega=0.287
-
-##############################################################
-## Plotting housekeeping
-
-matplotlib.rcParams['legend.frameon'] = False
-matplotlib.rcParams['axes.labelsize'] = 'large'
-## set up some colours
-n = 255
-mycols = plt.cm.viridis(np.linspace(0, 1,n))
-simba_mycols = plt.cm.cividis(np.linspace(0,1,n))
-mycols_m = plt.cm.inferno(np.linspace(0, 1,n))
-
-##############################################################
-
-## set a default spectral index
-si = -0.7
-zmin = 0.003  ## matches Mauch & Sadler 2007
-zmax = 0.3   ## matches Mauch & Sadler 2007
-dz = 0.0001  ## matches Cochrane and Kondapally
-
-## read in SIMBA
-'''
-Simba_SF = Table.read( paths.static / 'RLFS_50MYR_SF.csv', format='csv', delimiter=',' )
-Simba_AGN = Table.read( paths.static / 'RLFS_50MYR_AGN.csv', format='csv', delimiter=',' )
-## shift using spectral index
-Simba_SF['x']  = Simba_SF['x'] + np.log10( np.power( (144./1400.), si ) )
-Simba_AGN['x']  = Simba_AGN['x'] + np.log10( np.power( (144./1400.), si ) )
-## and convert from mag to Lum
-# mag = -2.5 * Lum + zeropoint
-# so Lum = mag / -2.5
-Simba_SF['Curve1'] = Simba_SF['Curve1']+np.log10(2.5)
-Simba_AGN['Curve2'] = Simba_AGN['Curve2']+np.log10(2.5)
-'''
-
-## read in SIMBA and convert to same values
-Simba = Table.read( paths.static / 'RLFs_flux_lim.csv', format='csv', delimiter=',' )
[…]

### 2024-10-30 src/scripts/sfrd.py

@@ -1,245 +1,0 @@
-import paths
-import numpy as np
-import matplotlib
-from matplotlib import pyplot as plt
-from astropy.table import Table, Column, join, vstack
-from helper_functions import *
-from astropy.io import fits
-## WMAP9 is Hinshaw et al. 2013, H_0=69.3, Omega=0.287
-
-##############################################################
-## Plotting housekeeping
-
-matplotlib.rcParams['legend.frameon'] = False
-matplotlib.rcParams['axes.labelsize'] = 'large'
-## set up some colours
-n = 255
-mycols = plt.cm.viridis(np.linspace(0, 1,n))
-mycols_m = plt.cm.inferno(np.linspace(0, 1,n))
-
-##############################################################
-
-## set a default spectral index
-si = -0.7
-zmin = 0.003  ## matches Mauch & Sadler 2007
-zmax = 0.3   ## matches Mauch & Sadler 2007
-dz = 0.0001  ## matches Cochrane and Kondapally
-
-
-t = Table.read( paths.static / 'redshift_bins.csv', format='csv' )
-zbin_starts = t['zbin_starts']
-zbin_ends = t['zbin_ends']
-
-fields = ['lockman','en1']
-
-z_lum_bins = []
-z_lum_func = []
-z_agn_lum_func = []
-z_sf_lum_func = []
-z_gal_agn_lum_func = []
-z_gal_sf_lum_func = []
-for i in np.arange(0,5):
-    zmin = zbin_starts[i]
-    zmax = zbin_ends[i]
-    keep_cols = ['Total_flux_dr','Z_BEST','vmax','agn_vmax','sf_vmax','AGN_flux','SF_flux', 'Overall_class','Mass_cons','SFR_cons']
-    vmaxes = Table()
-    for field in fields:
-        infile = '{:s}_vmaxes_zmin{:s}_zmax{:s}.fits'.format(field,str(zmin),str(zmax))
-        tmp = Table.read( paths.static / infile, format='fits' )
[…]

### 2024-10-30 src/scripts/simba_comparison.py

@@ -1,97 +1,0 @@
-import paths
-import numpy as np
-import matplotlib
-from matplotlib import pyplot as plt
-from astropy.table import Table, Column, join, vstack
-from helper_functions import *
-from astropy.io import fits
-## WMAP9 is Hinshaw et al. 2013, H_0=69.3, Omega=0.287
-
-##############################################################
-## Plotting housekeeping
-
-matplotlib.rcParams['legend.frameon'] = False
-matplotlib.rcParams['axes.labelsize'] = 'large'
-## set up some colours
-n = 255
-mycols = plt.cm.viridis(np.linspace(0, 1,n))
-mycols_m = plt.cm.inferno(np.linspace(0, 1,n))
-
-##############################################################
-
-## set a default spectral index
-si = -0.7
-zmin = 0.003  ## matches Mauch & Sadler 2007
-zmax = 0.3   ## matches Mauch & Sadler 2007
-dz = 0.0001  ## matches Cochrane and Kondapally
-
-## read in SIMBA
-'''
-Simba_SF = Table.read( paths.static / 'RLFS_50MYR_SF.csv', format='csv', delimiter=',' )
-Simba_AGN = Table.read( paths.static / 'RLFS_50MYR_AGN.csv', format='csv', delimiter=',' )
-## shift using spectral index
-Simba_SF['x']  = Simba_SF['x'] + np.log10( np.power( (144./1400.), si ) )
-Simba_AGN['x']  = Simba_AGN['x'] + np.log10( np.power( (144./1400.), si ) )
-'''
-
-## new simba check
-Simba = Table.read( paths.static / 'RLFs_flux_lim.csv', format='csv', delimiter=',' )
-z_idx = np.where( Simba['z'] == 0.0 )[0]
-Simba = Simba[z_idx]
-Simba_SF = Table()
-Simba_SF.add_column( Column( Simba['logP']+np.log10( np.power( (144./1400.), si ) ), name='x' ) )
-Simba_SF.add_column( Column( Simba['SF'], name='Curve1' ) )
-Simba_AGN = Table()
-Simba_AGN.add_column( Column( Simba['logP']+np.log10( np.power( (144./1400.), si ) ), name='x' ) )
[…]

### 2024-10-30 src/scripts/testplot.py

@@ -1,30 +1,0 @@
-import paths
-import numpy as np
-import matplotlib
-from matplotlib import pyplot as plt
-from astropy.table import Table, Column, join
-## WMAP9 is Hinshaw et al. 2013, H_0=69.3, Omega=0.287
-
-si = -0.8
-
-## read in SIMBA
-Simba_SF = Table.read( paths.static / 'RLFS_50MYR_SF.csv', format='csv', delimiter=',' )
-Simba_AGN = Table.read( paths.static / 'RLFS_50MYR_AGN.csv', format='csv', delimiter=',' )
-## shift using spectral index
-Simba_SF['x']  = Simba_SF['x'] + np.log10( np.power( (144./1400.), si ) )
-Simba_AGN['x']  = Simba_AGN['x'] + np.log10( np.power( (144./1400.), si ) )
-
-fig = plt.figure( figsize=(5,5) )
-## plot simba
-plt.plot( Simba_SF['x'], Simba_SF['Curve1'], color='green', linewidth=2.5, label='Simba SF' )
-plt.plot( Simba_AGN['x'], Simba_AGN['Curve2'], color='blue', linewidth=2.5, label='Simba AGN' )
-plt.xlim((20,28))
-plt.ylim(-7.5,-2)
-plt.xlabel('log('+r'$L_{\mathrm{144 MHz}}$'+' W Hz'+r'$^{-1}$'+'])')
-plt.ylabel('log('+r'$\Phi$'+' [mag'+r'$^{-1}$'+' Mpc'+r'$^{-3}$'+'])')
-plt.legend()
-plt.savefig(paths.figures / 'test.png',dpi=300)
-fig.clear()
-
-
-

### 2024-11-01 src/scripts/calculate_RLFs.py
updates

@@ -50,9 +50,6 @@ for i in np.arange(0,5):
     lum_bins, lum_func, agn_lum_func, sf_lum_func, gal_agn_lum_func, gal_sf_lum_func = get_RLFs( vmaxes, zmin, zmax, lmin=lmin, lmax=lmax, dl=dl, si=si )
     e_agn_lum_func, e_sf_lum_func, e_gal_agn_lum_func, e_gal_sf_lum_func = random_resample( agn_lum_func, sf_lum_func, gal_agn_lum_func, gal_sf_lum_func, vmaxes, zmin, zmax, lmin=lmin, lmax=lmax, dl=dl, si=si, nsamp=1000 )
 
-    print(len(lum_bins))
-    print(len(agn_lum_func))
-
     t = Table()
     t.add_column( lum_func, name='lum_func' )
     t.add_column( agn_lum_func, name='agn_lum_func' )

### 2024-11-01 src/scripts/comparison_plot.py
typo

@@ -43,7 +43,7 @@ lockman_tmp = lockman_vmaxes[keep_cols]
 elais_tmp = elais_vmaxes[keep_cols]
 vmaxes = vstack([lockman_tmp,elais_tmp])
 
-t = Table.read( paths.data / 'rlfs/rlfs_zmin0.003_zmax0.3_lmin20.5_lmax27_dl0.3.fits', fomrat='fits' )
+t = Table.read( paths.data / 'rlfs/rlfs_zmin0.003_zmax0.3_lmin20.5_lmax27_dl0.3.fits', format='fits' )
 
 lum_bins = np.arange( lmin, lmax, dl )
 lum_func = t['lum_func']

### 2024-11-01 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 All data used in this paper is either publicly available on \href{https://lofar-surveys.org}{https://lofar-surveys.org} or from relevant citations herein, including catalogues, rms images, and star masks.
-We have prepared this manuscript using using the showyourwork (\href{https://github.com/rodluger/showyourwork}{https://github.com/rodluger/showyourwork}) open source scientific article workflow, first introduced in \cite{luger_mapping_2021}.
-This ensures complete transparency.
-The code is available at \hl{update link when ready to publish}.
+This study was carried out using showyourwork (\href{https://github.com/showyourwork/showyourwork}{https://github.com/showyourwork/showyourwork}) which is open source scientific article workflow, first introduced in \cite{luger_mapping_2021}.
+This leverages continuous integration to
+programmatically download the data from \href{https://zenodo.org/}{zenodo.org}, create the figures, and compile the manuscript.
+The code is available at \href{https://github.com/lmorabit/hidden_AGN}{https://github.com/lmorabit/hidden\_AGN} which is linked to the starting datasets for this work, which are hosted on Zenodo with their own DOIs (10.5281/zenodo.14012620, 10.5281/zenodo.14013377, 14013423, 14013439, 14014490, 14014542, 14014849, and 14014853), ensuring complete reproducibility.
+
 \section*{Acknowledgements}

### 2024-11-07 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Title
-\title[A hidden AGN population]{A hidden Active Galactic Nuclei population: the first radio luminosity functions by physical process}
+\title[A hidden AGN population]{A hidden Active Galactic Nuclei population: the first radio luminosity functions constructed by physical process}
 
@@
 Both star formation (SF) and Active Galactic Nuclei (AGN) play an important role in galaxy evolution.
-Quantifying their relative importance in a statistical way can be done using radio luminosity functions.
-Until now these have relied on galaxy classifications, where sources with a mixture of radio emission from SF and AGN are labelled as either a star-forming galaxy or an AGN.
-This can lead to the under- or over-estimation of the relevance of AGN.
-Brightness temperature methods were demonstrated to be effective at separating radio emission from AGN and SF at 144 MHz using sub-arcsecond imaging with the International LOFAR Telescope (ILT).
-We use the combination of sub-arcsecond and arcsecond resolution imaging of 7,497 sources in the Lockman Hole and ELAIS-N1 fields to identify AGN components in the high resolution images and subtract them from the total flux density, leaving flux density from SF only.
-We construct, for the first time, radio luminosity functions by physical process, either SF or AGN activity, revealing a hidden AGN population at \llof $<10^{24}$\wphz .
+Statistically quantifying their relative importance can be done using radio luminosity functions.
+Until now these relied on galaxy classifications, where sources with a mixture of radio emission from SF and AGN are labelled as either a star-forming galaxy or an AGN.
+This can cause the misestimation of the relevance of AGN.
+Brightness temperature measurements at 144 MHz with the International LOFAR telescope can separate radio emission from AGN and SF.
+We use the combination of sub-arcsec and arcsec resolution imaging of 7,497 sources in the Lockman Hole and ELAIS-N1 fields to identify AGN components in the sub-arcsec resolution images and subtract them from the total flux density, leaving flux density from SF only.
+We construct, for the first time, radio luminosity functions by physical process, either SF or AGN activity, revealing a hidden AGN population at \llof$<10^{24}$\wphz .
 This population is 1.56$\pm$0.06 more than expected for $0.5<z<2.0$ when comparing to RLFs by galaxy classification.
-The star forming population is less impacted, with only 0.90$\pm$0.02 of the expected star formation.
-The contribution of `hidden' AGN to the overall radio population, especially at \llof $<10^{24}$\wphz , can have significant implications for the cosmic star formation rate and kinetic luminosity densities.
+The star forming population has only 0.90$\pm$0.02 of the expected SF.
+These ‘hidden’ AGN can have significant implications for the cosmic star formation rate and kinetic luminosity densities.
 \end{abstract}
